---
output:
  html_document:
  theme: yeti
pdf_document: default
editor_options: 
  chunk_output_type: console
---

# Rethinking: Chapter 7

**Ulysses' Compass**

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "#>", 
                      dev = "svg",         # for html rendering
                      # dev = "cairo_pdf", # for pdf rendering
                      fig.asp = .5, 
                      fig.align = "center",
                      message = FALSE,
                      warning = FALSE)

set.seed(42)
source("bayesian_settings.R")
source("knitr_matrix.R")
```

by [Richard McElreath](https://xcelab.net/rm/statistical-rethinking/), building on the Summaries by [Solomon Kurz](https://bookdown.org/content/4857/) and [Jake Thompson](https://sr2-solutions.wjakethompson.com/linear-models.html).

```{r, echo = FALSE, out.width = 250, fig.cap = "<i>Between Scylla and Charybdis</i> by <a href='https://commons.wikimedia.org/wiki/File:Adolf_Hiremy-Hirschl_-_Zwischen_Skylla_und_Charybdis.jpg'>Adolf Hirémy-Hirschl</a> (1910). "}
knitr::include_graphics("img/Skylla_and_Charybdis.svg")
```

## The Problem with Parameters

```{r data_brainsize, fig.asp = .55, fig.width = 7.5}
library(rethinking)
data_brainsize <- tibble(
  species = c("afarensis", "africanus", "habilis", "boisei", "rudolfensis", "ergaster", "sapiens"), 
  brain_size = c(438, 452, 612, 521, 752, 871, 1350), 
  mass = c(37.0, 35.5, 34.5, 41.5, 55.5, 61.0, 53.5)) %>% 
  mutate(brain_size_scl = brain_size/ max(brain_size),
         mass_std = standardize(mass))

data_brainsize %>% 
  ggplot(aes(x = mass, y = brain_size)) +
  geom_point(shape = 21, color = clr2, fill = fll2, size = 3) +
  ggrepel::geom_text_repel(aes(label = species),
                           force = 30, min.segment.length = unit(.1, "npc"),
                           family = fnt_sel, fontface = "italic") +
  coord_fixed(ratio = .03)
```

### The burial of R<sup>2</sup>

$$
R^{2} = \frac{var(outcome) - var(residuals)}{var(outcome)} = 1 - \frac{var(residuals)}{var(outcome)}
$$

#### Linear Model

$$
\begin{array}{rclr}
b_{i} & \sim & Normal( \mu_i, \sigma) & \textrm{[likelihood]}\\
\mu_i & = & \alpha + \beta_{m} m_{i} & \textrm{[linear model]}\\
\alpha & \sim & Normal(0.5, 1) & \textrm{[$\alpha$ prior]}\\
\beta_{m} & \sim & Normal(0, 10) & \textrm{[$\beta_{T}$ prior]}\\
\sigma & \sim & Log-Normal(0, 1) & \textrm{[$\sigma$ prior]}
\end{array}
$$

```{r}
model_brain_size <- quap(
  flist = alist(
    brain_size_scl ~ dnorm(mu, exp(log_sigma)),
    mu <- alpha + beta_m * mass_std,
    alpha ~ dnorm(0.5, 1),
    beta_m ~ dnorm(0, 10),
    log_sigma ~ dnorm( 0, 1 )
  ),
  data = data_brainsize
)

precis(model_brain_size) %>% 
  knit_precis()
```


```{r}
extract_r2 <- function(quap_fit, decimals = 5){
  data <- sim(quap_fit) %>% 
  as_tibble() %>% 
  set_names(nm = data_brainsize$species) %>% 
  summarise(across(everything(), mean, .names = "{.col}")) %>% 
  pivot_longer(everything(), names_to = "species", values_to = "mean_brainsize") %>% 
  mutate(brain_size_scl = data_brainsize$brain_size_scl,
         diff = mean_brainsize - brain_size_scl)

  round(1 - var2(data$diff) / var2(data$brain_size_scl), digits = decimals)
}

set.seed(12)
extract_r2(model_brain_size)
```

### Higher order polynomials

$$
\begin{array}{rclcr} 
b_i & {\sim} & Normal(\mu_i, \sigma) & &\textrm{[likelihood]}\\
\mu_i & = & \alpha + \beta_{1} m_{i} + \beta_{2} m_{i}^2 & &\textrm{[linear model]}\\
\alpha & \sim & Normal(0.5, 1) & & \textrm{[$\alpha$ prior]}\\
\beta_{j} & \sim & Normal(0, 10) & \textrm{for}~j = 1..2 & \textrm{[$\beta$ prior]}\\
\sigma & \sim & Log-Normal(0, 1) & &\textrm{[$\sigma$ prior]}
\end{array}
$$

```{r higher_order_polynomials}
model_brain_size2 <- quap(
  flist = alist(
    brain_size_scl ~ dnorm(mu, exp(log_sigma)),
    mu <- alpha + 
      beta_m[1] * mass_std + 
      beta_m[2] * mass_std ^ 2,
    alpha ~ dnorm(0.5, 1),
    beta_m ~ dnorm(0, 10),
    log_sigma ~ dnorm( 0, 1 )
  ),
  data = data_brainsize,
  start = list(beta_m = rep(0, 2))
)

model_brain_size3 <- quap(
  flist = alist(
    brain_size_scl ~ dnorm(mu, exp(log_sigma)),
    mu <- alpha + 
      beta_m[1] * mass_std + 
      beta_m[2] * mass_std ^ 2 + 
      beta_m[3] * mass_std ^ 3,
    alpha ~ dnorm(0.5, 1),
    beta_m ~ dnorm(0, 10),
    log_sigma ~ dnorm( 0, 1 )
  ),
  data = data_brainsize,
  start = list(beta_m = rep(0, 3))
)

model_brain_size4 <- quap(
  flist = alist(
    brain_size_scl ~ dnorm(mu, exp(log_sigma)),
    mu <- alpha + 
      beta_m[1] * mass_std + 
      beta_m[2] * mass_std ^ 2 + 
      beta_m[3] * mass_std ^ 3 + 
      beta_m[4] * mass_std ^ 4,
    alpha ~ dnorm(0.5, 1),
    beta_m ~ dnorm(0, 10),
    log_sigma ~ dnorm( 0, 1 )
  ),
  data = data_brainsize,
  start = list(beta_m = rep(0, 4))
)

model_brain_size5 <- quap(
  flist = alist(
    brain_size_scl ~ dnorm(mu, exp(log_sigma)),
    mu <- alpha + 
      beta_m[1] * mass_std + 
      beta_m[2] * mass_std ^ 2 + 
      beta_m[3] * mass_std ^ 3 + 
      beta_m[4] * mass_std ^ 4 + 
      beta_m[5] * mass_std ^ 5,
    alpha ~ dnorm(0.5, 1),
    beta_m ~ dnorm(0, 10),
    log_sigma ~ dnorm( 0, 1 )
  ),
  data = data_brainsize,
  start = list(beta_m = rep(0, 5))
)

model_brain_size6 <- quap(
  flist = alist(
    brain_size_scl ~ dnorm(mu, exp(log_sigma)),
    mu <- alpha + 
      beta_m[1] * mass_std + 
      beta_m[2] * mass_std ^ 2 + 
      beta_m[3] * mass_std ^ 3 + 
      beta_m[4] * mass_std ^ 4 + 
      beta_m[5] * mass_std ^ 5 + 
      beta_m[6] * mass_std ^ 6,
    alpha ~ dnorm(0.5, 1),
    beta_m ~ dnorm(0, 10),
    log_sigma ~ dnorm( 0, 1 )
  ),
  data = data_brainsize,
  start = list(beta_m = rep(0, 6))
)
```


```{r, fig.asp = .65}
mass_seq <- seq( from = min(data_brainsize$mass_std) - .15,
                 to = max(data_brainsize$mass_std) + .15,
                 length.out = 101)

plot_poly <- function(mod, ylim){
  model_posterior_samples <- extract.samples(mod) %>% 
    as.data.frame() %>% 
    as_tibble()
  
  model_posterior_prediction_samples <- link(mod, data = tibble(mass_std = mass_seq)) %>% 
    as_tibble() %>% 
    set_names(nm = mass_seq) %>% 
    pivot_longer(cols = everything(), names_to = "mass_std", values_to = "brain_size_scl") %>% 
    mutate(mass_std = as.numeric(mass_std)) 
  
  model_posterior_prediction_pi <- model_posterior_prediction_samples %>% 
    group_by(mass_std) %>% 
    summarise(mean = mean(brain_size_scl),
              PI_lower = PI(brain_size_scl)[1],
              PI_upper = PI(brain_size_scl)[2]) %>% 
    ungroup()
  
  p <- ggplot(mapping = aes(x = mass_std * sd(data_brainsize$mass) + mean(data_brainsize$mass))) +
    geom_smooth(data = model_posterior_prediction_pi, stat = "identity",
                aes(y = mean * max(data_brainsize$brain_size),
                    ymin = PI_lower  * max(data_brainsize$brain_size),
                    ymax = PI_upper * max(data_brainsize$brain_size)),
                color = clr2, fill = fll2, size = .2) +
    geom_point(data = data_brainsize, aes(y = brain_size_scl * max(data_brainsize$brain_size)),
               color = rgb(0,0,0,.5), size = 1) +
    labs(x = "mass",
         y = "brain_size",
         title = glue("*R<sup>2</sup>:* {extract_r2(mod, decimals = 2)}")) +
    coord_cartesian(ylim = ylim) +
    theme(plot.title = element_markdown())
  
  if(identical(mod, model_brain_size6)) {
    p <- p +
    geom_hline(yintercept = 0, color = clr_dark, linetype = 3 )
  }
  
  p
}

list(model_brain_size, model_brain_size2,model_brain_size3,
     model_brain_size4, model_brain_size5, model_brain_size6) %>% 
  purrr::map2(.y = list(c(420, 1400), c(420, 1400), c(420, 1400),
                        c(300, 1950), c(300, 1950), c(-400, 1500)),
              plot_poly) %>% 
  wrap_plots() +
  plot_annotation(tag_levels = "a")
```

### Underfitting

*Leave one out* (LOO)

```{r}
model_loo <- function(idx = 0, mod_degree = 1){
  data <- data_brainsize[-idx, ]
  
  if(mod_degree == 1){
    current_mod <- quap(
      flist = alist(
        brain_size_scl ~ dnorm(mu, exp(log_sigma)),
        mu <- alpha + 
          beta_m[1] * mass_std,
        alpha ~ dnorm(0.5, 1),
        beta_m ~ dnorm(0, 10),
        log_sigma ~ dnorm( 0, 1 )
      ),
      data = data,
      start = list(beta_m = rep(0, 1))
    )
  } else if(mod_degree == 4) {
    current_mod <- quap(
      flist = alist(
        brain_size_scl ~ dnorm(mu, exp(log_sigma)),
        mu <- alpha + 
          beta_m[1] * mass_std + 
          beta_m[2] * mass_std ^ 2 + 
          beta_m[3] * mass_std ^ 3 + 
          beta_m[4] * mass_std ^ 4,
        alpha ~ dnorm(0.5, 1),
        beta_m ~ dnorm(0, 10),
        log_sigma ~ dnorm( 0, 1 )
      ),
      data = data,
      start = list(beta_m = rep(0, 4))
    )
  } else { stop("`mod_degree` needs to be either 1 or 4") }
  
  model_posterior_prediction_samples <- link(current_mod, data = tibble(mass_std = mass_seq)) %>% 
    as_tibble() %>% 
    set_names(nm = mass_seq) %>% 
    pivot_longer(cols = everything(), names_to = "mass_std", values_to = "brain_size_scl") %>% 
    mutate(mass_std = as.numeric(mass_std)) 
  
  model_posterior_prediction_pi <- model_posterior_prediction_samples %>% 
    group_by(mass_std) %>% 
    summarise(mean = mean(brain_size_scl)) %>% 
    ungroup() %>% 
    mutate(idx = idx, mod_degree = mod_degree)
  
  model_posterior_prediction_pi
}

cross_df(list(idx = seq_along(data_brainsize$species),
              mod_degree = c(1, 4))) %>% 
  pmap_dfr(model_loo) %>% 
  ggplot(mapping = aes(x = mass_std * sd(data_brainsize$mass) + mean(data_brainsize$mass))) +
    geom_line(aes(y = mean * max(data_brainsize$brain_size),
                    group = factor(idx),
                  color = idx)) +
    geom_point(data = data_brainsize %>% mutate(idx = row_number()),
               aes(y = brain_size_scl * max(data_brainsize$brain_size),
                   fill = idx, color = after_scale(clr_darken(fill))),
               size = 2, shape = 21) +
    labs(x = "mass",
         y = "brain_size") +
  facet_wrap(mod_degree ~ ., labeller = label_both) +
  scale_color_gradientn(colors = c(clr0dd, clr0, clr2), guide = "none") +
  scale_fill_gradientn(colors = c(clr0dd, clr0, clr2), guide = "none") +
  coord_cartesian(ylim = c(0, 2e3)) +
  theme(plot.title = element_markdown())
```

## Entropy and Accuracy

### Entropy

Definition of *Information Entropy*

$$
H(p) = - E~\textrm{log}(p_{i}) = - \sum_{i = 1}^n p_{i}~\textrm{log}(p_{i})
$$

or verbally:

> *The uncertainty contained in a probability distribution is the average log-probability of an event.*

which fulfills the requirements:

- uncertainty should be continuous
- uncertainty should increase with the number of possible events
- uncertainty should be additive

Example for $p_1 = 0.3$ and $p_2 = 0.7$:

$$
H(p) = - \big( p_{1} \textrm{log}(p_{1}) + p_{2} \textrm{log}(p_{2}) \big) \approx 0.61
$$

```{r}
p <- c( .3, .7 )
- sum( p * log(p) )
```

Compared to Abu Dhabi (*"it hardly ever rains"*)

```{r}
p <- c( .01, .99 )
- sum( p * log(p) )
```

Entropy increases with th dimensionality of the prediction problem (eg. predicting `r emo::ji("cloud_with_rain")`/ `r emo::ji("cloud_with_snow")` / `r emo::ji("sunny")`)

```{r}
p <- c( .15, .5, .7 )
- sum( p * log(p) )
```

### Accuracy

> *Divergence: The additional uncertainty induced by using probabilities from one distribution to describe another distribution.*

The *Kullback-Leibler Divergence* (KL):

$$
D_{KL}(p, q) = \sum_{i} p_{i} \big( \textrm{log}(p_{i}) - \textrm{log}(q_{i})\big) = \sum_{i} p_{i} \textrm{log}\left(\frac{p_{i}}{q_{i}}\right)
$$

```{r, fig.asp = .75}
tibble(p1 = .3,
       p2 = .7,
       q1 = seq(from = .01, to = .99, by = .01),
       q2 = 1 - q1,
       d_kl = p1 * log(p1 / q1) + p2 * log(p2 / q2)) %>% 
  ggplot(aes(x = q1, y = d_kl)) +
  geom_line(color = clr2) +
  geom_vline(xintercept = .3, color = clr_dark, linetype = 3)
```

### Estimating Divergence

*Log-probability score* to compare the predictive accuracy of different models:

$$
S(q) = \sum_{i} \textrm(log) (q_{i})
$$

where $i$ indexes each case and $q_{i}$ is the likelihood for each case.

A (re-scaled) equivalent is given with the *deviance*:

$$
D(q) = -2 \sum_{i} \textrm(log) (q_{i})
$$

and it's Bayesian version the *Log-posterior-predictive density*:

$$
lppd(y, \Theta) = \sum_{i} \textrm{log} \frac{1}{S} \sum_{s} p (y_{i} | \Theta_{s}) 
$$

where $S$ is the number of samples and $\Theta_{s}$ is the s-th set of sampled parameter values in the posterior distribution.

```{r}
# lppd <- function (fit, ...) {
#     ll <- sim(fit, ll = TRUE, ...)
#     n <- ncol(ll)
#     ns <- nrow(ll)
#     f <- function(i) log_sum_exp(ll[, i]) - log(ns)
#     lppd <- sapply(1:n, f)
#     return(lppd)
# }
set.seed(1)
lppd(model_brain_size, n = 1e4)
```

### Scoring the right data

```{r}
tibble(model_degree = 1:6,
       model = list(model_brain_size, model_brain_size2,
                    model_brain_size3, model_brain_size4,
                    model_brain_size5, model_brain_size6)) %>% 
  mutate(log_prob_score = map_dbl(model, .f = function(mod){sum(lppd(mod))}))
```


```{r, eval = FALSE}
n_cores <- 8

run_sim <- function(k, n_samples, n_sim = 1e3, b_sigma = 100){
  mcreplicate(n_sim, sim_train_test(N = n_samples, k = k, b_sigma = b_sigma),
              mc.cores = n_cores) %>% 
    t() %>% 
    as_tibble() %>% 
    summarise(mean_p = mean(V1),
              mean_q = mean(V2),
              sd_p = sd(V1),
              sd_q = sd(V2)) %>% 
    mutate(k = k, n_samples = n_samples, b_sigma = b_sigma)
}

tictoc::tic()
data_sim <- cross_df(list(k = 1:5,
                          n_samples = c(20, 100))) %>% 
  pmap_dfr(run_sim)
tictoc::toc()
write_rds(data_sim, "data/rethinking_c6_data_sim.Rds")

tictoc::tic()
data_sim_var_beta <- crossing(k = 1:5,
         n_samples = c(20, 100),
         b_sigma = c(1, 0.5, 0.2)) %>% 
  pmap_dfr(run_sim)
tictoc::toc()
write_rds(data_sim_var_beta, "data/rethinking_c6_data_sim_var_beta.Rds")
```

```{r}
data_sim <- read_rds("data/rethinking_c6_data_sim.Rds")
x_dodge <- .3
data_sim %>% 
  ggplot() +
  geom_pointrange(aes(x = k - .5 * x_dodge,
                      ymin = mean_p - sd_p, y = mean_p, ymax = mean_p + sd_p,
                      color = "train", fill = after_scale(clr_lighten(color))),
                  shape = 21) +
    geom_pointrange(aes(x = k + .5 * x_dodge,
                      ymin = mean_q - sd_q, y = mean_q, ymax = mean_q + sd_q,
                      color = "test", fill = after_scale(clr_lighten(color))),
                  shape = 21) +
  scale_color_manual("", values = c(train = clr0dd, test = clr2)) +
  facet_wrap(n_samples ~ ., scales = "free", label = label_both) +
  labs(x = "number of parameters", y = "deviance") +
  theme(legend.position = "bottom")
```


```{r, fig.asp = .45}
p_curves <- ggplot() +
  stat_function(fun = function(x){dnorm(x = x, mean = 0, sd = .2)},
                color = clr0dd, linetype = 1, xlim = c(-3, 3), n = 301) +
  stat_function(fun = function(x){dnorm(x = x, mean = 0, sd = .5)},
                color = clr0dd, linetype = 2, xlim = c(-3, 3), n = 301) +
  stat_function(fun = function(x){dnorm(x = x, mean = 0, sd = 1)},
                color = clr0dd, linetype = 3, xlim = c(-3, 3), n = 501) +
  labs(x = "parameter value", y = "density")

data_sim_var_beta <- read_rds("data/rethinking_c6_data_sim_var_beta.Rds")

p_lines <- data_sim_var_beta %>%
  dplyr::select(mean_p, mean_q,k:b_sigma) %>% 
  pivot_longer(cols = mean_p:mean_q, names_to = "set", values_to = "mean") %>% 
  ggplot() +
  geom_line(aes(x = k,
                y = mean,
                linetype = factor(b_sigma),
                color = set)) +
  geom_point(data = data_sim %>%
               dplyr::select(mean_p, mean_q,k:b_sigma) %>% 
               pivot_longer(cols = mean_p:mean_q, names_to = "set", values_to = "mean"),
             aes(x = k,
                 y = mean,
                 color = set,
                 fill = after_scale(clr_lighten(color))),
             shape = 21, size = .9) +
  scale_color_manual("", 
                     values = c(mean_p = clr0dd, mean_q = clr2),
                     labels = c(mean_p = "test", mean_q = "train")) +
  scale_linetype_manual("beta prior width",
                        values = c(`1` = 3, `0.5` = 2, `0.2` = 1)) +
  facet_wrap(n_samples ~ ., scales = "free", label = label_both) +
  labs(x = "number of parameters", y = "deviance") +
  theme(legend.position = "bottom")

p_curves + p_lines + 
  plot_annotation(tag_levels = "a") +
  plot_layout(widths = c(.5, 1), guides = "collect") &
  theme(legend.position = "bottom")
```

## Golem taming: regularization

### Cross-validataion

- *leave-one-out cross-validataion* (loocv): dropping one data point in each *fold* (resulting in $n$ test sets, so $n$ refits of the model and $n$ posterior distributions)
- *Pareto-smoothed importance sampling cross-validataion* (PSIS): approximates loocv by sampling from the original posterior while taking the *importance*/*weight* of each data point into account (provides feedback about it's ow reliability, no model re-fitting necessary)

$$
s_{PSIS} = \sqrt{N~\textrm{var}(\textrm{psis}_{i})}
$$

The importance sampling estimate of out-of-sample lppd:

$$
lppd_{IS} = \sum_{i=1}^{N} \textrm{log} \frac{\Sigma_{s=1}^{S} r(\theta_{s}p(y_{i}|\theta_{s}))}{\sigma_{s=1}^{S} r(\theta_{s})}
$$

The *Pareto distribution*

$$
p(r | u, \sigma, k) = \sigma^{-1} \big(1 + k (r - u) \sigma ^{-1}\big)^{-\frac{1}{k}-1}
$$

```{r, fig.asp = .4}
ggplot() +
  stat_function(fun = function(x){dpareto(x = x, xmin = 1, alpha = 1)},aes(color = "1"), xlim = c(0, 5), n = 201)+
  stat_function(fun = function(x){dpareto(x = x, xmin = 1, alpha = .7)},aes(color = "0.7"), xlim = c(0, 5), n = 201)+
  stat_function(fun = function(x){dpareto(x = x, xmin = 1, alpha = .3)},aes(color = "0.3"), xlim = c(0, 5), n = 201)+
  stat_function(fun = function(x){dpareto(x = x, xmin = 1, alpha = .1)},aes(color = "0.1"), xlim = c(0, 5),  n = 201) +
  scale_color_manual("alpha", values = c(`1` = clr0dd, `0.7` = clr1, `0.3` = clr2, `0.1` = clr3)) +
  coord_cartesian(ylim = c(-.1, 15), expand = 0) +
  theme(legend.position = "bottom")
```

### Information Criteria

#### Akaike information criterion

Only for legacy reasons

$$
AIC = D_{train} + 2p = -2 lppd + 2p
$$

AIC is an approximation that depends on 

- flat priors
- posterior distribution is $\sim$ gaussian
- sample size $N$ is much greater than numbers of parameters $k$

#### Widely Applicable Information Criterion

$$
WAIC(y, \Theta) = -2 (lppd - \underbrace{\sum_{i} \textrm{var}_{\theta}~\textrm{log}~p(y_{i}|\theta)}_{\textrm{penalty term}})
$$

WAIC calculations

```{r}
data(cars)

set.seed(94)
model_cars <- quap(
  flist = alist(
    dist ~ dnorm(mu, sigma),
    mu <- alpha + beta_speed * speed,
    alpha ~ dnorm(0, 100),
    beta_speed ~ dnorm(0, 10),
    sigma ~ dexp(1)
  ),
  data = cars
)

set.seed(94)
n_samples <- 1e3
n_cases <-  nrow(cars)

cars_posterior_predictive_samples <- extract.samples(model_cars, n =  n_samples) %>% 
  as_tibble()

logprob <- sapply(1:n_samples,
                  function(idx){ 
  mu <- cars_posterior_predictive_samples$alpha[idx] + cars_posterior_predictive_samples$beta_speed[idx] * cars$speed
  dnorm( cars$dist, mean = mu, sd = cars_posterior_predictive_samples$sigma[idx], log = TRUE)})

lppd <- sapply(1:n_cases, function(i){log_sum_exp(logprob[i,]) - log(n_samples)})
pWAIC <- sapply(1:n_cases, function(i){var(logprob[i,])})

-2 * (sum(lppd) - sum(pWAIC))

waic_vec <- -2 * (lppd - pWAIC)
sqrt(n_cases * var(waic_vec))  
```


### Comparing CV, PSIS and WAIC

```{r, eval = FALSE}
make_sim <- function(n, k, b_sigma) {
  
  r <- mcreplicate(n_sim, 
                   sim_train_test(N = n,
                                  k = k,
                                  b_sigma = b_sigma,
                                  WAIC = TRUE,
                                  LOOCV = TRUE, 
                                  LOOIC = TRUE),
                   mc.cores = n_cores)
  
  t <- tibble(
      deviance_os = mean(unlist(r[2, ])),
      deviance_w = mean(unlist(r[3, ])),
      deviance_p = mean(unlist(r[11, ])),
      deviance_c = mean(unlist(r[19, ])),
      error_w = mean(unlist(r[7, ])),
      error_p = mean(unlist(r[15, ])),
      error_c = mean(unlist(r[20, ]))
      )
  
  return(t)
}

n_sim <- 1e3
n_cores <- 8

tictoc::tic()
data_sim_scores <- crossing(n = c(20, 100),
                            k = 1:5,
                            b_sigma = c(0.5, 100)) %>%
  mutate(sim = pmap(list(n, k, b_sigma), make_sim)) %>% 
  unnest(sim)
tictoc::toc()
# 119984.727 sec elapsed

write_rds(data_sim_scores, "data/rethinking_c6_data_sim_scores.Rds")
```

```{r}
data_sim_scores <- read_rds("data/rethinking_c6_data_sim_scores.Rds")

data_sim_scores %>% 
  pivot_longer(deviance_w:deviance_c) %>% 
  mutate(criteria = ifelse(name == "deviance_w", "WAIC",
                           ifelse(name == "deviance_p", "PSIS", "CV"))) %>% 
  ggplot(aes(x = k)) +
  geom_line(aes(y = value, color = criteria)) +
  geom_point(aes(y = deviance_os, shape = factor(b_sigma)), size = 1.5) +
  scale_shape_manual("sigma", values = c(19, 1)) +
  scale_color_manual(values = c(clr0dd, clr1, clr2)) +
  labs(x = "number of parameters (k)",
       y = "average deviance") +
  facet_grid(n ~ b_sigma, scales = "free_y", labeller = label_both)+
  theme(legend.position = "bottom")
```

## Model comparison

### Model mis-selection

```{r}
chapter6_models <- read_rds("envs/chapter6_models.rds")

set.seed(11)
WAIC(chapter6_models$model_fungus_post_treatment)
```

```{r}
set.seed(77)
comp_waic <- compare(chapter6_models$model_fungus_post_treatment,
        chapter6_models$model_fungus_no_treatment,
        chapter6_models$model_fungus_only_treatment,
        func = WAIC) %>% 
  as_tibble_rn() %>% 
  mutate(model = str_remove(param,".*\\$"),
         mod = model %>%  purrr::map(.f = function(m){chapter6_models[[m]]}),
         deviance = mod %>% 
          purrr::map_dbl(.f = rethinking::deviance),
        model = str_remove(model,"model_fungus_")) %>% 
  dplyr::select(-param)

comp_waic %>% 
  dplyr::select(-mod) %>%
  dplyr::select(model, everything())
```

WAIC and PSIS result in similar values:

```{r}
compare(chapter6_models$model_fungus_post_treatment,
        chapter6_models$model_fungus_no_treatment,
        chapter6_models$model_fungus_only_treatment,
        func = PSIS) %>% 
  as_tibble_rn()
```

```{r}
tibble(post_treatment = WAIC(chapter6_models$model_fungus_post_treatment, pointwise = TRUE)$WAIC,
       only_treatment = WAIC(chapter6_models$model_fungus_only_treatment, pointwise = TRUE)$WAIC,
       no_treatment = WAIC(chapter6_models$model_fungus_no_treatment, pointwise = TRUE)$WAIC) %>% 
  mutate(model_difference_post_only = post_treatment - only_treatment,
         model_difference_no_only = no_treatment - only_treatment) %>% 
  summarise(`post-only` = sqrt(n()[[1]] * var(model_difference_post_only)),
            `no-only` = sqrt(n()[[1]] * var(model_difference_no_only))) %>% 
  pivot_longer(everything(),
               names_to = "comparison",
               values_to = "se_of_model_differnce")
```

$\rightarrow$ compare to WAIC table `$dSE[[2]]` 

estimating the model difference for a *z-score* of $\sim$ 2.6 (99%)"

```{r}
comp_waic$dWAIC[[2]] + c(-1, 1) * comp_waic$dSE[[2]] * 2.6
```

```{r, fig.asp = .35}
library(tidybayes)
comp_waic %>% 
  ggplot() +
  geom_vline(xintercept = comp_waic$WAIC[[1]], color = clr_dark, linetype = 3) +
  geom_pointinterval(aes(y = model, x = WAIC, xmin = WAIC - SE, xmax = WAIC + SE,
                      color = "WAIC", fill = after_scale(clr_lighten(color))), 
                     shape = 21, size = 2.5) +
  geom_point(aes(y = model, x = deviance, color = "WAIC")) +
  geom_pointinterval(data = comp_waic %>% filter(row_number() > 1),
                     aes(y = as.numeric(factor(model)) + .25, 
                         x = WAIC, xmin = WAIC - dSE, xmax = WAIC + dSE,
                         color = "dSE"), 
                     shape = 17, size = .9) +
  scale_color_manual(values = c(WAIC = clr2, dSE = clr0dd)) +
  theme(legend.position = "bottom")
```


```{r}
set.seed(93)
compare(chapter6_models$model_fungus_post_treatment,
        chapter6_models$model_fungus_no_treatment,
        chapter6_models$model_fungus_only_treatment)@dSE %>%
  round(digits = 2) %>% 
  as_tibble() %>% 
  set_names(., nm =  names(.) %>% str_remove(pattern = ".*model_fungus_")) %>% 
  mutate(` ` = comp_waic$model) %>% 
  dplyr::select(` `, everything()) %>% 
  knitr::kable()
```

Weight of a model (last column of `compare()`, the relative support for each model):

$$
w_{i} = \frac{\textrm{exp}(-0.5\Delta_{i})}{\Sigma_{j}\textrm{exp}(-0.5\Delta_{j})}
$$

These weights are important for *model averaging*.

### Outliers and other illusions

```{r, fig.asp = .4}
chapter5_models <- read_rds("envs/chapter5_models.rds")

set.seed(24071847)
compare(chapter5_models$model_age,
        chapter5_models$model_marriage,
        chapter5_models$model_multiple) %>% 
  as_tibble_rn()

psis_k <- tibble(waic_penalty = (function(){set.seed(set.seed(23)); WAIC(chapter5_models$model_multiple, pointwise = TRUE)$penalty})(),
       psis_k = (function(){set.seed(set.seed(23)); PSIS(chapter5_models$model_multiple, pointwise = TRUE)$k})(),
       location = chapter5_models$data_waffle$Loc)

p_psis_k <- psis_k %>% 
  ggplot(aes(x = psis_k ,y = waic_penalty)) +
  geom_vline(xintercept = .5, color = clr_dark, linetype = 3) +
  geom_point(shape = 21, size = 2, color = clr2, fill = fll2) +
  geom_text(data = psis_k %>%  filter(location %in% c("ME", "ID")),
            aes(x = psis_k - .15, label = location))

p_dens <- ggplot() +
  stat_function(fun = function(x){dnorm(x = x, sd = .8)}, xlim = c(-4, 4), n = 201, aes(color = "gaussian"), linetype = 3)+
  stat_function(fun = function(x){dstudent(x = x, nu = 2, sigma = .55)}, xlim = c(-4, 4), n = 201, aes(color = "student t")) +
  labs(y = "density", x = "value")

p_logdens <- ggplot() +
  stat_function(fun = function(x){-log(dnorm(x = x, sd = .8))}, xlim = c(-4, 4), n = 201, aes(color = "gaussian"), linetype = 3)+
  stat_function(fun = function(x){-log(dstudent(x = x, nu = 2, sigma = .55))}, xlim = c(-4, 4), n = 201, aes(color = "student t")) +
  labs(y = "- log density", x = "value")


p_psis_k + p_dens + p_logdens +
  plot_layout(guides = "collect") &
  scale_color_manual("distribution", values = c(gaussian = clr0dd, `student t` = clr2)) &
  theme(legend.position = "bottom")
```


```{r}
chapter5_models$model_multiple

model_multiple_sudent <- quap(
  flist = alist(
    divorce_std ~ dstudent(2, mu, sigma),
    mu <- alpha + beta_M * marriage_std + beta_A * median_age_std,
    alpha ~ dnorm(0, 0.2),
    beta_A ~ dnorm(0, 0.5),
    beta_M ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = chapter5_models$data_waffle
)
```


```{r}
PSIS(chapter5_models$model_multiple)
```

With the *Student T* distribution as a the likelihood, $k$ is reduced as there is more mass in the tails of the distribution (thus, Idaho is less *surprising*)

```{r}
PSIS(model_multiple_sudent)
```


```{r}
precis(chapter5_models$model_multiple) %>% 
  knit_precis()

precis(model_multiple_sudent) %>% 
  knit_precis()
```

Also, as the influence of Idaho is reduced, the estimate of `beta_A` is decreased in the updated model.

```{r env_export}
library(rlang)
chapter7_models <- env(
  model_brain_size = model_brain_size
)
```

## Homework

**E1**

```{r E1}

```

**E2**

```{r E2}

```

**E3**

```{r E3}

```

**E4**

```{r E4}

```

**M1**

```{r M1}

```

**M2**

```{r M2}

```

**M3**

```{r M3}

```

**M4**

```{r M4}

```

**M5**

```{r M5}

```

**M6**

```{r M6}

```

**H1**

```{r H1}

```

**H2**

```{r H2}

```

**H3**

```{r H3}

```

**H4**

```{r H4}

```

**H5**

```{r H5}

```

## {brms} section

## pymc3 section

---

<div id="myModal" class="modal">
  <span class="close">&times;</span>
  <img class="modal-content" id="img01">
  <div id="caption"></div>
</div>

<script src="./js/zoom.js"></script>