---
output:
  html_document:
  theme: yeti
pdf_document: default
editor_options: 
  chunk_output_type: console
---

# Rethinking: Chapter 11

**God Spiked the Integers**

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "#>", 
                      dev = "svg",         # for html rendering
                      # dev = "cairo_pdf", # for pdf rendering
                      fig.asp = .5, 
                      fig.align = "center",
                      message = FALSE,
                      warning = FALSE)

source("bayesian_settings.R")
source("knitr_matrix.R")
library(tidybayes)
```

by [Richard McElreath](https://xcelab.net/rm/statistical-rethinking/), building on the Summaries by [Solomon Kurz](https://bookdown.org/content/4857/).


## Binomial Regression

### Logistic Regression: Prosocial chimpanzees

using a *logit* link function:

```{r}
library(rethinking)
data(chimpanzees)
```

translating the `prosoc_left` and the `condition` columns into the aggregated column `treatment`:

1. `prosoc_left == 0` and `condition == 0`
2. `prosoc_left == 1` and `condition == 0`
3. `prosoc_left == 0` and `condition == 1`
4. `prosoc_left == 1` and `condition == 1`

```{r}
data_chimp <- chimpanzees %>% 
  as_tibble() %>% 
  mutate(treatment = 1 + prosoc_left + 2 * condition,
         side_idx = prosoc_left + 1,     # right 1, left 2
         condition_idx = condition + 1)  # no partner 1, partner 2

xtabs(~ treatment + prosoc_left + condition, data_chimp)
```

$$
\begin{array}{rclr}
L_{i} & \sim & Binomial(1, p_{i}) & \textrm{[likelihood]}\\
\textrm{logit}(p_{i}) & = & \alpha_{ACTOR[i]} + \beta_{TREATMENT[i]} & \textrm{[linear model]}\\
\alpha_{j} & \sim & ... & \textrm{[$\alpha$ prior]}\\
\beta_{k} & \sim & ... & \textrm{[$\beta$ prior]}
\end{array}
$$

$$
Binomial(1, p_{i}) = Bernoulli(p_{i})
$$

Finding the right priors

$$
\begin{array}{rclr}
L_{i} & \sim & Binomial(1, p_{i}) & \textrm{[likelihood]}\\
\textrm{logit}(p_{i}) & = & \alpha & \textrm{[linear model]}\\
\alpha & \sim & Normal(0, \omega) & \textrm{[$\alpha$ prior]}\\
\end{array}
$$

Starting with $\omega = 10$ to demonstrate the effect of flat priors.

```{r}
model_omega10 <- quap(
  flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha,
    alpha ~ dnorm( 0, 10 )
  ),
  data = data_chimp
)

model_omega15 <- quap(
  flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha,
    alpha ~ dnorm( 0, 1.5 )
  ),
  data = data_chimp
)

set.seed(1999)
omega_prior <- extract.prior(model_omega10) %>% 
  as_tibble() %>%
  mutate(omega = 10) %>% 
  bind_rows(extract.prior(model_omega15) %>% 
  as_tibble() %>%
  mutate(omega = 1.5)) %>% 
  mutate(p = inv_logit(alpha))

p1 <- omega_prior %>% 
  ggplot(aes(x = p)) +
  geom_density(adjust = .15,
               aes(color = factor(omega), fill = after_scale(clr_alpha(color)))) +
  scale_color_manual("omega", values = c(`10` = clr0d, `1.5` = clr2)) +
  theme(legend.position = "bottom")
```

```{r, fig.asp = .4}
model_beta10 <- quap(
  flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha + beta[treatment],
    alpha ~ dnorm( 0, 1.5 ),
    beta[treatment] ~ dnorm(0,10)
  ),
  data = data_chimp 
)

model_beta05 <- quap(
  flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha + beta[treatment],
    alpha ~ dnorm( 0, 1.5 ),
    beta[treatment] ~ dnorm(0,0.5)
  ),
  data = data_chimp 
)

extract_logit <- function(model, beta_sd, seed = 42){
  
  set.seed(seed)

  extract.prior(model, n = 1e4) %>% 
  as.data.frame() %>% 
  as_tibble() %>%
  mutate(across(beta.1:beta.4, .fns = function(x){inv_logit(x + alpha)},
                .names = "p_{.col}"),
         beta_sd = beta_sd)
}

data_beta <- list(model_beta10, model_beta05) %>% 
  map2_dfr(.y = c(10, .5),
           .f = extract_logit) %>% 
  mutate(treatment_diff_1_2 = abs(`p_beta.1` - `p_beta.2`))

p2 <- data_beta %>% 
  ggplot(aes(x = treatment_diff_1_2)) +
  geom_density(adjust = .15,
               aes(color = factor(beta_sd), fill = after_scale(clr_alpha(color)))) +
  scale_color_manual("sd beta", values = c(`10` = clr0d, `0.5` = clr2)) +
  theme(legend.position = "bottom")

p1 + p2
```

$$
\begin{array}{rclr}
L_{i} & \sim & Binomial(1, p_{i}) & \textrm{[likelihood]}\\
\textrm{logit}(p_{i}) & = & \alpha_{ACTOR[i]} + \beta_{TREATMENT[i]} & \textrm{[linear model]}\\
\alpha_{j} & \sim & Normal(0, 1.5) & \textrm{[$\alpha$ prior]}\\
\beta_{k} & \sim & Normal(0, 0.5) & \textrm{[$\beta$ prior]}
\end{array}
$$

```{r, results = "hide"}
data_chimp_list <- data_chimp %>% 
  dplyr::select(pulled_left, actor, treatment) %>% 
  mutate(treatment = as.integer(treatment)) %>% 
  as.list()

model_chimp <- ulam(
  flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha[actor] + beta[treatment],
    alpha[actor] ~ dnorm( 0, 1.5 ),
    beta[treatment] ~ dnorm(0,0.5)
  ),
  data = data_chimp_list,
  chains = 4, cores = 4,
  log_lik = TRUE
)
```

```{r}
precis(model_chimp, depth = 2) %>% 
  knit_precis()

chimp_posterior <- extract.samples(model_chimp) %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  mutate(across(starts_with("alpha"),.fns = inv_logit, .names = "p_{.col}"))

p1 <- chimp_posterior %>% 
  dplyr::select(starts_with("p")) %>% 
  pivot_longer(everything(),
               names_to = "actor",
               values_to = "p") %>% 
  mutate(actor = str_remove(actor,"p_alpha.") %>% as.integer()) %>% 
  group_by(actor) %>% 
  summarise(p = list(quantile(p, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p) %>% 
  ggplot(aes(y = factor(actor, levels = 7:1))) +
  geom_linerange(aes(xmin = ll, xmax = hh), lwd = .2, color = clr2) +
  geom_linerange(aes(xmin = l, xmax = h), lwd = .7, color = clr2) +
  geom_point(aes(x = m),
             size = 2, shape = 21,
             color = clr2, fill = clr_lighten(clr2, .2)) +
  labs(y = "actor", x = "alpha") +
  coord_cartesian(xlim = c(0, 1))
```

```{r}
treatment_labels <- c("R|N", "L|N", "R|P", "L|P", "R|diff", "L|diff")

p2 <- chimp_posterior %>% 
  dplyr::select(starts_with("beta")) %>% 
  mutate(`beta.5` = `beta.1` - `beta.3` ,
         `beta.6` = `beta.2` - `beta.4` ) %>% 
  pivot_longer(everything(),
               names_to = "param",
               values_to = "p") %>% 
  mutate(treatment = treatment_labels[str_remove(param, "beta.") %>% as.integer()]) %>% 
  group_by(treatment) %>% 
  summarise(p = list(quantile(p, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p) %>% 
  ggplot(aes(y = factor(treatment, levels = rev(treatment_labels)))) +
  geom_rect(data = tibble(x = 1),inherit.aes = FALSE,
            aes(xmin = -Inf, xmax = Inf, ymin = .5, ymax = 2.5),
            color = clr0d, fill = fll0) +
  geom_vline(xintercept = 0, linetype = 3, color = clr_dark) +
  geom_linerange(aes(xmin = ll, xmax = hh), lwd = .2, color = clr2) +
  geom_linerange(aes(xmin = l, xmax = h), lwd = .7, color = clr2) +
  geom_point(aes(x = m),
             size = 2, shape = 21,
             color = clr2, fill = clr_lighten(clr2, .2)) +
  scale_y_discrete() +
  labs(y = "treatment", x = "beta")
```

```{r, fig.asp = .35}
p1 + p2
```

Posterior prediction plots:

```{r}
chimp_grid <- crossing(actor = 1:7, treatment = 1:4)

chimp_posterior_predictions <- link(model_chimp, data = chimp_grid) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(chimp_grid, .) %>% 
  pivot_longer(-c(actor, treatment), values_to = "pulled_left") %>% 
  dplyr::select(-name) %>% 
  group_by(actor, treatment) %>% 
  summarise(p = list(quantile(pulled_left, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p) %>% 
  mutate(type = "post. pred.",
         side = c("L", "R")[2 - (treatment == 1 | treatment == 3)],
         condition = c("N", "P")[1 + (treatment > 2)],
         lab = treatment_labels[treatment])

data_chimp %>% 
  group_by(actor, treatment) %>% 
  summarise(mean_data = mean(pulled_left),
            type = "data") %>% 
  mutate(side = c("L", "R")[2 - (treatment == 1 | treatment == 3)],
         condition = c("N", "P")[1 + (treatment > 2)],
         lab = treatment_labels[treatment]) %>% 
  ggplot(aes(x = treatment, y = mean_data)) +
  geom_hline(yintercept = .5, linetype = 3, color = clr_dark) +
  geom_line(aes(group = side), color = clr0dd) +
  geom_point(aes(shape = condition),
             color = clr0dd, fill = clr0, size = 1.8) +
  geom_text(data = . %>% filter(actor == 1),
            aes(y = mean_data - .2 *  (1.5 - as.numeric(factor(side))),
                label = lab), family = fnt_sel) +
  geom_line(data = chimp_posterior_predictions,
            aes(y = m, group = side), color = clr0dd) +
  geom_segment(data = chimp_posterior_predictions,
               inherit.aes = FALSE,
                  aes(x = treatment, xend = treatment,
                      y = ll, yend = hh),
                  color = clr0dd) +
    geom_point(data = chimp_posterior_predictions,
               inherit.aes = FALSE,
                  aes(x = treatment,
                      y = m, shape = condition),
             color = clr0dd, fill = clr0, size = 1.8) +
  scale_shape_manual(values = c("N" = 21, "P" = 19), guide = "none") +
  facet_grid(type ~ actor) +
  scale_x_discrete(expand = c(.2,.2)) +
  labs(x = NULL, y = "pulled_left") +
  lims(y = c(0,1)) +
  theme(panel.background = element_rect(color = clr0, fill = "transparent"))
```

```{r, results = 'hide'}
data_chimp_list2 <- data_chimp %>% 
  dplyr::select(pulled_left,
                actor, 
                side_idx,
                condition_idx)

model_chimp_interaction <- ulam(
  flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha[actor] + beta_s[side_idx] + beta_c[condition_idx],
    alpha[actor] ~ dnorm( 0, 1.5 ),
    beta_s[side_idx] ~ dnorm(0,0.5),
    beta_c[condition_idx] ~ dnorm(0,0.5)
  ),
  data = data_chimp_list2,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```


```{r}
compare(model_chimp,
        model_chimp_interaction, func = PSIS) %>% 
  knit_precis(param_name = "model")
```

```{r, fig.asp = .25}
chimp_grid2 <- crossing(actor = 1:7, side_idx = 1:2, condition_idx = 1:2)

chimp_interact_posterior_predictions <- link(model_chimp_interaction,
                                             data = chimp_grid2) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(chimp_grid2, .) %>% 
  pivot_longer(-c(actor, side_idx, condition_idx), values_to = "pulled_left") %>% 
  dplyr::select(-name) %>% 
  group_by(actor, side_idx, condition_idx) %>% 
  summarise(p = list(quantile(pulled_left, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p) %>% 
  mutate(treatment = factor(side_idx + 2 * (condition_idx - 1)))

chimp_interact_posterior_predictions %>% 
  ggplot(aes(x = treatment, y = m)) +
  geom_hline(yintercept = .5, linetype = 3, color = clr_dark) +
  geom_line(aes(y = m, group = factor(side_idx)), color = clr0dd) +
  geom_segment(data = chimp_posterior_predictions,
                  aes(x = treatment, xend = treatment,
                      y = ll, yend = hh),
                  color = clr0dd) +
    geom_point(aes(x = treatment,
                      y = m, shape = factor(condition_idx)),
             color = clr0dd, fill = clr0, size = 1.8) +
  scale_shape_manual(values = c(`1` = 21, `2` = 19), guide = "none") +
  facet_grid(. ~ actor) +
  scale_x_discrete(expand = c(.2,.2)) +
  labs(x = NULL, y = "pulled_left") +
  lims(y = c(0,1)) +
  theme(axis.text.x = element_blank(),
        panel.background = element_rect(color = clr0, fill = "transparent"),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.x = element_blank())
```

Adding log-probability calculations to a Stan model

```{r}
stan_posterior <- extract.samples(model_chimp, clean = FALSE)
str(stan_posterior)

model_chimp_stancode <- stancode(model_chimp)
```

```{r, results = 'hide'}
model_chimp_stan <- stan(model_code = model_chimp_stancode,
  data = data_chimp_list,
  chains = 4, cores = 4)
```

```{r}
compare(model_chimp,
        model_chimp_stan) %>% 
  knit_precis(param_name = "model")
```

### Relative Shark and Absolute Deer

Calculating the *proportional odds* relative effect size of switching from treatment two to four (adding a partner):

```{r}
(prop_odd <- extract.samples(model_chimp) %>%
  as.data.frame() %>% 
  as_tibble() %>% 
  mutate(prop_odds = exp( `beta.4` - `beta.2`)) %>% 
  summarise(average = mean(prop_odds)))
```

So this relates to a `r round(100 * (1 - prop_odd$average) , digits = 2)` % change in odds.


### Aggregated binomial: Chimpanzees again, condensed

```{r}
data_chimp_aggregated <- data_chimp %>% 
  group_by(treatment, actor, side_idx, condition_idx) %>% 
  summarise(left_pulls = sum(pulled_left))  %>% 
  ungroup() %>% 
  mutate(across(everything(), .fns = `as.integer`))
```

```{r, results = 'hide'}
model_chimp_aggregated <- ulam(
  flist = alist(
    left_pulls ~ dbinom( 18 , p ),
    logit(p) <- alpha[actor] + beta[treatment],
    alpha[actor] ~ dnorm(0, 1.5),
    beta[treatment] ~ dnorm(0, 0.5)
  ),
  data = data_chimp_aggregated,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```

```{r}
precis(model_chimp_aggregated, depth = 2) %>% 
  knit_precis(param_name = "model")

clr_current <- clr1
chimp_posterior_aggregated <- extract.samples(model_chimp_aggregated) %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  mutate(across(starts_with("alpha"),.fns = inv_logit, .names = "p_{.col}"))

p1 <- chimp_posterior_aggregated %>% 
  dplyr::select(starts_with("p")) %>% 
  pivot_longer(everything(),
               names_to = "actor",
               values_to = "p") %>% 
  mutate(actor = str_remove(actor,"p_alpha.") %>% as.integer()) %>% 
  group_by(actor) %>% 
  summarise(p = list(quantile(p, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p) %>% 
  ggplot(aes(y = factor(actor, levels = 7:1))) +
  geom_linerange(aes(xmin = ll, xmax = hh), lwd = .2, color = clr_current) +
  geom_linerange(aes(xmin = l, xmax = h), lwd = .7, color = clr_current) +
  geom_point(aes(x = m),
             size = 2, shape = 21,
             color = clr_current, fill = clr_lighten(clr_current, .2)) +
  labs(y = "actor", x = "alpha") +
  coord_cartesian(xlim = c(0, 1))
```

```{r}
p2 <- chimp_posterior_aggregated %>% 
  dplyr::select(starts_with("beta")) %>% 
  mutate(`beta.5` = `beta.1` - `beta.3` ,
         `beta.6` = `beta.2` - `beta.4` ) %>% 
  pivot_longer(everything(),
               names_to = "param",
               values_to = "p") %>% 
  mutate(treatment = treatment_labels[str_remove(param, "beta.") %>% as.integer()]) %>% 
  group_by(treatment) %>% 
  summarise(p = list(quantile(p, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p) %>% 
  ggplot(aes(y = factor(treatment, levels = rev(treatment_labels)))) +
  geom_rect(data = tibble(x = 1), inherit.aes = FALSE,
            aes(xmin = -Inf, xmax = Inf, ymin = .5, ymax = 2.5),
            color = clr0d, fill = fll0) +
  geom_vline(xintercept = 0, linetype = 3, color = clr_dark) +
  geom_linerange(aes(xmin = ll, xmax = hh), lwd = .2, color = clr_current) +
  geom_linerange(aes(xmin = l, xmax = h), lwd = .7, color = clr_current) +
  geom_point(aes(x = m),
             size = 2, shape = 21,
             color = clr_current, fill = clr_lighten(clr_current, .2)) +
  scale_y_discrete() +
  labs(y = "treatment", x = "beta")
```

```{r, fig.asp = .35}
p1 + p2
```

```{r}
compare(model_chimp,
        model_chimp_aggregated,
        func = PSIS) %>% 
  knit_precis(param_name = "model")
```

The huge difference in `PSIS` are due to the way the data is organized, which makes the *multiplicity* effect the aggregate model (eg. all they ways 6 successes in 9 trials can be arranged).

```{r}
#> deviance of aggregated 6 in 9
-2 * dbinom(x = 6 , size = 9, prob = .2, log = TRUE)

#> deviance of dis-aggregated
-2 * sum( dbern(x = rep(1:0, c(6, 3)), prob = .2, log = TRUE) )
```

```
Some Pareto k values are high (>0.5). Set pointwise=TRUE to inspect individual points.

.

Warning messages:
1: In compare(model_chimp, model_chimp_aggregated, func = PSIS) :
  Different numbers of observations found for at least two models.
Model comparison is valid only for models fit to exactly the same observations.
Number of observations for each model:
model_chimp 504 
model_chimp_aggregated 28 
```

:::beware
The high Pareto k values result from the fact that the aggregation turns the *leave one out* procedure into a *leave 18 out* procedure here.

> *If you want to calculate WAIC or PSIS, you should use a logistic regression data format, not an aggregated format.*

:::

### Aggregated binomial: Graduate School Admissions

```{r}
data(UCBadmit) 
data_ucb <- UCBadmit %>% 
  as_tibble() %>% 
  mutate(gid = 3L - as.integer(factor(applicant.gender)))
```

$$
\begin{array}{rclr}
A_{i} & \sim & Binomial(N_{i}, p_{i}) & \textrm{[likelihood]}\\
\textrm{logit}(p_{i}) & = & \alpha_{GID[i]} & \textrm{[linear model]}\\
\alpha_{j} & \sim & Normal(0, 1.5) & \textrm{[$\alpha$ prior]}\\
\end{array}
$$

```{r, results = "hide"}
data_ucb_list <- data_ucb %>% 
  dplyr::select(admit, applications,  gid) %>% 
  as.list()

model_ucb <- ulam(
  flist = alist(
    admit ~ dbinom( applications, p ),
    logit(p) <- alpha[gid],
    alpha[gid] ~ dnorm( 0, 1.5 )
  ),
  data = data_ucb_list,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```

```{r}
precis(model_ucb, depth = 2) %>% 
  knit_precis()
```

Calculating a contrast to quantify the difference in the posterior between `r emo::ji("female_sign")` and `r emo::ji("male_sign")` applicants.

```{r}
posterior_ucb <- extract.samples(model_ucb) %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  mutate(diff_absolute = `alpha.1` - `alpha.2`,
         diff_relative = inv_logit(`alpha.1`) - inv_logit(`alpha.2`))

clr_current <- clr2
precis(posterior_ucb) %>% 
  knit_precis()
```

Notice the discrepancy, between <span style="color:`r clr_current`">the model</span> and <span style="color:`r clr0dd`">the data</span>: the model expects `r emo::ji("female_sign")` applicants to do 14% worse than `r emo::ji("male_sign")`, however that is only the case in two departments (C and E).

```{r}
ucb_grid <- crossing(dept = factor(LETTERS[1:6]),
                     gid = 1:2) %>% 
  left_join(data_ucb %>% dplyr::select(dept, gid, applications))

ucb_posterior_predictions <- sim(model_ucb,
                                  data = ucb_grid) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(ucb_grid, .) %>% 
  pivot_longer(-c(dept, gid, applications), values_to = "admit") %>% 
  dplyr::select(-name) %>% 
  group_by(dept, gid, applications) %>% 
  summarise(p = quantile(admit/applications, probs = c(.055, .25, .5, .75, .955)),
            median = median(admit),
            breaks = c("ll", "l", "m", "h", "hh")) %>% 
  ungroup()  %>% 
  pivot_wider(names_from = breaks, values_from = p) %>% 
  mutate(type = "post. pred.",
         x = 2 * (as.integer(as.factor(dept)) - 1) + gid)

library(geomtextpath)
data_ucb %>% 
  group_by(dept, gid) %>% 
  summarise(mean_data = mean(admit/applications),
            type = "data") %>% 
  mutate(x = 2 * (as.integer(dept) - 1) + gid) %>% 
  ggplot(aes(x = x, y = mean_data)) +
  geom_segment(data = ucb_posterior_predictions,
               aes(x = x, xend = x,
                   y = ll, yend = hh),
               color = clr_current) +
  geom_point(data = ucb_posterior_predictions,
             aes(y = m, shape = factor(gid)),
             color = clr_current,
             fill = clr_lighten(clr_current),
             size = 1.8) +
  geom_textline(aes(label = dept, group = dept),
                color = clr0dd, family = fnt_sel) +
  geom_point(aes(shape = factor(gid)),
             color = clr0dd, fill = clr0, size = 1.8) +
  
  scale_shape_manual(values = c(`2` = 21, `1` = 19),
                     guide = "none") +
  scale_x_continuous(breaks = 2 * (1:6) - .5,
                     labels = LETTERS[1:6])+
  labs(x = "department", y = "admit") +
  lims(y = c(0,1)) +
  theme(panel.background = element_rect(color = clr0,
                                        fill = "transparent"), 
        panel.grid.major.x = element_blank())
```

$$
\begin{array}{rclr}
A_{i} & \sim & Binomial(N_{i}, p_{i}) & \textrm{[likelihood]}\\
\textrm{logit}(p_{i}) & = & \alpha_{GID[i]} + \delta_{DEPT[i]}& \textrm{[linear model]}\\
\alpha_{j} & \sim & Normal(0, 1.5) & \textrm{[$\alpha$ prior]}\\
\delta_{k} & \sim & Normal(0, 1.5) & \textrm{[$\delta$ prior]}\\
\end{array}
$$


```{r, results = "hide"}
data_ucb_list2 <- data_ucb %>% 
  mutate(dept_idx = as.numeric(dept)) %>% 
  dplyr::select(dept_idx, admit, applications,  gid) %>% 
  as.list()

model_ucb_dept <- ulam(
  flist = alist(
    admit ~ dbinom( applications, p ),
    logit(p) <- alpha[gid] + delta[dept_idx],
    alpha[gid] ~ dnorm( 0, 1.5 ),
    delta[dept_idx] ~ dnorm( 0, 1.5 )
  ),
  data = data_ucb_list2,
  iter = 4000,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```

```{r}
precis(model_ucb_dept, depth = 2) %>% 
  knit_precis()
```

```{r}
posterior_ucb_dept <- extract.samples(model_ucb_dept) %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  mutate(diff_absolute = `alpha.1` - `alpha.2`,
         diff_relative = inv_logit(`alpha.1`) - inv_logit(`alpha.2`))

precis(posterior_ucb_dept %>% dplyr::select(starts_with("diff"))) %>% 
  knit_precis(param_name = "column")
```

Now, `r emo::ji("male_sign")` applicants seem to have minimal worse chances than `r emo::ji("female_sign")` applicants (~2%).

> Why did adding departments to the model change inference about the gender so much?

```{r}
data_ucb %>% 
  group_by(dept) %>% 
  mutate(all_n = sum(applications)) %>% 
  ungroup() %>% 
  mutate(fraction = round(applications / all_n, digits = 2)) %>% 
  dplyr::select(dept, applicant.gender, fraction) %>% 
  pivot_wider(names_from = dept, values_from = fraction)
```

There is a causal path $G \rightarrow D \rightarrow A$ which is closed by conditioning on $D$ (this is an example of a **mediation analysis**).
However issues arise if there is an unobserved confound $U$.

```{r, fig.asp = .3}
dag_ucb <- dagify(
  A ~ D + G,
  D ~ G,
  exposure = "A",
  outcome = "G",
  coords = tibble(name = c("G", "D", "A"),
                  x = c(0, .5, 1),
                  y = c(0, 1, 0)))

dag_ucb_u <- dagify(
  A ~ D + G + U,
  D ~ G + U,
  exposure = "A",
  outcome = "G",
  coords = tibble(name = c("G", "D", "A", "U"),
                  x = c(0, .5, 1, 1),
                  y = c(0, 1, 0, 1)))

p1 <- dag_ucb %>%
  fortify() %>% 
  mutate(stage = if_else(name == "A", "response",
                         if_else(name %in% c("G", "D"),
                                 "predictor", "confounds"))) %>% 
  plot_dag(clr_in = clr_current) 

p2 <- dag_ucb_u %>%
  fortify() %>% 
  mutate(stage = if_else(name == "A", "response",
                         if_else(name %in% c("G", "D"),
                                 "predictor", "confounds"))) %>% 
  plot_dag(clr_in = clr_current) 

p1 + p2 &
  coord_fixed(ratio = .5) &
  scale_y_continuous(limits = c(-.1, 1.1)) &
  scale_x_continuous(limits = c(-.1, 1.1)) &
  theme(plot.title = element_text(hjust = .5, family = fnt_sel),
        plot.tag = element_text(family = fnt_sel))
```


```{r, fig.asp = .75, fig.width = 10}
as_draws_df(model_ucb_dept@stanfit) %>% 
  as_tibble() %>% 
  dplyr::select(starts_with("alpha"), 
                starts_with("delta")) %>% 
  ggpairs(
     lower = list(continuous = wrap(my_lower)),
        diag = list(continuous = wrap(my_diag, fill = fll0,
                                      color = clr0d, adjust = .7)),
        upper = list(continuous = wrap(my_upper ,
          size = 5, color = "black", family = fnt_sel)) ) +
    theme(panel.border = element_rect(color = clr_dark, fill = "transparent"))
```

## Poisson Regression

```{r}
y <- rbinom( 1e5, 1000, 1/1000)
c(mean(y), var(y))
```

Models build upon the poisson distribution (for uncountable, or uncoutably large number of trials $N$), are even simpler than binomial or Gaussian model, because there is only a single parameter that describes its shape:

$$
y_{i} \sim
$$


The conventional link function for a Poisson Model is the log link:

$$
\begin{array}{rclr}
y_{i} & \sim & Poisson(\lambda_{i}) & \textrm{[likelihood]}\\
\textrm{log}(\lambda_{i}) & = & \alpha + \beta(x_{i} - \bar{x}) & \textrm{[linear model]}
\end{array}
$$

### Oceanic Tool Complexity

```{r}
library(sf)
data(Kline)

data_kline <- Kline %>% 
  as_tibble() %>% 
  mutate(pop_log_scl = scale(log(population))[,1],
         contact_idx = 3L - as.integer(factor(contact)))

data_kline_sf <- read_sf("data/pacific.gpkg") %>% 
  left_join(Kline) %>%
  mutate(coords = st_centroid(geom) %>% st_coordinates() %>% as_tibble() ) %>%
  mutate(x = coords$X,
         y = coords$Y) %>% 
  dplyr::select(-coords)
  
```


```{r}
library(rnaturalearth)
crs_pac <- 3832
long_lim <- c(100,-140)
lat_lim <- c(-25, 25)
bb <- tibble(x = long_lim[c(1,1,2,2,1)],
       y = lat_lim[c(1,2,2,1,1)]) %>% 
  st_as_sf(coords = c("x","y"), crs = 4326) %>% 
  st_transform(crs = crs_pac) %>% 
  st_combine() %>% 
  st_cast("POLYGON") 

dateline <- tibble(x = c(180,180),
       y = lat_lim) %>% 
  st_as_sf(coords = c("x","y"), crs = 4326) %>% 
  st_transform(crs = crs_pac) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

equator <- tibble(x = long_lim,
       y = c(0, 0)) %>% 
  st_as_sf(coords = c("x","y"), crs = 4326) %>% 
  st_transform(crs = crs_pac) %>% 
  st_combine() %>% 
  st_cast("LINESTRING")

coast <- read_sf("~/work/geo_store/natural_earth/ne_110m_land.shp") %>%
  st_transform(crs = crs_pac) %>%
  st_intersection(bb)

library(raster)
library(ggspatial)
# library(marmap)
# bat_l <- marmap::getNOAA.bathy(long_lim[1], 180,
#                                lat_lim[1], lat_lim[2], resolution = 5)
# 
# bat_r <- marmap::getNOAA.bathy(-180, long_lim[2],
#                                lat_lim[1], lat_lim[2], resolution = 5)
# 
# bat_l_proj <- projectRaster(marmap::as.raster(bat_l), crs = crs(bb))
# bat_r_proj <- projectRaster(marmap::as.raster(bat_r), crs = crs(bb))
# bat <- mosaic(bat_l_proj, bat_r_proj,tolerance = 0.1, fun = mean)

bat <- raster("data/pacific.tif") %>%
  raster::crop(y = as_Spatial(bb)) %>%
  raster::mask(mask = as_Spatial(bb))

ggplot() +
  geom_raster(data = df_spatial(bat),
              aes(x = x, y = y, fill = band1)) +
  geom_sf(data = bb, fill = "transparent", color = clr0d) +
  geom_sf(data = dateline, size = .3, linetype = 3, color = clr_dark) +
  geom_sf(data = equator, size = .3, linetype = 3, color = clr_dark) +
  geom_sf(data = coast, color = clr0dd, fill = clr0, size = .2) +
  geom_sf(data = data_kline_sf,
          color = clr_current,
          fill = clr_lighten(clr_current),
          size = .4) +
  ggrepel::geom_text_repel(data = data_kline_sf,
                           aes(x = x,
                               y = y,
                               label = culture),
                           family = fnt_sel) +
  annotation_north_arrow(style = north_arrow_fancy_orienteering(text_family = fnt_sel)) +
  annotation_scale(style = "ticks", text_family = fnt_sel) +
  scale_fill_gradientn(colours = c("white", "black") %>% 
                         clr_lighten(.35),
                       guide = "none") +
  coord_sf(crs = crs_pac, expand = 0) +
  theme(panel.grid = element_blank(),
        axis.title = element_blank())
```

```{r}
data_kline
```

modeling the idea that:

- the number of tools increases with the log `population` size
- the number of tools increases with the `contact` rate among islands
- the impact of `population` on tool counts is moderated by high `contact` (looking for a positive interaction)

$$
\begin{array}{rclr}
T_{i} & \sim & Poisson(\lambda_{i}) & \textrm{[likelihood]}\\
\textrm{log}(\lambda_{i}) & = & \alpha_{CID[i]} + \beta_{CID[i]}~\textrm{log}~P_{i}& \textrm{[linear model]}\\
\alpha_{j}& \sim & ... & \textrm{[$\alpha$ prior]}\\
\beta_{j}& \sim & ... & \textrm{[$\beta$ prior]}\\
\end{array}
$$

Prior considerations (first for $\alpha$):

$$
\begin{array}{rclr}
T_{i} & \sim & Poisson(\lambda_{i}) & \textrm{[likelihood]}\\
\textrm{log}(\lambda_{i}) & = & \alpha & \textrm{[linear model]}\\
\alpha& \sim & Normal( 0, 10 ) & \textrm{[$\alpha$ prior]}\\
\end{array}
$$

```{r}
ggplot() +
    stat_function(fun = function(x){dlnorm(x, 3, .5)},
                xlim = c(0, 100), n = 201,
                geom = "area", color = clr_current, fill = fll_current())+
  stat_function(fun = function(x){dlnorm(x, 0, 10)},
                xlim = c(0, 100), n = 201,
                geom = "area", color = clr0d, fill = fll0) +
  geom_richtext(data = tibble(x = c(12.5, 37.5),
                              y = c(.065, .035),
                              lab = c(glue("*{mth('\U03B1')}* {mth('\U007E')} dnorm(0, 10)"),
                                      glue("*{mth('\U03B1')}* {mth('\U007E')} dnorm(3, 0.5)")),
                              prior = c("wide", "narrow")),
                aes(x = x, y = y, color = prior, label = lab ),
                hjust = -.05, size = 5,
                fill = NA, label.color = NA, 
                label.padding = grid::unit(rep(0, 4), "pt"),
                family = fnt_sel) +
  scale_color_manual(values = c(narrow = clr_current, wide = clr0dd),
                     guide = "none") +
  labs(x = "mean number of tools", y = "density")
```

Mean of the <span style="color:`r clr0dd`">narrow curve</span> ($exp(\mu + \sigma^2 / 2)$):

```{r}
alpha <- rnorm(1e4, 0, 10)
lambda <- exp(alpha)
mean(lambda)
```

```{r}
tibble(prior = c( "wide", "narrow"),
       mean = c(exp(0 + 10 ^2 /2),
                exp(3 + .5 ^2 /2)))
```


```{r}
mean_weight <- function(mu, sigma){exp(mu + sigma ^2 /2)}

(prior_tab <- crossing(mu = c(0, 1.5, 3, 4),
         sigma = c(5, 1, .5, .1)) %>% 
  mutate(RN = row_number(),
         mean = map2_dbl(mu, sigma, mean_weight)) )

ggplot() +
  (pmap(prior_tab, .f = function(mu, sigma,...){
    force(mu)
    force(sigma)
    return(geom_function(data = tibble(#x = 0:150,
                                       mu = mu),
                  fun = function(x, mu, sigma,...){
                  dlnorm(x, mu, sigma)},
                  aes(color = factor(mu)), 
                  xlim = c(0, 150),
                  n = 201,
                args = list(mu = mu, sigma = sigma)))
  }))  +
  facet_wrap(mu ~ ., scales = "free_y", labeller = label_both) +
  scale_color_manual(values = c(`0` = clr0d, `1.5` = clr1, `3` = clr2, `4` = clr3),
                     guide = "none") +
  labs(x = "mean number of tools", y = "density")
```

Now for the $\beta$ prior:

```{r}
n <- 100
label_strp <- function(prior){glue("beta {mth('\U007E')} Normal({str_remove(prior,'b')})")}

tibble(idx = 1:n,
       a = rnorm(n, mean = 3, sd = 0.5)) %>% 
  mutate(`b0, 10`  = rnorm(n, mean = 0 , sd = 10),
         `b0, 0.2` = rnorm(n, mean = 0 , sd = 0.2)) %>% 
  pivot_longer(contains("b"),
               values_to = "b",
               names_to = "prior") %>% 
  mutate(prior = label_strp(prior),
         prior = factor(prior, levels = label_strp(c("b0, 10", "b0, 0.2"))) ) %>% 
  expand(nesting(idx, a, b, prior),
         x = seq(from = -2, to = 2, length.out = 100)) %>% 
  ggplot(aes(x = x, y = exp(a + b * x), group = idx)) +
  geom_line(size = .2, color = clr_alpha(clr_current, .25)) +
  labs(x = "log population (std)",
       y = "total tools") +
  coord_cartesian(ylim = c(0, 100)) +
  facet_wrap(~ prior) +
  theme(strip.text = element_markdown())
```


```{r}
prior <- tibble(idx = 1:n,
                a = rnorm(n, mean = 3, sd = 0.5),
                b = rnorm(n, mean = 0, sd = 0.2)) %>% 
  expand(nesting(idx, a, b),
         x = seq(from = log(100), to = log(200000),
                 length.out = 100))

p1 <- prior %>% 
  ggplot(aes(x = x, y = exp(a + b * x), group = idx)) +
  geom_line(size = .2, color = clr_alpha(clr_current, .25)) +
  geom_hline(yintercept = 100, linetype = 3, size = .6, 
             color = clr_dark) +
  labs(subtitle = glue("alpha {mth('\U007E')} Normal(3, 0.5); beta {mth('\U007E')} Normal(0, 0.2)"),
       x = "log population",
       y = "total tools") +
  coord_cartesian(xlim = c(log(100), log(200000)),
                  ylim = c(0, 500))

p2 <- prior %>% 
  ggplot(aes(x = exp(x), y = exp(a + b * x), group = idx)) +
  geom_line(size = .2, color = clr_alpha(clr_current, .4)) +
  geom_hline(yintercept = 100, linetype = 3, size = .6, 
             color = clr_dark) +
  labs(subtitle = glue("alpha {mth('\U007E')} Normal(3, 0.5); beta {mth('\U007E')} Normal(0, 0.2)"),
       x = "population",
       y = "total tools") +
  coord_cartesian(xlim = c(100, 200000),
                  ylim = c(0, 500))

p1 + p2 & theme(plot.subtitle = element_markdown(size = 8, hjust = .5))
```


Using a log link function, induces diminishing returns due to establishing a log-linear relationship.

Finally, we can create the model(s) - one with *interaction* and one simpler one that is *intercept-only*.

```{r, results = "hide"}
data_kline_list <- data_kline %>% 
  dplyr::select(total_tools,
                pop_log_scl,
                contact_idx) %>% 
  as.list()

#> intercept only
model_ocean_intercept <- ulam(
  flist = alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- alpha,
    alpha ~ dnorm(3, 0.5)
  ),
  data = data_kline_list,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)

#> interaction model
model_ocean_interact <- ulam(
  flist = alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- alpha[contact_idx] + beta[contact_idx] * pop_log_scl,
    alpha[contact_idx] ~ dnorm(3, 0.5),
    beta[contact_idx] ~ dnorm(0, 0.2)
  ),
  data = data_kline_list,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```

```{r}
compare(model_ocean_intercept,
        model_ocean_interact,
        func = PSIS) %>% 
  knit_precis(param_name = "model")
```

```r
#> Some Pareto k values are high (>0.5). Set pointwise=TRUE to inspect individual points.
#> Some Pareto k values are high (>0.5). Set pointwise=TRUE to inspect individual points.
```

The *effective number of parameters* (`pPSIS` / `pWAIC`) is actually larger for the model with fewer parameters (`model_ocean_intercept`) `r emo::ji("astonished")`.

$\rightarrow$ the clear relationship of overfitting and the number of parameters only holds for simple linear regressions with flat priors!

Checking the influential points (with high `Pareto k values`) through **posterior prediction plots**: 

```{r}
ocean_k_values <- PSIS(model_ocean_interact, pointwise = TRUE) %>% 
  bind_cols(data_kline, .)

n <- 101
ocean_grid <- crossing(pop_log_scl = seq(-1.4, 3, length.out = n),
                        contact_idx = 1:2)

ocean_posterior_predictions <- link(model_ocean_interact,
                                             data = ocean_grid) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(ocean_grid, .) %>% 
  pivot_longer(-c(pop_log_scl, contact_idx), values_to = "total_tools") %>% 
  dplyr::select(-name) %>% 
  group_by(pop_log_scl, contact_idx) %>% 
  summarise(p = list(quantile(total_tools, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p)  %>%
  mutate(population = exp((pop_log_scl *
                             sd(log(data_kline$population))) +
                            mean(log(data_kline$population))))

p1 <- ocean_posterior_predictions %>% 
  ggplot(aes(x = pop_log_scl, y = total_tools)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh,
                  group = factor(contact_idx),
                  color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color))),
              stat = "identity", size = .3) +
  geom_point(data = ocean_k_values,
             aes(color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color)),
                 size = k),
             shape = 21) +
  ggrepel::geom_text_repel(data = ocean_k_values %>% filter(k > .5),
             aes(label = str_c(culture, " (", round(k, digits = 2), ")")),
             family = fnt_sel,nudge_y = 6, min.segment.length = 15) +
  scale_color_manual(values = c(`1` = clr_current, `2` = clr0dd), guide = "none") +
  scale_size_continuous(guide = "none")+
  coord_cartesian(ylim = c(0, 90), x = c(-1.4, 3), expand = 0) +
  labs(subtitle = "on log scale")

p2 <- ocean_posterior_predictions %>% 
  ggplot(aes(x = population, y = total_tools)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh,
                  group = factor(contact_idx),
                  color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color))),
              stat = "identity", size = .3) +
  geom_point(data = ocean_k_values,
             aes(color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color)),
                 size = k),
             shape = 21) +
  scale_color_manual(values = c(`1` = clr_current, `2` = clr0dd), guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(breaks = 1e5 * (0:3), labels = scales::comma) +
  coord_cartesian(ylim = c(0, 90),
                  x = c(-1e3, 3e5),
                  expand = 1) +
  labs(subtitle = "on natural scale")

p1 + p2 &
  theme(plot.subtitle = element_text(hjust = .5))
```

Notice how the posterior predictions expect fewer tools for <span style="color:`r clr0dd`">well connected cultures</span> compared to  <span style="color:`r clr_current`">less connected</span> ones at high population sizes (after the mean lines cross).

This asks for a better model that incorporates scientific knowledge (eg. forcing the model to go through the origin for bot culture types).

### Modeling Tool Innovation

$$
\Delta T = \alpha P^{\beta} - \gamma T
$$

Setting $\Delta T$ to 0 at the equilibrium:

$$
\hat{T} = \frac{\alpha P^{\beta}}{\gamma}
$$

Plugging this into a Poisson model:

$$
\begin{array}{rclr}
T_{i} & \sim & Poisson(\lambda_{i}) & \textrm{[likelihood]}\\
\textrm{log}(\lambda_{i}) & = & \alpha P_{i}^{\beta} / \gamma & \textrm{[linear model]}\\
\end{array}
$$

```{r, results = "hide"}
data_kline_list_sc <- data_kline %>% 
  dplyr::select(total_tools,
                population,
                contact_idx) %>% 
  as.list()

model_ocean_scientific <- ulam(
  flist = alist(
    total_tools ~ dpois(lambda),
    lambda <- exp(alpha[contact_idx]) * population^beta[contact_idx] / gamma,
    alpha[contact_idx] ~ dnorm(1, 1),
    beta[contact_idx] ~ dexp(1),
    gamma ~ dexp(1)
  ),
  data = data_kline_list_sc,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```


```{r}
ocean_k_values_sc <- PSIS(model_ocean_scientific, pointwise = TRUE) %>% 
  bind_cols(data_kline, .)

n <- 101
ocean_grid_sc <- crossing(population = seq(0, 3e5, length.out = n),
                        contact_idx = 1:2)

ocean_posterior_predictions_sc <- link(model_ocean_scientific,
                                             data = ocean_grid_sc) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(ocean_grid_sc, .) %>% 
  pivot_longer(-c(population, contact_idx), values_to = "total_tools") %>% 
  dplyr::select(-name) %>% 
  group_by(population, contact_idx) %>% 
  summarise(p = list(quantile(total_tools, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p)

ocean_posterior_predictions_sc %>% 
  ggplot(aes(x = population, y = total_tools)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh,
                  group = factor(contact_idx),
                  color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color))),
              stat = "identity", size = .3) +
  geom_point(data = ocean_k_values_sc,
             aes(color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color)),
                 size = k),
             shape = 21) +
  scale_color_manual(values = c(`1` = clr_current, `2` = clr0dd), guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(breaks = 1e5 * (0:3), labels = scales::comma) +
  coord_cartesian(ylim = c(0, 80),
                  x = c(-1e3, 3e5),
                  expand = 1) +
  labs(subtitle = "the scientific model") +
  theme(plot.subtitle = element_text(hjust = .5))
```

### Negative Binomial (gamma-Poisson) models

This relates to the Poisson distribution, like the Student-t distribution reates to the normal.
The *negative binomial* (aka. *gamma-Poisson*) distribution is just a mixtrue of several different Poisson distributions

### Exposure and the Offset

In Poisson models $\lambda$ is both the expected value, as well as a *rate*: $\lambda = \mu / \tau$.

$$
\begin{array}{rcl}
y_{i} & \sim & Poisson(\lambda_{i}) \\
\textrm{log}(\lambda_{i}) & = & \textrm{log} \frac{\mu_{i}}{\tau_{i}}~=~\alpha + \beta x_{i} \\
\textrm{log}(\lambda_{i}) & = & \textrm{log}~\mu_{i} - \textrm{log}~\tau_{i}~=~\alpha + \beta x_{i} \\
\end{array}
$$

In here $\tau$ are the *exposures* - if different observations $i$ do have different exposures:

$$
\textrm{log}~\mu_{i} = \textrm{log}~\tau_{i} + \alpha + \beta
$$

We can use $\tau$ (needs to be a data column) to scale the expected number of events:

$$
\begin{array}{rcl}
y_{i} & \sim & Poisson(\mu_{i}) \\
\textrm{log}(\mu_{i}) & = & \textrm{log}~\tau_{i} + \alpha + \beta x_{i} \\
\end{array}
$$

The practical gain here is that we can introduce an *offset* (scaling of the rate), if our sampling is uneven over time (eg. varying lengths of observations, area of sampling, intensity of sampling).

To test this we are modelling book-completions by different medieval monasteries (starting with $\lambda_{M1} = 1.5$ books/day and $\lambda_{M2} = .5$ books/day - but tallied in weeks).

```{r}
n_days <- 30
n_weeks <- 4

books_m1 <- rpois(n_days, 1.5)
books_m2 <- rpois(n_weeks, .5 *7 )

data_books <- tibble(books = c(books_m1, books_m2),
       monastery = rep(0:1, c(n_days,n_weeks)),
       days = rep(c(1,7), c(n_days,n_weeks))) %>%   # exposure
  mutate(days_log = log(days))

model_books <- quap(
  flist = alist(
    books ~ dpois( lambda ),
    log(lambda) <- days_log + alpha + beta * monastery,
    alpha ~ dnorm( 0, 1 ),
    beta ~ dnorm(0 ,1 )
  ),
  data = data_books
)
```

Computing the posterior distributions:

```{r}
books_posterior <- extract.samples(model_books) %>% 
  as_tibble() %>% 
  mutate(lambda_old = exp(alpha),
         lambda_new = exp(alpha + beta))

precis(books_posterior %>% dplyr::select(starts_with("lambda"))) %>% 
  knit_precis(param_name = "column")
```

## Multinomial and Categorical Models

The *multinomial distribution* is the maximum entropy distribution if more than two types of events can happen and if the probabilities for each event is constant.

For $K$ types of events, with the probabilities $p_{1}, ..., p_{K}$, the probability of observing $y_{1}, ..., y_{K}$ in $n$ trials is:

$$
Pr(y_{1}, ..., y_{K}| n, p_{1}, ..., p_{K}) = \frac{n!}{\prod_{i} y_{i}!} \prod_{i=1}^{K} p_{i}^{y_{i}}
$$

The typical link function for multinomial models (categorical models) is the *multinomial logit* (aka. *softmax*) which takes a score vector of the $K$ event types:

$$
Pr(k | s_{1},...,s_{K}) =  \frac{\textrm{exp}(s_{K})}{\sum_{i=1}^{K}\textrm{exp}(s_{i})}
$$

This link function produces a *multinomial logistic regression*.

### Predictor Matches to Outcomes

Simulating career choices based on expected income:

```{r}
n <- 500

career_p <- tibble(income = c(1, 2, 5),
       score = .5 * income,
       p = softmax(score))

set.seed(34302)
career <- 1:n %>% 
  map_dbl(.f = function(x){
    sample(x = 1:3, size = 1, prob = career_p$p)
    })
```

Writing the Stan model by hand:

```{r}
code_model_career <- "
data{
    int N;              // number of individuals
    int K;              // number of possible careers
    int career[N];      // outcome
    vector[K] career_income;
}
parameters{
    vector[K-1] alpha;  // intercepts
    real<lower=0> beta; // association of income with slope
}
model{
    vector[K] p;
    vector[K] s;
    alpha ~ normal( 0, 1 );
    beta ~ normal( 0, 0.5 );
    s[1] = alpha[1] + beta * career_income[1];
    s[2] = alpha[2] + beta * career_income[2];
    s[3] = 0;           // pivot
    p = softmax( s );
    career ~ categorical( p );
}
"
```

```{r, results = "hide"}
data_career_list <- list(
  N = n,
  K = 3,
  career = career,
  career_income = career_p$income
)

model_career <- stan(
  model_code = code_model_career,
  data = data_career_list,
  chains = 4,
  cores = 4
)
```

```{r}
precis(model_career, depth = 2) %>% 
  knit_precis()
```

Counterfactual simulations (doubling income #2):

```{r}
career_counterfactual <- extract.samples(model_career) %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  mutate(idx = row_number()) %>% 
  pivot_longer(cols = starts_with("alpha"), names_to = "case", values_to = "alpha") %>% 
  mutate(case = str_remove(case, "alpha.") %>% as.integer(),
         score = alpha + beta * career_p$income[case],
         double_score = alpha + beta * career_p$income[case] * 2 ) %>% 
  pivot_wider(names_from = case, values_from = alpha:double_score,id_cols = c(idx, beta)) %>% 
  mutate(s2_orig = map2_dbl(score_1, score_2, .f = function(x,y){softmax(c(x,y,0))[2]}),
         s2_new = map2_dbl(score_1, double_score_2, .f = function(x,y){softmax(c(x,y,0))[2]}),
         p_two_dif = s2_new - s2_orig)
  
career_counterfactual %>% 
  dplyr::select(p_two_dif) %>% 
  precis() %>% 
  knit_precis(param_name = "column")
```

:::beware
This section does get quite different values compared to the book - they are similar to the [summary by Solomon Kurz](https://bookdown.org/content/4857/god-spiked-the-integers.html#predictors-matched-to-outcomes) though (and also raised [an issue](https://github.com/rmcelreath/rethinking/issues/296) within the rethinking repo)
:::

### Predictors Matched to Observations

```{r}
n <- 500
beta <- c(-2, 0, 2)

career_fam <- tibble(family_income = runif(n)) %>% 
  mutate(score = purrr::map(family_income, function(x){.5 * 1:3 + beta * x}),
         p = purrr::map(score, function(x){softmax(x[1], x[2], x[3])}),
         career = map_int(p, function(x){sample(1:3, size = 1,prob = x)}))
```

```{r}
code_model_career_fam <- "
data{
    int N;                 // number of observations
    int K;                 // number of outcome values
    int career[N];         // outcome
    real family_income[N]; //
}
parameters{
    vector[K-1] alpha;     // intercepts
    vector[K-1] beta;      // coefficients on family income
}
model{
    vector[K] p;
    vector[K] s;
    alpha ~ normal( 0, 1.5 );
    beta ~ normal( 0, 1 );
    for( i in 1:N ) {
        for( j in 1:(K-1) ){
            s[j] = alpha[j] + beta[j] * family_income[i];
        }
        s[K] = 0;
        p = softmax( s );
        career[i] ~ categorical( p );
    }
}
"
```

```{r, results = "hide"}
data_career_fam_list <- list(
  N = n,
  K = 3,
  career = career_fam$career,
  family_income = career_fam$family_income
)

model_career_fam <- stan( 
  model_code = code_model_career_fam,
  data = data_career_fam_list,
  chains = 4,
  cores = 4
  )
```

```{r}
precis(model_career_fam, depth = 2) %>% 
  knit_precis()
```

### Multinomial in Disguise as Poisson

```{r, results = "hide"}
model_ucb_binom <- quap(
  flist = alist(
    admit ~ dbinom( applications, p),
    logit(p) <- alpha,
    alpha ~ dnorm( 0, 1.5)
  ),
  data = data_ucb
)

data_ucb_list_poisson <- list(
  adm = data_ucb$admit,
  rej = data_ucb$reject
)

model_ucb_poisson <- ulam(
  flist = alist(
    adm ~ dpois( lambda_1 ),
    rej ~ dpois( lambda_2 ),
    log( lambda_1 ) <- alpha_1,
    log( lambda_2 ) <- alpha_2,
    c( alpha_1, alpha_2 ) ~ dnorm( 0, 1.5 )
  ),
  data = data_ucb_list_poisson,
  cores = 4,
  chains = 4
)
```

```{r}
inv_logit(coef(model_ucb_binom))
```


The implied probability of admission within the Poisson model:

$$
p_{ADMIT} = \frac{\lambda_{1}}{\lambda_{1} + \lambda_{2}} = \frac{\textrm{exp}(\alpha_{1})}{\textrm{exp}(\alpha_{1}) + \textrm{exp}(\alpha_{2})}
$$

```{r}
k <- coef(model_ucb_poisson)
alpha_1 <- k["alpha_1"]
alpha_2 <- k["alpha_2"]

exp(alpha_1) / (exp(alpha_1) + exp(alpha_2)) 
```


---

```{r env_export}
library(rlang)
chapter11_models <- env(
  data_chimp = data_chimp,
  model_omega10 = model_omega10,
  model_omega15 = model_omega15,
  model_beta10 = model_beta10,
  model_beta05 = model_beta05,
  model_chimp = model_chimp,
  data_chimp_list = data_chimp_list,
  model_chimp_interaction = model_chimp_interaction,
  data_chimp_list2 = data_chimp_list2,
  model_chimp_stan = model_chimp_stan,
  data_chimp_aggregated = data_chimp_aggregated,
  model_chimp_aggregated = model_chimp_aggregated,
  data_ucb = data_ucb,
  model_ucb = model_ucb,
  data_ucb_list2 = data_ucb_list2,
  model_ucb_dept = model_ucb_dept,
  data_kline = data_kline,
  data_kline_sf = data_kline_sf,
  data_kline_list = data_kline_list,
  model_ocean_intercept = model_ocean_intercept,
  model_ocean_interact = model_ocean_interact,
  data_kline_list_sc = data_kline_list_sc,
  model_ocean_scientific = model_ocean_scientific,
  data_books = data_books,
  model_books = model_books,
  data_career_list = data_career_list,
  model_career = model_career,
  data_career_fam_list = data_career_fam_list,
  model_career_fam = model_career_fam,
  model_ucb_binom = model_ucb_binom,
  data_ucb_list_poisson = data_ucb_list_poisson,
  model_ucb_poisson = model_ucb_poisson
)

write_rds(chapter11_models, "envs/chapter11_models.rds")
```


## Homework

**E1**

Based on the definition of log-odds:

$$
\textrm{log-odds} = \textrm{log}(\frac{p}{1-p})
$$

```{r E1}
log(.35 / (1 - .35))
```

**E2**

Rearranging the log-odds formula to represent the *logistic* function (s. page 317):

$$
p_{i} = \frac{e^\textrm{log-odds}}{1 - e^\textrm{log-odds}}
$$

```{r E2}
logistic(3.2)

exp(3.2) / (1 + exp(3.2))

x <- exp(3.2) / (1 + exp(3.2))
log(x / (1 - x))
```

**E3**

A coefficient in a logistic regression with a value of 1.7 implies rate 70% increase for every unit change.

**E4**

The offset is needed if rates are compared that are measured in different intervals.
It provides a way of standardizing - in our example this was two monasteries tallying their manuscript productions over different time intervals (days vs. weeks).

**M1**

This is likely an effect of the multiplicity that is inflated when using the dis-aggregated format (because the ordering for individual trials is flexible, while the aggregated form is just interpreted as a singular event).
In a way, this also has to happen if the posteriors are the same, but the priors change (because we are not assuming a Bernoulli distribution as prior for the aggregated data). Therefore the likelihood has to balance this difference if we are still to arrive at the same posterior distribution.

**M2**

An increase in 1.7 in a Poisson regression translates to an increase in 1.7 orders of magnitude (I think `r emo::ji("thinking")` - since a log link function is commonly used here).

**M3**

The inverse-logit transforms all values from the $\mathbb{R}$ space (real numbers) to the range of probabilities, which is $p \in [0, 1]$.
It is thus suited to transform variables that are far from any boundary to the probability scale.

```{r M3}
p1 <- ggplot() +
  stat_function(fun = logit,
                xlim = c(0, 1), 
                geom = "line",
                color = clr2, fill = fll2) +
  geom_vline(xintercept = c(0, 1), linetype = 3,
             color = clr_dark) +
  labs(subtitle = "the logit function",
       x = "p", y = "natural scale") 

p2 <- ggplot() +
  stat_function(fun = inv_logit,
                xlim = c(-10, 10), 
                geom = "area",
                color = clr2, fill = fll2) +
  geom_hline(yintercept = c(0, 1), linetype = 3,
             color = clr_dark) +
  labs(subtitle = "the inverse logit",
       y = "p", x = "natural scale")

p1 + p2 &
  theme(plot.subtitle = element_text(hjust = .5))
```

**M4**

The inverse-logit transforms all values from the $\mathbb{R}$ space (real numbers) to the space  $\mathbb{R}_{*}^{+}$ (positive real numbers).
It is thus suited to transform variables that are far from any boundary to the probability scale.

```{r M4}
p1 <- ggplot() +
  stat_function(fun = log,
                xlim = c(0, 100), 
                geom = "line",n = 501,
                color = clr2, fill = fll2) +
  geom_vline(xintercept = 0, linetype = 3,
             color = clr_dark) +
  labs(subtitle = "the log function",
       x = "count", y = "parameter scale") +
  coord_cartesian(ylim = c(-1.25, 4.5))

p2 <- ggplot() +
  stat_function(fun = exp,
                xlim = c(-1.25, 4.6), 
                geom = "area",
                n = 501,
                color = clr2, fill = fll2) +
  geom_hline(yintercept = 0, linetype = 3,
             color = clr_dark) +
  labs(subtitle = "the exponential function",
       y = "count", x = "parameter scale")

p1 + p2 &
  theme(plot.subtitle = element_text(hjust = .5))
```

**M5**

Using the logit link for a Poisson GLM would imply that the rate $\lambda$ also does have a theoretical upper bound that needs to be considered.

**M6**

Constraints of the binomial distribution

- discrete (count distribution)
- $n$ trials
- constant expected value $np$

Constraints of the Poisson distribution (special case of the Binomial)

- discrete (count distribution)
- $n$ is very large (even unknown)
- $p$ is very small
- expected rate per unit time $\lambda = np$ 

The constraints are similar since the Poisson is a special case of the Binomial distribution.
There could be an argument that knowing the magnitude of $n$ and $p$ implies further previous knowledge about the system.

**M7**

```{r M7, results = "hide"}
model_chimp_quap <- quap(
   flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha[actor] + beta[treatment],
    alpha[actor] ~ dnorm( 0, 1.5 ),
    beta[treatment] ~ dnorm(0,0.5)
  ),
  data = data_chimp_list
)

model_chimp_a10 <- ulam(
  flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha[actor] + beta[treatment],
    alpha[actor] ~ dnorm( 0, 10 ),
    beta[treatment] ~ dnorm(0,0.5)
  ),
  data = data_chimp_list,
  chains = 4, cores = 4,
  log_lik = TRUE
)

model_chimp_a10_quap <- quap(
   flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha[actor] + beta[treatment],
    alpha[actor] ~ dnorm( 0, 10 ),
    beta[treatment] ~ dnorm(0,0.5)
  ),
  data = data_chimp_list
)
```


```{r}
plot_chimps <- function(model, plot_title){
  is_ulam <- grepl("ulam", plot_title)
  clr_current <- c(clr1, clr2)[1 + is_ulam]
  
  chimp_posterior <- extract.samples(model) %>% 
    as.data.frame() %>% 
    as_tibble() %>% 
    mutate(across(starts_with("alpha"),.fns = inv_logit, .names = "p_{.col}"))
  
  p1 <- chimp_posterior %>% 
    dplyr::select(starts_with("p")) %>% 
    pivot_longer(everything(),
                 names_to = "actor",
                 values_to = "p") %>% 
    mutate(actor = str_remove(actor,"p_alpha.") %>% as.integer()) %>% 
    group_by(actor) %>% 
    summarise(p = list(quantile(p, probs = c(.055, .25, .5, .75, .955))),
              breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
    ungroup() %>% 
    unnest(c(p, breaks)) %>% 
    pivot_wider(names_from = breaks, values_from = p) %>% 
    ggplot(aes(y = factor(actor, levels = 7:1))) +
    geom_linerange(aes(xmin = ll, xmax = hh), lwd = .2, color = clr_current) +
    geom_linerange(aes(xmin = l, xmax = h), lwd = .7, color = clr_current) +
    geom_point(aes(x = m),
               size = 2, shape = 21,
               color = clr_current, fill = clr_lighten(clr_current, .2)) +
    labs(y = "actor", x = "alpha", subtitle = plot_title) +
    coord_cartesian(xlim = c(0, 1)) +
    theme(plot.subtitle = element_text(hjust = .5))
  
  p2 <- chimp_posterior %>% 
    dplyr::select(starts_with("beta")) %>% 
    mutate(`beta.5` = `beta.1` - `beta.3` ,
           `beta.6` = `beta.2` - `beta.4` ) %>% 
    pivot_longer(everything(),
                 names_to = "param",
                 values_to = "p") %>% 
    mutate(treatment = treatment_labels[str_remove(param, "beta.") %>% as.integer()]) %>% 
    group_by(treatment) %>% 
    summarise(p = list(quantile(p, probs = c(.055, .25, .5, .75, .955))),
              breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
    ungroup() %>% 
    unnest(c(p, breaks)) %>% 
    pivot_wider(names_from = breaks, values_from = p) %>% 
    ggplot(aes(y = factor(treatment, levels = rev(treatment_labels)))) +
    geom_rect(data = tibble(x = 1),inherit.aes = FALSE,
              aes(xmin = -Inf, xmax = Inf, ymin = .5, ymax = 2.5),
              color = clr0d, fill = fll0) +
    geom_vline(xintercept = 0, linetype = 3, color = clr_dark) +
    geom_linerange(aes(xmin = ll, xmax = hh), lwd = .2, color = clr_current) +
    geom_linerange(aes(xmin = l, xmax = h), lwd = .7, color = clr_current) +
    geom_point(aes(x = m),
               size = 2, shape = 21,
               color = clr_current, fill = clr_lighten(clr_current, .2)) +
    scale_y_discrete() +
    labs(y = "treatment", x = "beta")
  
  p1 + p2
}
```

```{r, fig.asp = 1.2}
list(model_chimp,
     model_chimp_quap,
     model_chimp_a10,
     model_chimp_a10_quap) %>% 
  map2(.y = c("ulam (0, 1.5)",
              "quap (0, 1.5)",
              "ulam (0, 10)",
              "quap (0, 10)"),
       plot_chimps) %>% 
  wrap_plots(ncol = 1)
```

**M8**

```{r M8, results = "hide"}
data_kline_non_hawaii <- data_kline %>% 
  filter(culture != "Hawaii")

data_kline_nH_list <- data_kline_non_hawaii %>% 
  dplyr::select(total_tools,
                pop_log_scl,
                contact_idx) %>% 
  as.list()

model_ocean_interact_non_hawaii <- ulam(
  flist = alist(
    total_tools ~ dpois(lambda),
    log(lambda) <- alpha[contact_idx] + beta[contact_idx] * pop_log_scl,
    alpha[contact_idx] ~ dnorm(3, 0.5),
    beta[contact_idx] ~ dnorm(0, 0.2)
  ),
  data = data_kline_nH_list,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)

data_kline_nH_list_sc <- data_kline_non_hawaii %>% 
  dplyr::select(total_tools,
                population,
                contact_idx) %>% 
  as.list()

model_ocean_non_hawaii_scientific <- ulam(
  flist = alist(
    total_tools ~ dpois(lambda),
    lambda <- exp(alpha[contact_idx]) * population ^ beta[contact_idx] / gamma,
    alpha[contact_idx] ~ dnorm(1, 1),
    beta[contact_idx] ~ dexp(1),
    gamma ~ dexp(1)
  ),
  data = data_kline_nH_list_sc,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```

```{r}
ocean_nH_k_values <- PSIS(model_ocean_interact_non_hawaii, pointwise = TRUE) %>% 
  bind_cols(data_kline_non_hawaii, .)

ocean_nH_posterior_predictions <- link(model_ocean_interact_non_hawaii,
                                             data = ocean_grid) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(ocean_grid, .) %>% 
  pivot_longer(-c(pop_log_scl, contact_idx), values_to = "total_tools") %>% 
  dplyr::select(-name) %>% 
  group_by(pop_log_scl, contact_idx) %>% 
  summarise(p = list(quantile(total_tools, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p)  %>%
  mutate(population = exp((pop_log_scl *
                             sd(log(data_kline_non_hawaii$population))) +
                            mean(log(data_kline_non_hawaii$population))))

p1 <- ocean_nH_posterior_predictions %>% 
  ggplot(aes(x = pop_log_scl, y = total_tools)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh,
                  group = factor(contact_idx),
                  color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color))),
              stat = "identity", size = .3) +
  geom_point(data = ocean_nH_k_values,
             aes(color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color)),
                 size = k),
             shape = 21) +
  ggrepel::geom_text_repel(data = ocean_nH_k_values %>% filter(k > .5),
             aes(label = str_c(culture, " (", round(k, digits = 2), ")")),
             family = fnt_sel,nudge_y = 6, min.segment.length = 15) +
  scale_color_manual(values = c(`1` = clr_current, `2` = clr0dd), guide = "none") +
  scale_size_continuous(guide = "none")+
  coord_cartesian(ylim = c(0, 90), x = c(-1.4, 3), expand = 0) +
  labs(subtitle = "on log scale")

p2 <- ocean_nH_posterior_predictions %>% 
  ggplot(aes(x = population, y = total_tools)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh,
                  group = factor(contact_idx),
                  color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color))),
              stat = "identity", size = .3) +
  geom_point(data = ocean_nH_k_values,
             aes(color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color)),
                 size = k),
             shape = 21) +
  scale_color_manual(values = c(`1` = clr_current, `2` = clr0dd), guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(breaks = 1e4 * (0:3), labels = scales::comma) +
  coord_cartesian(ylim = c(0, 90),
                  x = c(-1e3, 3e4),
                  expand = 1) +
  labs( subtitle = "on natural scale" )

p1 + p2 &
  theme(plot.subtitle = element_text(hjust = .5))
```

```{r}
ocean_nH_k_values_sc <- PSIS(model_ocean_non_hawaii_scientific, pointwise = TRUE) %>% 
  bind_cols(data_kline_non_hawaii, .)


n <- 101
ocean_grid_nH_sc <- crossing(population = seq(0, 3e4, length.out = n),
                        contact_idx = 1:2)

ocean_nH_posterior_predictions_sc <- link(model_ocean_non_hawaii_scientific,
                                             data = ocean_grid_nH_sc) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(ocean_grid_nH_sc, .) %>% 
  pivot_longer(-c(population, contact_idx), values_to = "total_tools") %>% 
  dplyr::select(-name) %>% 
  group_by(population, contact_idx) %>% 
  summarise(p = list(quantile(total_tools, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p)

ocean_nH_posterior_predictions_sc %>% 
  ggplot(aes(x = population, y = total_tools)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh,
                  group = factor(contact_idx),
                  color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color))),
              stat = "identity", size = .3) +
  geom_point(data = ocean_nH_k_values_sc,
             aes(color = factor(contact_idx),
                  fill = after_scale(clr_alpha(color)),
                 size = k),
             shape = 21) +
  scale_color_manual(values = c(`1` = clr_current, `2` = clr0dd), guide = "none") +
  scale_size_continuous(guide = "none") +
    scale_x_continuous(breaks = 1e4 * (0:3), labels = scales::comma) +
  coord_cartesian(ylim = c(0, 80),
                  x = c(-1e3, 3e4),
                  expand = 1) +
  labs(subtitle = "the scientific model") +
  theme(plot.subtitle = element_text(hjust = .5))
```

**H1**

```{r H1, results = 'hide'}
model_beta05_ulam <- ulam(
  flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha + beta[treatment],
    alpha ~ dnorm( 0, 1.5 ),
    beta[treatment] ~ dnorm(0,0.5)
  ),
  data = data_chimp_list,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)

model_omega15_ulam <- ulam(
  flist = alist(
    pulled_left ~ dbinom( 1, p ),
    logit(p) <-  alpha,
    alpha ~ dnorm( 0, 1.5 )
  ),
  data = data_chimp_list,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```

```{r}
compare(model_chimp,
        model_chimp_interaction,
        model_beta05_ulam,
        model_omega15_ulam) %>% 
  knit_precis(param_name = "model")
```

The models without the individual intercepts is doing a way worse job according to WAIC compared to the two models that do include individual intercepts.
Including the intercepts is also a major improvement and more important than including a slope for the treatment (which has only a marginal effect).

**H2**

```{r H2}
data(eagles,package = "MASS")

data_eagles <- eagles %>% 
  as_tibble() %>% 
  rename(sucesses = "y",
         n_total = "n",
         pirate_size = "P",
         pirate_age = "A",
         victim_size = "V") %>% 
  mutate(p = 2L - as.integer(pirate_size),
         a = 2L - as.integer(pirate_age),
         v = 2L - as.integer(victim_size))

rm(eagles)
```

**a.**

$$
\begin{array}{rclr}
y_{i} & \sim & Binomial(n_{i}, p_{i}) & \textrm{[likelihood]}\\
\textrm{logit}(p_{i}) & = & \alpha + \beta_{P} P_{i} + \beta_{V} V_{i}  + \beta_{A} A_{i} & \textrm{[linear model]}\\
\alpha &\sim & Normal(0, 1.5) & \textrm{[$\alpha$ prior]}\\
\beta_{P}, \beta_{V}, \beta_{A} & \sim & Normal(0, 0.5) & \textrm{[$\beta$ priors]}\\
\end{array}
$$

```{r, results = "hide"}
model_eagle_quap <- quap(
  flist = alist(
    sucesses ~ dbinom(n_total, p_i),
    logit(p_i) <- alpha + beta_p * p + beta_v *v + beta_a * a,
    alpha ~ dnorm( 0, 1.5 ),
    c( beta_p, beta_v, beta_a ) ~ dnorm( 0, 0.5 )
  ),
  data = data_eagles
)

model_eagle_ulam <- ulam(
  flist = alist(
    sucesses ~ dbinom(n_total, p_i),
    logit(p_i) <- alpha + beta_p * p + beta_v *v + beta_a * a,
    alpha ~ dnorm( 0, 1.5 ),
    c( beta_p, beta_v, beta_a ) ~ dnorm( 0, 0.5 )
  ),
  data = data_eagles,
  chain = 4, 
  cores = 4, 
  log_lik = 4
)
```

```{r}
precis(model_eagle_quap, depth = 2) %>%
  knit_precis()

precis(model_eagle_ulam, depth = 2) %>%
  knit_precis()
```

```{r}
clr_current <- clr2
eagle_posterior_quap <- extract.samples(model_eagle_quap) %>%
  as_tibble() %>% 
  pivot_longer(everything(), names_to = "parameter") %>% 
  mutate(model = "quap")

eagle_posterior_ulam <- extract.samples(model_eagle_ulam) %>%
  as_tibble() %>% 
  pivot_longer(everything(), names_to = "parameter") %>% 
  mutate(model = "ulam")

eagle_quap_params <- precis(model_eagle_quap, depth = 2) %>%
  data.frame() %>%
  rownames_to_column(var = "parameter") %>% 
  mutate(rowid = as.integer(factor(parameter)))

add_quap_theory <- function(rowid, mean, sd, scale = .3, ...){
  stat_function(fun = function(x, mean, sd, rowid){ rowid + dnorm(x, mean, sd) * scale }, 
                geom = "line", args = list(rowid = rowid,
                                           mean = mean,
                                           sd = sd), 
                color = clr0dd, linetype = 3)
}

bind_rows(eagle_posterior_quap,
          eagle_posterior_ulam) %>% 
  ggplot(aes(x = value, y = parameter, color = model)) +
  stat_slab( height = .5,
             size = .5, 
             aes(side = c(quap = "right", ulam = "left")[model],
                 fill = after_scale(clr_alpha(color)))) +
  (pmap(eagle_quap_params, add_quap_theory)) +
  scale_color_manual(values = c(quap = clr0d, ulam = clr_current)) +
  theme(legend.position = "bottom")
```

$\rightarrow$ yes, I would say the quadratic approximation does a very good job here.

**b.**

$$
\textrm{logit}(p_{i}) = \alpha + \beta_{P} P_{i} + \beta_{V} V_{i}  + \beta_{A} A_{i}
$$

```{r}
# eagles_grid <- data_eagles %>% 
#   distinct(p,a,v, n_total = 1e4)

eagles_mod_val <- link(model_eagle_quap) %>% 
  as_tibble() %>%
  summarise(across(everything(),
                   .fns = function(x){quantile(x, probs = c(.055, .5, .945))})) %>% 
  mutate(percentile = c("lower_89", "median", "upper_89")) %>% 
  pivot_longer(-percentile, names_to = "rowid") %>% 
  pivot_wider(names_from = percentile) %>% 
  bind_cols(data_eagles,. ) %>% 
  mutate(rn = row_number())

eagles_sim <- sim(model_eagle_quap) %>% 
  as_tibble() %>%
  summarise(across(everything(),
                   .fns = function(x){quantile(x, probs = c(.055, .5, .945))})) %>% 
  mutate(percentile = c("lower_89", "median", "upper_89")) %>% 
  pivot_longer(-percentile, names_to = "rowid") %>% 
  pivot_wider(names_from = percentile) %>% 
  bind_cols(data_eagles,. ) %>%
  mutate(across(c(lower_89, median, upper_89),
                .fns = function(x, n){x/n},
                n = n_total, .names = c("{.col}_p")),
         rn = row_number()) 

eagles_mod_val %>% 
  ggplot(aes(y = glue("v {victim_size}, p {pirate_size}, a {pirate_age}, n {n_total}"))) +
  geom_pointinterval(aes(xmin = lower_89, x = median, xmax = upper_89))+
  geom_point(data = eagles_sim %>% pivot_longer(ends_with("89_p")), 
             aes(x = value), size = .4) +
  geom_point(aes(x = sucesses / n_total), color = clr_current, shape = 1, size = 2) +
  labs(y = "case") +
  lims(x = 0:1)

eagles_sim %>% 
  ggplot(aes(y = glue("v {victim_size}, p {pirate_size}, a {pirate_age}, n {n_total}"))) +
  geom_pointinterval(aes(xmin = lower_89, x = median, xmax = upper_89))+
  geom_point(aes(x = sucesses), color = clr_current, shape = 1, size = 2) +
  labs(y = "case") 
```

**c.**

```{r}
model_eagle_interact_quap <- quap(
  flist = alist(
    sucesses ~ dbinom(n_total, p_i),
    logit(p_i) <- alpha + beta_p * p + beta_v *v + beta_a * a + beta_ap * a * p,
    alpha ~ dnorm( 0, 1.5 ),
    c( beta_p, beta_v, beta_a, beta_ap ) ~ dnorm( 0, 0.5 )
  ),
  data = data_eagles
)

sim(model_eagle_interact_quap) %>% 
  as_tibble() %>%
  summarise(across(everything(),
                   .fns = function(x){quantile(x, probs = c(.055, .5, .945))})) %>% 
  mutate(percentile = c("lower_89", "median", "upper_89")) %>% 
  pivot_longer(-percentile, names_to = "rowid") %>% 
  pivot_wider(names_from = percentile) %>% 
  bind_cols(data_eagles,. ) %>% 
  ggplot(aes(y = glue("v {victim_size}, p {pirate_size}, a {pirate_age}"))) +
  geom_pointinterval(aes(xmin = lower_89, x = median, xmax = upper_89))+
  geom_point(aes(x = sucesses), color = clr_current, shape = 1, size = 2) +
  labs(y = "case") 

compare(model_eagle_interact_quap,
        model_eagle_quap)
```

**H3**

```{r H3}
data(salamanders)
data_salamander <- salamanders %>% 
  as_tibble() %>% 
  set_names(nm = c("site_idx", "salamander_n", "cover_pct", "forest_age")) %>% 
  mutate(across(cover_pct:forest_age, standardize, .names = "{.col}_std"))

rm(salamanders)
```

**a.**

```{r, results = 'hide'}
model_salamander_quap <- quap(
  flist = alist(
    salamander_n ~ dpois(lambda),
    log(lambda) <- alpha + beta_c * cover_pct_std,
    alpha ~ dnorm(1.5, .5),
    beta_c ~ dnorm(0, 0.2)
  ),
  data = data_salamander
)

model_salamander_ulam <- ulam(
  flist = alist(
    salamander_n ~ dpois(lambda),
    log(lambda) <- alpha + beta_c * cover_pct_std,
    alpha ~ dnorm(1.5, .5),
    beta_c ~ dnorm(0, 0.2)
  ),
  data = data_salamander,
  chains = 4, 
  cores = 4,
  log_lik = TRUE
)
```

```{r}
salamander_posterior_quap <- extract.samples(model_salamander_quap) %>%
  as_tibble() %>% 
  pivot_longer(everything(), names_to = "parameter") %>% 
  mutate(model = "quap")

salamander_posterior_ulam <- extract.samples(model_salamander_ulam) %>%
  as_tibble() %>% 
  pivot_longer(everything(), names_to = "parameter") %>% 
  mutate(model = "ulam")

salamander_quap_params <- precis(model_salamander_quap, depth = 2) %>%
  data.frame() %>%
  rownames_to_column(var = "parameter") %>% 
  mutate(rowid = as.integer(factor(parameter)))

bind_rows(salamander_posterior_quap,
          salamander_posterior_ulam) %>% 
  ggplot(aes(x = value, y = parameter, color = model)) +
  stat_slab( height = .55,
             size = .5, 
             aes(side = c(quap = "right", ulam = "left")[model],
                 fill = after_scale(clr_alpha(color))),
             trim = FALSE, normalize = "xy") +
  (pmap(salamander_quap_params, add_quap_theory, scale = .1)) +
  scale_color_manual(values = c(quap = clr0d, ulam = clr_current)) +
  theme(legend.position = "bottom")
```

```{r}
salamander_mod_val <- link(model_salamander_ulam) %>% 
  as_tibble() %>%
  summarise(across(everything(),
                   .fns = function(x){quantile(x, probs = c(.055, .5, .945))})) %>% 
  mutate(percentile = c("lower_89", "median", "upper_89")) %>% 
  pivot_longer(-percentile, names_to = "rowid") %>% 
  pivot_wider(names_from = percentile) %>% 
  bind_cols(data_salamander,. ) %>% 
  mutate(rn = row_number())

salamander_sim <- sim(model_salamander_ulam) %>% 
  as_tibble() %>%
  summarise(across(everything(),
                   .fns = function(x){quantile(x, probs = c(.055, .5, .945))})) %>% 
  mutate(percentile = c("lower_89", "median", "upper_89")) %>% 
  pivot_longer(-percentile, names_to = "rowid") %>% 
  pivot_wider(names_from = percentile) %>% 
  bind_cols(data_salamander, . ) %>% 
  mutate(rn = row_number())

salamander_mod_val %>% 
  mutate(rn = fct_reorder(as.character(rn), -salamander_n)) %>% 
  ggplot(aes(x = rn, color = log10(forest_age))) +
  geom_pointinterval(aes(ymin = lower_89, y = median, ymax = upper_89)) +
  geom_point(data = salamander_sim %>% pivot_longer(ends_with("89")), 
             aes(y = value), size = .4) +
  geom_point(aes(y = salamander_n), color = clr_dark, shape = 1, size = 2) +
  scale_color_gradientn(colours = c(clr_lighten(clr_current), clr_dark)) +
  labs(x = "case") +
  guides(color = guide_colorbar(title.position = "top",
                                barwidth = unit(.7, "npc"),
                                barheight = unit(4, "pt")))+
  theme(legend.position = "bottom")
```

```{r}
n <- 301
new_salamander <- tibble(cover_pct_std = seq(-1.8, 1.8, length.out = n))

salamander_posterior_predictions <- link(model_salamander_ulam,
                                             data = new_salamander) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(new_salamander, .) %>% 
  pivot_longer(-cover_pct_std, values_to = "salamander_n") %>% 
  dplyr::select(-name) %>% 
  group_by(cover_pct_std) %>% 
  summarise(p = list(quantile(salamander_n, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p)

salamander_posterior_predictions %>% 
  ggplot(aes(x = cover_pct_std, y = salamander_n)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh),
              color = clr2, fill = fll2,
              stat = "identity", size = .3) +
  geom_point(data = data_salamander,
             color = clr2, fill = fll2,
             shape = 21)
```


**b.**

```{r, results = "hide"}
model_salamander_age <- ulam(
  flist = alist(
    salamander_n ~ dpois(lambda),
    log(lambda) <- alpha + beta_c * cover_pct_std + beta_a * forest_age_std,
    alpha ~ dnorm(-.5, .5),
    beta_c ~ dnorm(0, 0.2),
    beta_a ~ dnorm(0, 0.2)
  ),
  data = data_salamander,
  chains = 4, 
  cores = 4,
  log_lik = TRUE
)
```

```{r}
library(ggmcmc)
clr_chains <- function(n = 4, alpha = .7, col = clr2){
  scales::colour_ramp(colors = c(clr0dd, col))(seq(0,1,length.out = n)) %>% 
    clr_lighten(.2) %>% 
    clr_alpha(alpha = alpha)
  }

ggs(model_salamander_age@stanfit, inc_warmup = TRUE) %>%
  mutate(chain = factor(Chain)) %>% 
  filter(Parameter %in% c("alpha", "beta_c", "beta_a")) %>% 
  ggplot(aes(x = Iteration, y = value)) +
  annotate(geom = "rect", 
           xmin = 0, xmax = 500, ymin = -Inf, ymax = Inf,
           fill = clr0d, alpha = .3, size = 0) +
  geom_line(aes(color = chain),
            size = .15) +
  scale_color_manual(values = clr_chains() ) +
  facet_wrap(~ Parameter, scales = "free_y") +
    labs(x = NULL, y = NULL) +
  theme(legend.position = "bottom") 
```

```{r}
compare(model_salamander_ulam,
        model_salamander_age) %>% 
  knit_precis(param_name = "model")
```


```{r}
salamander_age_mod_val <- link(model_salamander_age) %>% 
  as_tibble() %>%
  summarise(across(everything(),
                   .fns = function(x){quantile(x, probs = c(.055, .5, .945))})) %>% 
  mutate(percentile = c("lower_89", "median", "upper_89")) %>% 
  pivot_longer(-percentile, names_to = "rowid") %>% 
  pivot_wider(names_from = percentile) %>% 
  bind_cols(data_salamander,. ) %>% 
  mutate(rn = row_number())

salamander_age_sim <- sim(model_salamander_age) %>% 
  as_tibble() %>%
  summarise(across(everything(),
                   .fns = function(x){quantile(x, probs = c(.055, .5, .945))})) %>% 
  mutate(percentile = c("lower_89", "median", "upper_89")) %>% 
  pivot_longer(-percentile, names_to = "rowid") %>% 
  pivot_wider(names_from = percentile) %>% 
  bind_cols(data_salamander, . ) %>% 
  mutate(rn = row_number())

salamander_age_mod_val %>% 
  mutate(rn = fct_reorder(as.character(rn), -salamander_n)) %>% 
  ggplot(aes(x = rn, color = log10(forest_age))) +
  geom_pointinterval(aes(ymin = lower_89, y = median, ymax = upper_89)) +
  geom_point(data = salamander_age_sim %>% pivot_longer(ends_with("89")), 
             aes(y = value), size = .4) +
  geom_point(aes(y = salamander_n), color = clr_dark, shape = 1, size = 2) +
  scale_color_gradientn(colours = c(clr_lighten(clr_current), clr_dark)) +
  labs(x = "case") +
  guides(color = guide_colorbar(title.position = "top",
                                barwidth = unit(.7, "npc"),
                                barheight = unit(4, "pt")))+
  theme(legend.position = "bottom")
```

```{r}
n <- 301

salamander_grid <- crossing(forest_age_std = seq(-1, 3, length.out = n),
                        cover_pct_std = c(-1, 0, 1)) %>% 
  mutate(cover_class = cut(cover_pct_std, c(-2, -.5 , .5, 2)))

salamander_age_posterior_predictions <- link(model_salamander_age,
                                             data = salamander_grid) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(salamander_grid, .) %>% 
  pivot_longer(-c(forest_age_std:cover_class), values_to = "salamander_n") %>% 
  dplyr::select(-name) %>% 
  group_by(forest_age_std, cover_pct_std, cover_class) %>% 
   summarise(p = list(quantile(salamander_n, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p)

salamander_age_posterior_predictions %>% 
  ggplot(aes(x = forest_age_std, y = salamander_n)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh,
                  color = cover_class, fill = after_scale(clr_alpha(color))),
              stat = "identity", size = .3) +
  geom_point(data = data_salamander %>% 
               mutate(cover_class = cut(cover_pct_std, c(-2, -.5 , .5, 2))) ,
             aes(color = cover_class, fill = after_scale(clr_alpha(color))),
             shape = 21) +
  ggrepel::geom_text_repel(data = data_salamander %>% 
              filter(site_idx %in% 1:5) %>% 
               mutate(cover_class = cut(cover_pct_std, c(-2, -.5 , .5, 2))) ,
             aes(label = site_idx),
             faily = fnt_sel)+
  scale_color_manual(values = c(clr0dd, clr_dark, clr_current), guide = "none") +
  facet_wrap(cover_class ~ . , labeller = label_both)
```


**H4**

```{r H4}
data(NWOGrants)
data_nwo <- NWOGrants %>% 
  as_tibble() %>% 
  mutate(gender_idx = as.integer(gender),
         discipline_idx = as.integer(discipline))

rm(NWOGrants)
clr_current <- clr3
```

```{r}
data_nwo %>% 
  ggplot(aes(x = discipline, color = gender)) +
  geom_bar(position = position_dodge2(padding  = .15),
           width = .9, stat = 'identity',
           aes(y = awards, fill = after_scale(clr_alpha(color,.8)))) +
  geom_bar(position = position_dodge2(padding  = .15),
           width = .9, stat = 'identity',
           aes(y = -applications*.15, fill = after_scale(clr_alpha(color,.4))))+
  geom_hline(yintercept = 0, lietype = 3, color = clr_dark) +
  scale_color_manual(values = c(f = clr_current, m = clr0dd))
```


```{r}
dag_nwo <- dagify(
  A ~ D + G,
  D ~ G,
  exposure = "A",
  outcome = "G",
  coords = tibble(name = c("G", "D", "A"),
                  x = c(0, .5, 1),
                  y = c(0, 1, 0)))

dag_nwo %>%
  fortify() %>% 
  mutate(stage = if_else(name == "A", "response",
                         if_else(name %in% c("G", "D"),
                                 "predictor", "confounds"))) %>% 
  plot_dag(clr_in = clr_current) +
  coord_fixed(ratio = .5) +
  scale_y_continuous(limits = c(-.1, 1.1)) +
  scale_x_continuous(limits = c(-.1, 1.1)) +
  theme(plot.title = element_text(hjust = .5, family = fnt_sel),
        plot.tag = element_text(family = fnt_sel))
```

```{r, results = 'hide'}
model_nwo <- ulam(
  flist = alist(
    awards ~ dbinom( applications, p ),
    logit(p) <- alpha[gender_idx],
    alpha[gender_idx] ~ dnorm( 0, 1.5 ),
    delta[discipline_idx] ~ dnorm( 0, 1.5 )
  ),
  data = data_nwo,
  iter = 4000,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)

model_nwo_dept <- ulam(
  flist = alist(
    awards ~ dbinom( applications, p ),
    logit(p) <- alpha[gender_idx] + delta[discipline_idx],
    alpha[gender_idx] ~ dnorm( 0, 1.5 ),
    delta[discipline_idx] ~ dnorm( 0, 1.5 )
  ),
  data = data_nwo,
  iter = 4000,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```


```{r}
precis(model_nwo, depth = 2) %>% 
  knit_precis()

precis(model_nwo_dept, depth = 2) %>% 
  knit_precis()
```


```{r}
compare(model_nwo,
        model_nwo_dept)
```

```{r}
nwo_grid <- distinct(data_nwo, 
                     discipline_idx, gender_idx, applications)


nwo_posterior_predictions <- sim(model_nwo,
                                  data = nwo_grid) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(nwo_grid, .) %>% 
  pivot_longer(-c(discipline_idx, gender_idx, applications), values_to = "awards") %>% 
  dplyr::select(-name) %>% 
  group_by(discipline_idx, gender_idx, applications) %>% 
  summarise(p = quantile(awards/applications, probs = c(.055, .25, .5, .75, .955)),
            median = median(awards),
            breaks = c("ll", "l", "m", "h", "hh")) %>% 
  ungroup()  %>% 
  pivot_wider(names_from = breaks, values_from = p) %>% 
  mutate(type = "post. pred.",
         x = 2 * (as.integer(as.factor(discipline_idx)) - 1) + gender_idx)

data_nwo %>% 
  group_by(discipline_idx, gender_idx) %>% 
  summarise(mean_data = mean(awards/applications),
            type = "data") %>% 
  mutate(x = 2 * (as.integer(discipline_idx) - 1) + gender_idx) %>% 
  ggplot(aes(x = x, y = mean_data)) +
  geom_segment(data = nwo_posterior_predictions,
               aes(x = x, xend = x,
                   y = ll, yend = hh),
               color = clr_current) +
  geom_point(data = nwo_posterior_predictions,
             aes(y = m, shape = factor(gender_idx)),
             color = clr_current,
             fill = clr_lighten(clr_current),
             size = 1.8) +
  geom_textline(aes(label = discipline_idx, group = discipline_idx),
                color = clr0dd, family = fnt_sel) +
  geom_point(aes(shape = factor(gender_idx)),
             color = clr0dd, fill = clr0, size = 1.8) +
  
  scale_shape_manual(values = c(`2` = 21, `1` = 19)) +
  scale_x_continuous(breaks = 2 * (seq_along(levels(data_nwo$discipline))) - .5,
                     labels = levels(data_nwo$discipline))+
  labs(x = "discipline", y = "awards") +
  theme(panel.background = element_rect(color = clr0,
                                        fill = "transparent"), 
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 90),
        legend.position = "bottom")
```


```{r}
nwo_dept_posterior_predictions <- sim(model_nwo_dept,
                                  data = nwo_grid) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(nwo_grid, .) %>% 
  pivot_longer(-c(discipline_idx, gender_idx, applications), values_to = "awards") %>% 
  dplyr::select(-name) %>% 
  group_by(discipline_idx, gender_idx, applications) %>% 
  summarise(p = quantile(awards/applications, probs = c(.055, .25, .5, .75, .955)),
            median = median(awards),
            breaks = c("ll", "l", "m", "h", "hh")) %>% 
  ungroup()  %>% 
  pivot_wider(names_from = breaks, values_from = p) %>% 
  mutate(type = "post. pred.",
         x = 2 * (as.integer(as.factor(discipline_idx)) - 1) + gender_idx)

data_nwo %>% 
  group_by(discipline_idx, gender_idx) %>% 
  summarise(mean_data = mean(awards/applications),
            type = "data") %>% 
  mutate(x = 2 * (as.integer(discipline_idx) - 1) + gender_idx) %>% 
  ggplot(aes(x = x, y = mean_data)) +
  geom_segment(data = nwo_dept_posterior_predictions,
               aes(x = x, xend = x,
                   y = ll, yend = hh),
               color = clr_current) +
  geom_point(data = nwo_dept_posterior_predictions,
             aes(y = m, shape = factor(gender_idx)),
             color = clr_current,
             fill = clr_lighten(clr_current),
             size = 1.8) +
  geom_textline(aes(label = discipline_idx, group = discipline_idx),
                color = clr0dd, family = fnt_sel) +
  geom_point(aes(shape = factor(gender_idx)),
             color = clr0dd, fill = clr0, size = 1.8) +
  
  scale_shape_manual(values = c(`2` = 21, `1` = 19)) +
  scale_x_continuous(breaks = 2 * (seq_along(levels(data_nwo$discipline))) - .5,
                     labels = levels(data_nwo$discipline))+
  labs(x = "discipline", y = "awards") +
  theme(panel.background = element_rect(color = clr0,
                                        fill = "transparent"), 
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(angle = 90),
        legend.position = "bottom")
```


**H5**

```{r H5}
dag_nwo_u <- dagify(
  A ~ D + G + U,
  D ~ G + U,
  exposure = "A",
  outcome = "G",
  coords = tibble(name = c("G", "D", "A", "U"),
                  x = c(0, .5, 1, 1),
                  y = c(0, 1, 0, 1)))

dag_nwo_u %>%
  fortify() %>% 
  mutate(stage = if_else(name == "A", "response",
                         if_else(name %in% c("G", "D"),
                                 "predictor", "confounds"))) %>% 
  plot_dag(clr_in = clr_current) +
  coord_fixed(ratio = .5) +
  scale_y_continuous(limits = c(-.1, 1.1)) +
  scale_x_continuous(limits = c(-.1, 1.1)) +
  theme(plot.title = element_text(hjust = .5, family = fnt_sel),
        plot.tag = element_text(family = fnt_sel))
```

Conditioning on `department` opens the collider between `G` and `U` - so this is not closing the backdoor path.

**H6**

```{r H6}
data("Primates301")
data_primates <- Primates301 %>% 
  as_tibble() %>% 
  mutate(brain_size_log = log(brain),
         research_effort_log = log(research_effort),
         across(c(brain_size_log, research_effort_log), standardize, .names = "{.col}_std")) %>% 
  dplyr::select(genus, species, spp_id, social_learning, research_effort, brain,
                brain_size_log:research_effort_log_std)
rm(Primates301)

data_primates_complete <- data_primates %>% 
  filter(complete.cases(brain, social_learning))
```

**a.**

```{r}
range(data_primates$social_learning,na.rm = TRUE)

ggplot() +
    stat_function(fun = function(x){dlnorm(x, 4, .5)},
                xlim = c(0, 250), n = 201,
                geom = "area", color = clr_current, fill = fll_current())
```

```{r, results = 'hide'}
data_primates_list <- data_primates_complete %>% 
  dplyr::select(spp_id, social_learning, brain_size_log_std) %>% 
  as.list()
  
model_primates <- ulam(
  flist = alist(
    social_learning ~ dpois(lambda),
    log(lambda) <- alpha + beta * brain_size_log_std,
    alpha ~ dnorm(4, 0.5),
    beta ~ dnorm(0, 0.2)
  ),
  data = data_primates_list,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```

```{r}
precis(model_primates) %>% 
  knit_precis()
```

```{r}
n <- 101
new_primates <- tibble(brain_size_log_std = seq(-3, 2.5, length.out = n))

primates_posterior_predictions <- link(model_primates,
                                       data = new_primates) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(new_primates, .) %>% 
  pivot_longer(-brain_size_log_std, values_to = "social_learning") %>% 
  dplyr::select(-name) %>% 
  group_by(brain_size_log_std) %>% 
  summarise(p = list(quantile(social_learning, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p)  %>%
  mutate(brain_size_log = brain_size_log_std * 
           sd(data_primates$brain_size_log, na.rm = TRUE) +
           mean(data_primates$brain_size_log, na.rm = TRUE),
         brain = exp(brain_size_log))

p1 <- primates_posterior_predictions %>% 
  ggplot(aes(x = brain_size_log_std, y = social_learning)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh),
              stat = "identity", size = .3,
              color = clr_current, fill = clr_alpha(clr_current)) +
  geom_point(data = data_primates_complete,
              color = clr_dark, fill = clr_alpha(clr_dark),
             shape = 21) +
  labs(subtitle = "on log scale")

p2 <- primates_posterior_predictions %>% 
  ggplot(aes(x = brain, y = social_learning)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh),
              stat = "identity", size = .3,
              color = clr_current, fill = clr_alpha(clr_current)) +
  geom_point(data = data_primates_complete,
              color = clr_dark, fill = clr_alpha(clr_dark),
             shape = 21) +
  labs(subtitle = "on natural scale")

p1 + p2
```

**b.**

```{r, results = 'hide'}
data_primates_effort_list <- data_primates_complete %>% 
  filter(complete.cases(research_effort_log_std)) %>% 
  dplyr::select(spp_id, social_learning, brain_size_log_std, research_effort_log_std) %>% 
  as.list()
  
model_primates_effort <- ulam(
  flist = alist(
    social_learning ~ dpois(lambda),
    log(lambda) <- alpha + beta_b * brain_size_log_std + beta_e * research_effort_log_std,
    alpha ~ dnorm(4, 0.5),
    beta_b ~ dnorm(0, 0.2),
    beta_e ~ dnorm(0, 0.2)
  ),
  data = data_primates_effort_list,
  chains = 4,
  cores = 4,
  log_lik = TRUE
)
```


```{r}
precis(model_primates_effort) %>%
  knit_precis()
```

$\rightarrow$ the more research is conducted on a species, the more likely it is to observe social learning events.

```{r}
compare(model_primates, model_primates_effort) %>% # ! differnet sample sizes.... 
  knit_precis(param_name = "model")
```

```{r}
n <- 101
grid_primates <- crossing(brain_size_log_std = c(-1.5, 0, 1.5),
                          research_effort_log_std = seq(-3, 3, length.out = n)) %>% 
  mutate(brain_class = cut(brain_size_log_std, c(-3.1, -1 , 1, 3.1)))

primates_posterior_predictions <- link(model_primates_effort,
                                       data = grid_primates) %>% 
  as.matrix() %>% 
  t() %>% 
  as_tibble() %>% 
  bind_cols(grid_primates, .) %>% 
  pivot_longer(-c(brain_size_log_std, research_effort_log_std, brain_class),
               values_to = "social_learning") %>% 
  dplyr::select(-name) %>% 
  group_by(brain_size_log_std, research_effort_log_std, brain_class) %>% 
  summarise(p = list(quantile(social_learning, probs = c(.055, .25, .5, .75, .955))),
            breaks = list(c("ll", "l", "m", "h", "hh"))) %>% 
  ungroup() %>% 
  unnest(c(p, breaks)) %>% 
  pivot_wider(names_from = breaks, values_from = p)  %>%
  mutate(brain_size_log = brain_size_log_std * 
           sd(data_primates$brain_size_log, na.rm = TRUE) +
           mean(data_primates$brain_size_log, na.rm = TRUE),
         brain = exp(brain_size_log))

primates_posterior_predictions %>% 
  ggplot(aes(x = research_effort_log_std, y = social_learning)) +
  geom_smooth(aes(ymin = ll, y = m, ymax = hh),
              stat = "identity", size = .3,
              color = clr_current, fill = clr_alpha(clr_current)) +
  geom_point(data = data_primates_complete %>% 
               mutate(brain_class = cut(brain_size_log_std,
                                        c(-3.1, -1 , 1, 3.1))),
              color = clr_dark, fill = clr_alpha(clr_dark),
             shape = 21) +
  facet_wrap(brain_class ~ . , labeller = label_both) +
  labs(subtitle = "on log scale")
```

**c.**

```{r}
data_primates %>% 
  dplyr::select(social_learning, brain_size_log_std, research_effort_log_std) %>% 
  ggpairs(
            lower = list(continuous = wrap(ggally_points, colour = clr_current, size = .9, alpha = .7)),
         diag = list(continuous = wrap(my_diag, fill = fll0, col = clr_darken(clr_current),
                                      color = clr0d, adjust = .7)),
        upper = list(continuous = wrap(my_upper ,
          size = 5, color = "black", family = fnt_sel)) ) 
```



```{r}
dag_primates <- dagify(
  S ~ B + R,
  R ~ B,
  exposure = "B",
  outcome = "S",
  coords = tibble(name = c("B", "R", "S"),
                  x = c(0, .5, 1),
                  y = c(0, 1, 0)))

dag_primates %>%
  fortify() %>% 
  mutate(stage = if_else(name == "S", "response",
                         if_else(name %in% c("B", "R"),
                                 "predictor", "confounds"))) %>% 
  plot_dag(clr_in = clr_current) +
  coord_fixed(ratio = .5) +
  scale_y_continuous(limits = c(-.1, 1.1)) +
  scale_x_continuous(limits = c(-.1, 1.1)) +
  theme(plot.title = element_text(hjust = .5, family = fnt_sel),
        plot.tag = element_text(family = fnt_sel))
```

## {brms} section

### Binomial Regression

> In the `brm()` formula syntax, including a `|` bar on the left side of a formula indicates we have extra supplementary information about our criterion. In this case, that information is that each `pulled_left` value corresponds to a single trial (i.e., `trials(1)`), which itself corresponds to the $n = 1$ portion of the statistical formula, above.

```{r}
brms_c11_model_omega10 <- brm(
  data = data_chimp, 
  family = binomial,
  pulled_left | trials(1) ~ 1,
  prior(normal(0, 10), 
        class = Intercept),
  seed = 42,
  sample_prior = TRUE,
  file = "brms/brms_c11_model_omega10")

brms_c11_model_omega15 <- brm(
  data = data_chimp, 
  family = binomial,
  pulled_left | trials(1) ~ 1,
  prior(normal(0, 1.5), 
        class = Intercept),
  seed = 42,
  sample_prior = TRUE,
  file = "brms/brms_c11_model_omega15")
```


```{r}
p1 <- bind_rows(prior_draws(brms_c11_model_omega10),
                prior_draws(brms_c11_model_omega15)) %>% 
  mutate(p = inv_logit_scaled(Intercept),
         w = factor(rep(c(10, 1.5), each = n() / 2),
                    levels = c(10, 1.5))) %>% 
  as_tibble() %>% 
  ggplot(aes(x = p, color = w)) +
  geom_density(size = .3, adjust = 0.1,
               aes(fill = after_scale(clr_alpha(color)))) +
  scale_color_manual(expression(italic(w)),
                     values = c(`1.5` = clr0d,
                                `10` = clr_dark)) +
  scale_y_continuous("denstiy") +
  labs(title = expression(alpha%~%Normal(0*", "*italic(w))),
       x = "prior prob pull left")
```


```{r}
data_chimp_f <- data_chimp %>% 
  mutate(actor = factor(actor),
         treatment = factor(treatment),
         side_idx = factor(side_idx),
         condition_idx = factor(condition_idx))

brms_c11_model_beta10 <- brm(
  data = data_chimp_f, 
  family = binomial,
  bf(pulled_left | trials(1) ~ a + b,
     a ~ 1, 
     b ~ 0 + treatment,
     nl = TRUE),
  prior = c(prior(normal(0, 1.5), nlpar = a),
            prior(normal(0, 10), nlpar = b, coef = treatment1),
            prior(normal(0, 10), nlpar = b, coef = treatment2),
            prior(normal(0, 10), nlpar = b, coef = treatment3),
            prior(normal(0, 10), nlpar = b, coef = treatment4)),
  iter = 2000, warmup = 1000, 
  chains = 4, cores = 4,
  seed = 42,
  sample_prior = TRUE,
  file = "brms/brms_c11_model_beta10")

brms_c11_model_beta05 <- brm(
  data = data_chimp_f, 
  family = binomial,
  bf(pulled_left | trials(1) ~ a + b,
     a ~ 1, 
     b ~ 0 + treatment,
     nl = TRUE),
  prior = c(prior(normal(0, 1.5), nlpar = a),
            prior(normal(0, 0.5), nlpar = b, coef = treatment1),
            prior(normal(0, 0.5), nlpar = b, coef = treatment2),
            prior(normal(0, 0.5), nlpar = b, coef = treatment3),
            prior(normal(0, 0.5), nlpar = b, coef = treatment4)),
  iter = 2000, warmup = 1000, 
  chains = 4, cores = 4,
  seed = 42,
  sample_prior = TRUE,
  file = "brms/brms_c11_model_beta05")
```

> You could have just included a single line, `prior(normal(0, 0.5), nlpar = b)`, that did not include a coef argument. The problem with this approach is wed only get one column for treatment when using the `prior_samples()` function to retrieve the prior samples. To get separate columns for the prior samples of each of the levels of treatment, you need to take the verbose approach, above.

```{r}
prior_beta <- bind_rows(prior_draws(brms_c11_model_beta10),
                        prior_draws(brms_c11_model_beta05)) %>% 
  mutate(w  = factor(rep(c(10, 0.5), each = n() / 2),
                     levels = c(10, 0.5)),
         p1 = inv_logit_scaled(b_a + b_b_treatment1),
         p2 = inv_logit_scaled(b_a + b_b_treatment2)) %>% 
  mutate(diff = abs(p1 - p2)) 

p2 <- prior_beta %>% 
  ggplot(aes(x = diff, color = w)) +
  geom_density(size = .3, adjust = 0.1,
               aes(fill = after_scale(clr_alpha(color)))) +
  scale_color_manual(expression(italic(beta)),
                     values = c(`0.5` = clr0d,
                                `10` = clr_dark)) +
  labs(title = expression(alpha%~%Normal(0*", "*italic(w))),
       x = "prior diff between treatments")

p1 + p2 &
  theme(legend.position = "bottom")
```

```{r}
prior_beta %>% 
  group_by(w) %>% 
  summarise(mean = mean(diff))
```


```{r}
as_draws_df(brms_c11_model_beta05) %>%
    as_tibble() %>% 
    transmute(alpha = inv_logit_scaled(b_a_Intercept)) %>% 
    mean_qi()
```

Empirical treatment means:

```{r}
data_chimp_f %>% 
  group_by(treatment) %>% 
  summarise(mean = mean(pulled_left))
```

Posterior:

```{r}
posterior_samples(brms_c11_model_beta05) %>% 
  pivot_longer(b_b_treatment1:b_b_treatment4) %>% 
  mutate(treatment = str_remove(name, "b_b_treatment"),
         mean      = inv_logit_scaled(b_a_Intercept + value)) %>%
  group_by(treatment) %>% 
  mean_qi(mean)
```


```{r}
brms_c11_model_chimp <- brm(
  data = data_chimp_f, 
  family = binomial,
  bf(pulled_left | trials(1) ~ a + b,
     a ~ 0 + actor, 
     b ~ 0 + treatment,
     nl = TRUE),
  prior = c(prior(normal(0, 1.5), nlpar = a),
            prior(normal(0, 0.5), nlpar = b)),
  iter = 2000, warmup = 1000,
  chains = 4, cores = 4,
  seed = 42,
  file = "brms/brms_c11_model_chimp")
```

```{r}
chimp_post <- as_draws_df(brms_c11_model_chimp)

p1 <- chimp_post %>% 
  pivot_longer(contains("actor")) %>%
  mutate(probability = inv_logit_scaled(value),
         actor       = factor(str_remove(name, "b_a_actor"),
                              levels = 7:1)) %>% 
  ggplot(aes(x = probability, y = actor)) +
  geom_vline(xintercept = .5, color = clr_dark, linetype = 3) +
  stat_pointinterval(.width = .95, size = 1/2,
                     color = clr_dark) +
  scale_x_continuous(expression(alpha[actor]), limits = 0:1) +
  ylab(NULL) +
  theme(axis.ticks.y = element_blank())

p2 <- chimp_post %>% 
  dplyr::select(contains("treatment")) %>% 
  set_names("R|N","L|N","R|P","L|P") %>% 
  mutate(`R|diff`= `R|N` - `R|P`,
         `L|diff`= `L|N` - `L|P`) %>% 
  pivot_longer(everything()) %>%
  mutate(probability = inv_logit_scaled(value),
         treatment   = factor(name, levels = treatment_labels)) %>% 
  mutate(treatment = fct_rev(treatment)) %>% 
  ggplot(aes(x = value, y = treatment)) +
  geom_rect(data = tibble(x = 1),inherit.aes = FALSE,
            aes(xmin = -Inf, xmax = Inf, ymin = .5, ymax = 2.5),
            color = clr0d, fill = fll0) +
  geom_vline(xintercept = 0, color = clr_dark, linetype = 3) +
  stat_pointinterval(.width = .95, size = 1/2,
                     color = clr_dark) +
  labs(x = expression(beta[treatment]),
       y = NULL) +
  scale_y_discrete() +
  theme(axis.ticks.y = element_blank())

p1 + p2
```

```{r}
p1 <- data_chimp_f %>%
  group_by(actor, treatment) %>%
  summarise(proportion = mean(pulled_left)) %>% 
  left_join(data_chimp_f %>%
              mutate(labels = treatment_labels[treatment]) %>% 
              distinct(actor, treatment, labels, condition, prosoc_left),
            by = c("actor", "treatment")) %>% 
  mutate(condition = factor(condition)) %>% 
  
  ggplot(aes(x = labels, y = proportion)) +
  geom_hline(yintercept = .5, color = clr_dark, linetype = 3) +
  geom_line(aes(group = prosoc_left),
            size = 1/4, color = clr_dark) +
  geom_point(aes(color = condition),
             size = 2.5, show.legend = F) + 
  labs(subtitle = "observed proportions") 

new_chimp <- data_chimp_f %>% 
  distinct(actor, treatment, condition_idx, prosoc_left)

p2 <- fitted(brms_c11_model_chimp,
         newdata = new_chimp) %>% 
  as_tibble() %>% 
  bind_cols(new_chimp) %>% 
  mutate(labels = treatment_labels[treatment]) %>% 
  ggplot(aes(x = labels, y = Estimate, ymin = Q2.5, ymax = Q97.5)) +
  geom_hline(yintercept = .5, color = clr_dark, linetype = 3) +
  geom_line(aes(group = prosoc_left),
            size = 1/4, color = clr_dark) +
  geom_pointrange(aes(color = condition_idx),
                  fatten = 2.5, show.legend = F) + 
  labs(subtitle = "posterior predictions")

p1 + p2 +
  plot_layout( ncol = 1 ) & 
  scale_color_manual(values = c(clr0d, clr_dark)) &
  scale_y_continuous("proportion left lever", 
                     breaks = c(0, .5, 1), limits = c(0, 1)) &
  facet_wrap(~ actor, nrow = 1, labeller = label_both) &
  xlab(NULL) &
  theme(axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.background = element_rect(fill = "transparent",
                                        color = clr0)) 

```

```{r}
brms_c11_model_chimp_no_interaction <- brm(
  data = data_chimp_f, 
  family = binomial,
  bf(pulled_left | trials(1) ~ a + bs + bc,
     a ~ 0 + actor, 
     bs ~ 0 + side_idx, 
     bc ~ 0 + condition_idx,
     nl = TRUE),
  prior = c(prior(normal(0, 1.5), nlpar = a),
            prior(normal(0, 0.5), nlpar = bs),
            prior(normal(0, 0.5), nlpar = bc)),
  iter = 2000, warmup = 1000,
  chains = 4, cores = 4,
  seed = 42,
  file = "brms/brms_c11_model_chimp_no_interaction")
```

```{r}
brms_c11_model_chimp <- add_criterion(brms_c11_model_chimp, c("loo", "waic"))
brms_c11_model_chimp_no_interaction <- add_criterion(brms_c11_model_chimp_no_interaction,
                                                     c("loo", "waic"))

loo_compare(brms_c11_model_chimp,
            brms_c11_model_chimp_no_interaction,
            criterion = "loo") %>%
  print(simplify = FALSE)

model_weights(brms_c11_model_chimp,
              brms_c11_model_chimp_no_interaction, weights = "loo") %>%
  round(digits = 2)
```

```{r, fig.asp = .3}
new_chimp <- data_chimp_f %>% 
  distinct(actor, treatment, condition_idx, side_idx, prosoc_left)

fitted(brms_c11_model_chimp_no_interaction,
       newdata = new_chimp) %>% 
  data.frame() %>% 
  bind_cols(new_chimp) %>% 
  mutate(labels = treatment_labels[treatment]) %>% 
  ggplot(aes(x = labels, y = Estimate, ymin = Q2.5, ymax = Q97.5)) +
  geom_hline(yintercept = .5, color = clr_dark) +
  geom_line(aes(group = side_idx),
            size = 1/4, color = clr_dark) +
  geom_pointrange(aes(color = condition_idx),
                  fatten = 2.5, show.legend = FALSE) + 
  scale_color_manual(values = c(clr0, clr_dark)) +
  scale_y_continuous("proportion left lever", 
                     breaks = c(0, .5, 1), limits = c(0, 1)) +
  labs(subtitle = "posterior predictions for b11.5",
       x = NULL) +
  facet_wrap(~ actor, nrow = 1, labeller = label_both)+
  theme(axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.background = element_rect(fill = "transparent",
                                        color = clr0))
```


```{r}
log_lik(brms_c11_model_chimp) %>% 
  str()
```

#### Relative Shark and Absolute Deer

```{r}
as_draws_df(brms_c11_model_chimp) %>% 
  mutate(proportional_odds = exp(b_b_treatment4 - b_b_treatment2)) %>% 
  mean_qi(proportional_odds)
```

####  Aggregated Binomial: Chimpanzees Again, Condensed

```{r}
data_chimp_aggregated_f <- data_chimp_f %>% 
  group_by(treatment, actor, side_idx, condition_idx) %>% 
  summarise(left_pulls = sum(pulled_left))  %>% 
  ungroup()
```

```{r}
brms_c11_model_chimp_aggregated <- brm(
  data = data_chimp_aggregated_f, 
  family = binomial,
  bf(left_pulls | trials(18) ~ a + b,
     a ~ 0 + actor, 
     b ~ 0 + treatment,
     nl = TRUE),
  prior = c(prior(normal(0, 1.5), nlpar = a),
            prior(normal(0, 0.5), nlpar = b)),
  iter = 2000, warmup = 1000,
  chains = 4, cores = 4,
  seed = 42,
  file = "brms/brms_c11_model_chimp_aggregated")
```


```{r}
bind_rows(as_draws_df(brms_c11_model_chimp),
          as_draws_df(brms_c11_model_chimp_aggregated)) %>% 
  mutate(fit = rep(c("brms_c11_model_chimp", 
                     "brms_c11_model_chimp_aggregated"),
                   each = n() / 2)) %>% 
  pivot_longer(b_a_actor1:b_b_treatment4) %>% 
  # plot
  ggplot(aes(x = value, y = name, color = fit)) +
  stat_pointinterval(.width = .95, size = 2/3,
                     position = position_dodge(width = .8)) +
  scale_color_manual(values = c(clr0d, clr_dark)) +
  labs(x = "posterior (log-odds scale)",
       y = NULL) +
  theme(axis.ticks.y = element_blank(),
        legend.position = "bottom")
```


```{r, error = TRUE}
brms_c11_model_chimp <- add_criterion(brms_c11_model_chimp, "loo")
brms_c11_model_chimp_aggregated <- add_criterion(brms_c11_model_chimp_aggregated, "loo")

loo_compare(brms_c11_model_chimp,
            brms_c11_model_chimp_aggregated,
            criterion = "loo") %>%
  # does not work on models based on different data - that is the point here...
  print(simplify = FALSE) 
```


```{r}
loo(brms_c11_model_chimp_aggregated) %>% 
  loo::pareto_k_table() %>% 
  as.data.frame() %>% 
  knitr::kable()

```

#### Aggregated Binomial: Graduate School Admissions

```{r}
data_ucb_f <- data_ucb %>% 
  mutate(gid  = factor(applicant.gender, levels = c("male", "female")),
         case = factor(1:n()))
```

```{r}
brms_c11_model_ucb <- brm(
  data = data_ucb_f, 
  family = binomial,
  admit | trials(applications) ~ 0 + gid,
  prior(normal(0, 1.5), class = b),
  iter = 2000, warmup = 1000,
  cores = 4, chains = 4,
  seed = 42,
  file = "brms/brms_c11_model_ucb") 
```

```{r}
as_draws_df(brms_c11_model_ucb) %>% 
  mutate(diff_a = b_gidmale - b_gidfemale,
         diff_p = inv_logit_scaled(b_gidmale) - inv_logit_scaled(b_gidfemale)) %>% 
  pivot_longer(contains("diff")) %>% 
  group_by(name) %>% 
  mean_qi(value, .width = .89)
```

```{r}
ucb_predict <- predict(brms_c11_model_ucb) %>% 
  data.frame() %>% 
  bind_cols(data_ucb_f)

ucb_predict %>% 
  ggplot(aes(x = case, y = admit / applications)) +
  geom_pointrange(aes(y    = Estimate / applications,
                      ymin = Q2.5     / applications ,
                      ymax = Q97.5    / applications),
                  color = clr0d, fill = clr0,
                  shape = 21) +
  geom_point(color = clr_dark) +
  geom_textpath(aes(group = dept, label = dept),
            color = clr_dark, family = fnt_sel) +
  scale_y_continuous("Proportion admitted", limits = 0:1) +
  ggtitle("Posterior validation check") +
  theme(axis.ticks.x = element_blank())
```


```{r}
model_ucb_dept

brms_c11_model_ucb_dept <- brm(
  data = data_ucb_f, 
  family = binomial,
  bf(admit | trials(applications) ~ a + d,
     a ~ 0 + gid, 
     d ~ 0 + dept,
     nl = TRUE),
  prior = c(prior(normal(0, 1.5), nlpar = a),
            prior(normal(0, 1.5), nlpar = d)),
  iter = 2000, warmup = 1000,
  cores = 4, chains = 4,
  seed = 42,
  file = "brms/brms_c11_model_ucb_dept") 
```

```{r}
ucb_dept_predict <- predict(brms_c11_model_ucb_dept) %>% 
  data.frame() %>% 
  bind_cols(data_ucb_f)

ucb_dept_predict %>% 
  ggplot(aes(x = case, y = admit / applications)) +
  geom_pointrange(aes(y    = Estimate / applications,
                      ymin = Q2.5     / applications ,
                      ymax = Q97.5    / applications),
                  color = clr0d, fill = clr0,
                  shape = 21) +
  geom_point(color = clr_dark) +
  geom_textpath(aes(group = dept, label = dept),
            color = clr_dark, family = fnt_sel) +
  scale_y_continuous("Proportion admitted", limits = 0:1) +
  ggtitle("Posterior validation check") +
  theme(axis.ticks.x = element_blank())
```


```{r, fig.width = 11, fig.asp = .75}
as_draws_df(brms_c11_model_ucb_dept) %>% 
  as_tibble() %>% 
  dplyr::select(starts_with("b_a") | starts_with("b_d")) %>% 
  set_names(c("alpha[male]", "alpha[female]", 
              str_c("delta[", LETTERS[1:6], "]"))) %>%
  ggpairs(
     lower = list(continuous = wrap(my_lower, col = clr_dark)),
        diag = list(continuous = wrap(my_diag, fill = fll0,
                                      col = clr_dark, adjust = .7)),
        upper = list(continuous = wrap(my_upper ,
          size = 3, col = "black", family = fnt_sel)) ) +
    theme(panel.border = element_rect(color = clr_dark, fill = "transparent"))
```

### Poisson Regression

```{r}
data_kline_f <- data_kline %>% 
  mutate(contact_idx = factor(contact_idx))

# intercept only
brms_c11_model_ocean_intercept <- brm(
  data = data_kline_f, 
  family = poisson,
  total_tools ~ 1,
  prior(normal(3, 0.5), class = Intercept),
  iter = 2000, warmup = 1000,
  chains = 4, cores = 4,
  seed = 42,
  file = "brms/brms_c11_model_ocean_intercept") 

# interaction model
brms_c11_model_ocean_interact <- brm(
  data = data_kline_f, 
  family = poisson,
  bf(total_tools ~ a + b * pop_log_scl,
     a + b ~ 0 + contact_idx,
     nl = TRUE),
  prior = c(prior(normal(3, 0.5), nlpar = a),
            prior(normal(0, 0.2), nlpar = b)),
  iter = 2000, warmup = 1000,
  chains = 4, cores = 4,
  seed = 42,
  file = "brms/brms_c11_model_ocean_interact") 
```


```{r}
brms_c11_model_ocean_intercept  <- add_criterion(brms_c11_model_ocean_intercept, "loo")
brms_c11_model_ocean_interact <- add_criterion(brms_c11_model_ocean_interact, "loo")

loo_compare(brms_c11_model_ocean_intercept,
            brms_c11_model_ocean_interact,
            criterion = "loo") %>%
  print(simplify = FALSE)
```

```{r}
model_weights(brms_c11_model_ocean_intercept,
            brms_c11_model_ocean_interact,
            weights = "loo") %>%
  round(digits = 2)
```

```{r}
loo(brms_c11_model_ocean_interact) %>% 
  loo::pareto_k_table() %>% 
  as.data.frame()%>% 
  knitr::kable()
```

```{r}
tibble(culture = data_kline_f$culture,
       k       = brms_c11_model_ocean_interact$criteria$loo$diagnostics$pareto_k) %>% 
  arrange(desc(k)) %>% 
  mutate_if(is.double, round, digits = 2)
```

```{r}
high_k_cultures <- tibble(culture = data_kline_f$culture,
       k = brms_c11_model_ocean_interact$criteria$loo$diagnostics$pareto_k) %>% 
  arrange(desc(k)) %>% 
  filter(k > .5) %>% 
  transmute(culture) %>% 
  unlist()

new_ocean <- distinct(data_kline_f, contact_idx) %>% 
  expand(contact_idx, 
         pop_log_scl = seq(from = -2, to = 2.5, length.out = 100))

ocean_fitted <- fitted(brms_c11_model_ocean_interact,
         newdata = new_ocean,
         probs = c(.055, .945)) %>%
  data.frame() %>%
  bind_cols(new_ocean)

p1 <- ocean_fitted %>%
  ggplot(aes(x = pop_log_scl, group = contact_idx, color = contact_idx)) +
  geom_smooth(aes(y = Estimate, ymin = Q5.5, ymax = Q94.5, fill = after_scale(clr_alpha(color))),
              stat = "identity",
              alpha = 1/4, size = 1/2) +
  geom_point(data = bind_cols(data_kline_f,
                              brms_c11_model_ocean_interact$criteria$loo$diagnostics),
             aes(y = total_tools, size = pareto_k, fill = after_scale(clr_alpha(color))),
             shape = 21) +
  ggrepel::geom_text_repel(data = 
                    bind_cols(data_kline_f,
                              brms_c11_model_ocean_interact$criteria$loo$diagnostics) %>% 
                    filter(culture %in% high_k_cultures) %>% 
                    mutate(label = str_c(culture, " (", round(pareto_k, digits = 2), ")")),
                  aes(y = total_tools, label = label), 
                  size = 3, seed = 11, color = "black", family = fnt_sel) +
  labs(x = "log population (std)",
       y = "total tools") +
  coord_cartesian(xlim = range(brms_c11_model_ocean_interact$data$pop_log_scl),
                  ylim = c(0, 80))


p2 <- ocean_fitted %>%
  mutate(population = exp((pop_log_scl * sd(log(data_kline_f$population))) +
                            mean(log(data_kline_f$population)))) %>% 
  ggplot(aes(x = population, group = contact_idx, color = contact_idx)) +
  geom_smooth(aes(y = Estimate, ymin = Q5.5, ymax = Q94.5, fill = after_scale(clr_alpha(color))),
              stat = "identity",
              alpha = 1/4, size = 1/2) +
  geom_point(data = bind_cols(data_kline_f,
                              brms_c11_model_ocean_interact$criteria$loo$diagnostics),
             aes(y = total_tools, size = pareto_k, fill = after_scale(clr_alpha(color))),
             shape = 21) +
  scale_x_continuous("population", breaks = c(0, 50000, 150000, 250000)) +
  ylab("total tools") +
  coord_cartesian(xlim = range(data_kline_f$population),
                  ylim = c(0, 80))

p1 + p2 &
  scale_color_manual(values = c(clr_dark, clr0dd)) &
  scale_size(range = c(2, 5)) &
  theme(legend.position = "none")
```

#### Modeling Tool Innovation

```{r}
brms_c11_model_ocean_scientific <- brm(
  data = data_kline_f, 
  family = poisson(link = "identity"),
  bf(total_tools ~ exp(a) * population^b / g,
     a + b ~ 0 + contact_idx,
     g ~ 1,
     nl = TRUE),
  prior = c(prior(normal(1, 1), nlpar = a),
            prior(exponential(1), nlpar = b, lb = 0),
            prior(exponential(1), nlpar = g, lb = 0)),
  iter = 2000, warmup = 1000, chains = 4, cores = 4,
  seed = 42,
  control = list(adapt_delta = .95),
  file = "brms/brms_c11_model_ocean_scientific") 
```

```{r}
brms_c11_model_ocean_scientific <- add_criterion(brms_c11_model_ocean_scientific,
                                                 criterion = "loo",
                                                 moment_match = TRUE)
loo(brms_c11_model_ocean_scientific)
```

```{r}
new_oceans <- distinct(data_kline_f, contact_idx) %>% 
  expand(contact_idx, 
         population = seq(from = 0, to = 300000, length.out = 100))

# compute the poster predictions for lambda
fitted(brms_c11_model_ocean_scientific,
       newdata = new_oceans,
       probs = c(.055, .945)) %>%
  data.frame() %>%
  bind_cols(new_oceans) %>%
  ggplot(aes(x = population, group = contact_idx, color = contact_idx)) +
  geom_smooth(aes(y = Estimate, ymin = Q5.5, ymax = Q94.5, fill = after_scale(clr_alpha(color))),
              stat = "identity",
              alpha = 1/4, size = 1/2) +
  geom_point(data = bind_cols(data_kline_f,
                              brms_c11_model_ocean_scientific$criteria$loo$diagnostics),
             aes(y = total_tools, size = pareto_k, fill = after_scale(clr_alpha(color))),
             shape = 21) +
  scale_color_manual(values = c(clr_dark, clr0dd)) +
  scale_size(range = c(2, 5)) +
  scale_x_continuous("population", breaks = c(0, 50000, 150000, 250000)) +
  ylab("total tools") +
  coord_cartesian(xlim = range(data_kline_f$population),
                  ylim = range(data_kline_f$total_tools)) +
  theme(legend.position = "none")
```

#### Exposure and the Offset

```{r}
data_books_f <- data_books %>% 
  mutate(monastery = factor(monastery))

brms_c11_model_books <- brm(
  data = data_books_f, 
  family = poisson,
  books ~ 1 + offset(days_log) + monastery,
  prior = c(prior(normal(0, 1), class = Intercept),
            prior(normal(0, 1), class = b)),
  iter = 2000, warmup = 1000,
  cores = 4, chains = 4,
  seed = 42,
  file = "brms/brms_c11_model_books")
```


```{r}
as_draws_df(brms_c11_model_books) %>%
  mutate(lambda_old = exp(b_Intercept),
         lambda_new = exp(b_Intercept + b_monastery1)) %>%
  pivot_longer(contains("lambda")) %>% 
  mutate(name = factor(name, levels = c("lambda_old", "lambda_new"))) %>%
  group_by(name) %>%
  mean_hdi(value, .width = .89) %>% 
  mutate_if(is.double, round, digits = 2)
```

### Multinomial and Categorical Models

#### Predictors Matched to Outcomes

```{r, fig.asp = .35}
data_career <- tibble(career = factor(career)) %>% 
  mutate(career_income = ifelse(career == 3, 5, career))

data_career %>%
  ggplot(aes(y = career)) +
  ggstance::geom_barh(size = 0, fill = clr_dark)
```

```{r}
brms_c11_model_career <- brm(
  data = data_career, 
  family = categorical(link = logit, refcat = 3),
  career ~ 1,
  prior = c(prior(normal(0, 1), class = Intercept, dpar = mu1),
            prior(normal(0, 1), class = Intercept, dpar = mu2)),
  iter = 2000, warmup = 1000,
  cores = 4, chains = 4,
  seed = 42,
  file = "brms/brms_c11_model_career")
```

```{r}
fitted(brms_c11_model_career)[1, , ] %>% 
  round(digits = 2) %>% 
  t() %>% 
  as.data.frame() %>% 
  knitr::kable()
```

```{r}
tibble(income = c(1, 2, 5)) %>% 
  mutate(score = 0.5 * income) %>% 
  mutate(p = exp(score) / sum(exp(score)))
```

```{r}
as_draws_df(brms_c11_model_career) %>% 
  mutate(b_mu3_Intercept = 0) %>% 
  mutate(p1 = exp(b_mu1_Intercept) / (exp(b_mu1_Intercept) + exp(b_mu2_Intercept) + exp(b_mu3_Intercept)),
         p2 = exp(b_mu2_Intercept) / (exp(b_mu1_Intercept) + exp(b_mu2_Intercept) + exp(b_mu3_Intercept)),
         p3 = exp(b_mu3_Intercept) / (exp(b_mu1_Intercept) + exp(b_mu2_Intercept) + exp(b_mu3_Intercept))) %>% 
  pivot_longer(p1:p3) %>% 
  group_by(name) %>% 
  mean_qi(value) %>% 
  mutate_if(is.double, round, digits = 2)
```

Two alternative ways to fit multinomial models within {brms}:

```{r}
# verbose syntax
brms_c11_model_career_verbose <- brm(
  data = data_career, 
  family = categorical(link = logit, refcat = 3),
  bf(career ~ 1,
     mu1 ~ 1,
     mu2 ~ 1),
  prior = c(prior(normal(0, 1), class = Intercept, dpar = mu1),
            prior(normal(0, 1), class = Intercept, dpar = mu2)),
  iter = 2000, warmup = 1000,
  cores = 4, chains = 4,
  seed = 42,
  file = "brms/brms_c11_model_career_verbose")

# nonlinear syntax
brms_c11_model_career_nonlinear <- brm(
  data = data_career, 
  family = categorical(link = logit, refcat = 3),
  bf(career ~ 1,
     nlf(mu1 ~ a1),
     nlf(mu2 ~ a2),
     a1 + a2 ~ 1),
  prior = c(prior(normal(0, 1), class = b, nlpar = a1),
            prior(normal(0, 1), class = b, nlpar = a2)),
  iter = 2000, warmup = 1000,
  cores = 4, chains = 4,
  seed = 42,
  file = "brms/brms_c11_model_career_nonlinear")
```

```{r}
brms_c11_model_career_d <- brm(
  data = data_career, 
  family = categorical(link = logit, refcat = 3),
  bf(career ~ 1,
     nlf(mu1 ~ a1 + b * 1),
     nlf(mu2 ~ a2 + b * 2),
     a1 + a2 + b ~ 1),
  prior = c(prior(normal(0, 1), class = b, nlpar = a1),
            prior(normal(0, 1), class = b, nlpar = a2),
            prior(normal(0, 0.5), class = b, nlpar = b, lb = 0)),
  iter = 2000, warmup = 1000,
  cores = 4, chains = 4,
  seed = 42,
  control = list(adapt_delta = .99),
      file = "brms/brms_c11_model_career_d")
```

```{r}
income <- c(1, 2, 5)
as_draws_df(brms_c11_model_career_d) %>% 
  transmute(s1 = b_a1_Intercept + b_b_Intercept * income[1],
            s2_orig = b_a2_Intercept + b_b_Intercept * income[2],
            s2_new  = b_a2_Intercept + b_b_Intercept * income[2] * 2) %>% 
  mutate(p_orig = purrr::map2_dbl(s1, s2_orig, ~softmax(.x, .y, 0)[2]),
         p_new  = purrr::map2_dbl(s1, s2_new, ~softmax(.x, .y, 0)[2])) %>% 
  mutate(p_diff = p_new - p_orig) %>% 
  mean_qi(p_diff) %>% 
  mutate_if(is.double, round, digits = 2)
```

#### Predictors Matched to Observations

```{r}
p1 <- career_fam %>% 
  mutate(career = as.factor(career)) %>% 
  ggplot(aes(x = family_income, color = career)) +
  geom_density(size = .5, aes(fill = after_scale(clr_alpha(color)))) +
  scale_color_manual(values = c(clr_dark, clr0dd, clr0)) +
  theme(legend.position = "none")


p2 <- career_fam %>% 
  mutate(career = as.factor(career)) %>%
  mutate(fi = santoku::chop_width(family_income,
                                  width = .1,
                                  start = 0, 
                                  labels = 1:10)) %>% 
  count(fi, career) %>% 
  group_by(fi) %>% 
  mutate(proportion = n / sum(n)) %>% 
  mutate(f = as.double(fi)) %>% 
  ggplot(aes(x = (f - 1) / 9, y = proportion, color = career)) +
  geom_area(size = .5, aes(fill = after_scale(clr_alpha(color)))) +
  scale_color_manual(values = c(clr_dark, clr0dd, clr0)) +
  xlab("family_income, descritized")

p1 + p2 +
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")
```

```{r}
data_career_fam <- career_fam %>% 
  dplyr::select(family_income, career) %>% 
  mutate(career = factor(career))

brms_c11_model_career_fam <- brm(
  data = data_career_fam, 
  family = categorical(link = logit, refcat = 3),
  bf(career ~ 1,
     nlf(mu1 ~ a1 + b1 * family_income),
     nlf(mu2 ~ a2 + b2 * family_income),
     a1 + a2 + b1 + b2 ~ 1),
  prior = c(prior(normal(0, 1.5), class = b, nlpar = a1),
            prior(normal(0, 1.5), class = b, nlpar = a2),
            prior(normal(0, 1), class = b, nlpar = b1),
            prior(normal(0, 1), class = b, nlpar = b2)),
  iter = 2000, warmup = 1000,
  cores = 4, chains = 4,
  seed = 42,
  file = "brms/brms_c11_model_career_fam")
```

```{r}
brms_c11_model_career_fam <- add_criterion(brms_c11_model_career_fam, "loo")

loo(brms_c11_model_career_fam)
```

```{r}
new_careers <- tibble(family_income = seq(from = 0, to = 1, length.out = 60))

careers_fitted <- fitted(brms_c11_model_career_fam,
         newdata = new_careers)
```

```{r}
p1 <- rbind(careers_fitted[, , 1],
            careers_fitted[, , 2],
            careers_fitted[, , 3]) %>% 
  as_tibble() %>% 
  bind_cols(new_careers %>% expand(career = 1:3, family_income)) %>% 
  mutate(career = str_c("career: ", career)) %>% 
  ggplot(aes(x = family_income, y = Estimate,
             ymin = Q2.5, ymax = Q97.5,
             color = career)) +
  geom_smooth(size = .5, stat = "identity", aes(fill = after_scale(clr_alpha(color)))) +
  scale_color_manual(values = c(clr_dark, clr0dd, clr0)) +
  scale_x_continuous(breaks = 0:2 / 2) +
  scale_y_continuous("probability", limits = c(0, 1),
                     breaks = 0:3 / 3, labels = c("0", ".33", ".67", "1")) +
  theme(axis.text.y = element_text(hjust = 0),
        legend.position = "none") 

p2 <- rbind(careers_fitted[, , 1],
      careers_fitted[, , 2],
      careers_fitted[, , 3]) %>% 
  as_tibble() %>% 
  bind_cols(new_careers %>% expand(career = 1:3, family_income)) %>% 
  group_by(family_income) %>% 
  mutate(proportion = Estimate / sum(Estimate),
         career = factor(career)) %>% 
  ggplot(aes(x = family_income, y = proportion)) +
  geom_area(aes(color = career, fill = after_scale(clr_alpha(color))),
            size = .5) +
  scale_color_manual(values = c(clr_dark, clr0dd, clr0))

p1 + p2 +
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")
```

#### Multinomial in Disguise as Poisson

```{r}
data_ucb_f <- data_ucb %>% 
  rename(rej = "reject") # 'reject' is a reserved word

brms_c11_model_ucb_binom <- brm(
  data = data_ucb_f, 
      family = binomial,
      admit | trials(applications) ~ 1,
      prior(normal(0, 1.5), class = Intercept),
      iter = 2000, warmup = 1000,
      cores = 3, chains = 3,
      seed = 42,
      file = "brms/brms_c11_model_ucb_binom")

# Poisson model of overall admission rate and rejection rate
brms_c11_model_ucb_poisson <- brm(
  data = data_ucb_f, 
      family = poisson,
      mvbind(admit, rej) ~ 1,
      prior(normal(0, 1.5), class = Intercept),
      iter = 2000, warmup = 1000,
      cores = 3, chains = 3,
      seed = 42,
      file = "brms/brms_c11_model_ucb_poisson")
```

```{r}
ucb_posterior <- as_draws_df(brms_c11_model_ucb_poisson)

ucb_posterior %>%
  mutate(admit  = exp(b_admit_Intercept), 
         reject = exp(b_rej_Intercept)) %>% 
  pivot_longer(admit:reject) %>% 
  ggplot(aes(x = value, y = name, fill = name)) +
  stat_halfeye(point_interval = median_qi, .width = .95,
               color = "black") +
  scale_fill_manual(values = c(clr0d, clr_dark)) +
  labs(title = " Mean admit/reject rates across departments",
       x = "# applications",
       y = NULL) +
  theme(axis.ticks.y = element_blank(),
        legend.position = "none")
```

```{r}
summary(brms_c11_model_ucb_binom)$fixed
summary(brms_c11_model_ucb_poisson)$fixed
```

```{r}
fixef(brms_c11_model_ucb_binom)[, "Estimate"] %>%
  inv_logit_scaled()
```

```{r}
k <- fixef(brms_c11_model_ucb_poisson) %>%
  as.numeric()

exp(k[1]) / (exp(k[1]) + exp(k[2]))
```

```{r}
bind_cols(
  as_draws_df(brms_c11_model_ucb_poisson) %>% 
    mutate(`the Poisson` = exp(b_admit_Intercept) /
             (exp(b_admit_Intercept) + 
                exp(b_rej_Intercept))),
  as_draws_df(brms_c11_model_ucb_binom) %>% 
    mutate(`the binomial` = inv_logit_scaled(b_Intercept)) ) %>% 
  pivot_longer(starts_with("the")) %>% 
  ggplot(aes(x = value, y = name, fill = name)) +
  stat_halfeye(point_interval = median_qi, .width = c(.89, .5),
               color = "black") +
  scale_fill_manual(values = c(clr0d, clr_dark)) +
  labs(title = "Two models, same marginal posterior",
       x = "admissions probability",
       y = NULL) +
  coord_cartesian(ylim = c(1.5, 2.35)) +
  theme(axis.text.y = element_text(hjust = 0),
        axis.ticks.y = element_blank(),
        legend.position = "none")
```

### Survival analysis

```{r}
data(AustinCats, package = "rethinking")

data_cats <- AustinCats %>% 
  as_tibble() %>% 
  mutate(black = ifelse(color == "Black", "black", "other"),
         adopted  = ifelse(out_event == "Adoption", 1, 0),
         censored = ifelse(out_event != "Adoption", 1, 0))

data_cats %>% 
  count(color) %>%
  arrange(-n) %>% 
  slice(1:10)
```

```{r, fig.asp = .35}
data_cats %>% 
  count(black) %>% 
  mutate(percent = 100 * n / sum(n)) %>% 
  mutate(label = str_c(round(percent, digits = 1), "%")) %>% 
  ggplot(aes(y = black)) +
  geom_col(aes(x = n, color = black, fill = after_scale(clr_alpha(color,.65))), size = .5) +
  geom_text(aes(x = n - 250, label = label),
            color = "white", family = fnt_sel, hjust = 1.1) +
  scale_color_manual(values = c("black", clr0dd), breaks = NULL) +
  scale_x_continuous(expression(italic(n)), breaks = c(0, count(data_cats, black) %>%
                                                         pull(n))) +
  labs(title = "Cat color",
       y = NULL) +
  theme(axis.ticks.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank())
```

```{r}
data_cats %>% 
  count(out_event)
```

```{r}
data_cats %>% 
  mutate(censored = factor(censored)) %>% 
  filter(days_to_event < 300) %>% 
  
  ggplot(aes(x = days_to_event, y = censored)) +
  # let's just mark off the 50% intervals
  stat_halfeye(.width = .5, fill =  fll0dd, height = 4) +
  scale_y_discrete(NULL, labels = c("censored == 0", "censored == 1")) +
  coord_cartesian(ylim = c(1.5, 2.6)) +
  theme(axis.ticks.y = element_blank())
```

> McElreath fit his survival model using the exponential likelihood. [...] If we let $y$ be a non-negative continuous variable, the probability density function for the exponential distribution is

$$
f(y) = \lambda e^{-\lambda y}
$$

> where $\lambda$ is called the rate. The mean of the exponential distribution is the inverse of the rate:

$$
E[y] =\frac{1}{\lambda}
$$

> we can write our continuous-time survival model as:

$$
\begin{align*}
\text{days_to_event}_i | \text{censored}_i = 0 & \sim \operatorname{Exponential}(\lambda_i) \\
\text{days_to_event}_i | \text{censored}_i = 1 & \sim \operatorname{Exponential-CCDF}(\lambda_i) \\
\lambda_i & = 1 / \mu_i \\
\log \mu_i & = \alpha_{\text{black}[i]} \\
\alpha & \sim \operatorname{Normal}(0, 1).
\end{align*}
$$

> When you fit a continuous-time survival analysis with `brm()`, youll want to tell the software about how the data have been censored with help from the `cens()` function. For many of the models in this chapter, we used the `trials()` function to include the $n_{i}$ information into our binomial models. Both `trials()` and `cens()` are members of a class of functions designed to provide supplemental information about our criterion variables to `brm()`. The `cens()` function lets us add in information about censoring. [...] We will feed this information into the model with the formula code `days_to_event | cens(censored)`, where `censored` is the name of the variable in our data that indexes the censoring.

```{r}
brms_c11_model_cats <- brm(
  data = data_cats,
  family = exponential,
  days_to_event | cens(censored) ~ 0 + black,
  prior(normal(0, 1), class = b),
  iter = 2000, warmup = 1000, chains = 4, cores = 4,
  seed = 42,
  file = "brms/brms_c11_model_cats")
```

> Since we modeled $\log \mu_{i}$, we need to transform our $\alpha$ parameters back into the $\lambda$ metric using the formula:

$$
\begin{align*}
\log \mu             & = \alpha_\text{black}, && \text{and} \\
\lambda              & = 1 / \mu,             && \text{therefore} \\
\lambda_\text{black} & = 1 / \exp(\alpha_\text{black}).
\end{align*}
$$

```{r}
(1 / exp(fixef(brms_c11_model_cats)[, -2])) %>% 
  as.data.frame() %>% 
  knitr::kable()
```

```{r}
cats_fixed <- fixef(brms_c11_model_cats) %>% 
  data.frame() %>% 
  rownames_to_column() %>% 
  as_tibble() %>% 
  mutate(color = str_remove(rowname, "black")) %>% 
  expand(nesting(Estimate, Q2.5, Q97.5, color),
         days = 0:100) %>% 
  mutate(m  = 1 - pexp(days, rate = 1 / exp(Estimate)),
         ll = 1 - pexp(days, rate = 1 / exp(Q2.5)),
         ul = 1 - pexp(days, rate = 1 / exp(Q97.5)))
  
p1 <- cats_fixed %>% 
  ggplot(aes(x = days)) +
  geom_hline(yintercept = .5, linetype = 3, color = clr_dark) +
  geom_smooth(aes(ymin = ll, y = m, ymax = ul, 
                  color = color, fill = after_scale(clr_alpha(color))),
              stat = "identity", size = .4) +
  scale_color_manual(values = c(black = "black", other = clr0d)) +
  scale_y_continuous("proportion remaining", breaks = c(0, .5, 1), limits = 0:1) +
  xlab("days to adoption")
```

>  About how many days would it take for half of the cats of a given color to be adopted?

```{r}
qexp(p = .5, rate = 1 / exp(fixef(brms_c11_model_cats)[1, 1]))
```


```{r}
cats_posterior <- as_draws_df(brms_c11_model_cats) %>% 
  pivot_longer(starts_with("b_")) %>% 
  mutate(color = str_remove(name, "b_black"),
         days  = qexp(p = .5, rate = 1 / exp(value))) 

# axis breaks
cats_medians <-
  group_by(cats_posterior, color) %>% 
  summarise(med = median(days)) %>% 
  pull(med) %>% 
  round(., digits = 1)

p2 <- cats_posterior %>% 
  ggplot(aes(x = days, y = color)) +
  stat_halfeye(.width = .95, aes(fill = color), height = 4) +
  scale_x_continuous("days untill 50% are adopted",
                     breaks = c(30, cats_medians, 45), labels = c("30", cats_medians, "45"),
                     limits = c(30, 45)) +
  scale_fill_manual(values = c("black", clr0d) %>% 
                      clr_alpha() %>% 
                      set_names(nm = c("black", "other")),
                    guide = "none") +
  ylab(NULL) +
  coord_cartesian(ylim = c(1.5, 5.1)) +
  theme(axis.ticks.y = element_blank(),
        panel.grid.minor.x = element_blank())

p1 + p2 +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```


## pymc3 section

---

<div id="myModal" class="modal">
  <span class="close">&times;</span>
  <img class="modal-content" id="img01">
  <div id="caption"></div>
</div>

<script src="./js/zoom.js"></script>