---
output:
  html_document:
  theme: yeti
pdf_document: default
editor_options: 
  chunk_output_type: console
---

# Rethinking: Chapter 5

**Spurious waffles**

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = "#>", 
                      dev = "svg",         # for html rendering
                      # dev = "cairo_pdf", # for pdf rendering
                      fig.asp = .5, 
                      fig.align = "center",
                      message = FALSE,
                      warning = FALSE)

source("bayesian_settings.R")
source("knitr_matrix.R")
```

by [Richard McElreath](https://xcelab.net/rm/statistical-rethinking/), building on the Summaries by [Solomon Kurz](https://bookdown.org/content/4857/) and [Jake Thompson](https://sr2-solutions.wjakethompson.com/linear-models.html).

```{r}
library(sf)
library(rethinking)
library(ggfx)

data(WaffleDivorce)

WaffleDivorce <- WaffleDivorce %>% as_tibble()

usa <- read_sf("~/work/geo_store/USA/usa_states_albers_revised.gpkg") %>% 
  left_join(WaffleDivorce, by = c(name = "Location" ))

p_waffle <- usa %>% 
  ggplot(aes(fill = WaffleHouses / Population)) +
  scale_fill_gradientn(colours = c(clr0d, clr2) %>%
                         clr_lighten(.3))

p_divorce <- usa %>% 
  ggplot(aes(fill = Divorce))+
  scale_fill_gradientn(colours = c(clr0d, clr1) %>%
                         clr_lighten(.3))

p_age <- usa %>% 
  ggplot(aes(fill = MedianAgeMarriage))+
  scale_fill_gradientn(colours = c(clr_lighten(clr0d, .3), clr3))
```

```{r, fig.asp = .38}
p_waffle +
  p_divorce +
  p_age +
  plot_layout(guides = "collect") & 
  with_shadow(geom_sf(aes(color = after_scale(clr_darken(fill)))),
              x_offset = 0, y_offset = 0, sigma = 3) & 
  guides(fill = guide_colorbar(title.position = "top",
                               barheight = unit(5,"pt"))) &
  theme(legend.position = "bottom")
```

Age Model: first model (divorce rate depends on age at marriage)

$$
\begin{array}{cccr} 
D_i & {\sim} & Normal(\mu, \sigma) & \textrm{[likelihood]}\\
\mu_i & = & \alpha + \beta_{A} A_{i} & \textrm{[linear model]}\\
\alpha & \sim & Normal(0, 0.2) & \textrm{[$\alpha$ prior]}\\
\beta_{A} & \sim & Normal(0, 0.5) & \textrm{[$\beta$ prior]}\\
\sigma & \sim & Exponential(1) & \textrm{[$\sigma$ prior]}
\end{array}
$$

```{r}
data_waffle <- WaffleDivorce %>% 
  mutate(across(.cols = c(Divorce, Marriage, MedianAgeMarriage),
                .fns = standardize,
                .names = "{str_to_lower(.col)}_std"),
         waffle_pop = WaffleHouses / Population) %>% 
  rename(median_age_std = "medianagemarriage_std") 

sd(data_waffle$MedianAgeMarriage)

model_age <- quap(
  flist = alist(
    divorce_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_A * median_age_std ,
    alpha ~ dnorm( 0, 0.2 ),
    beta_A ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_waffle
)

set.seed(10)
age_priors <- extract.prior(model_age) %>% 
  as_tibble()

prior_prediction_range <- c(-2, 2)
age_prior_predictions <- link(model_age,
                        post = age_priors,
                        data = list(median_age_std = prior_prediction_range)) %>% 
  as_tibble() %>% 
  set_names(nm = as.character(prior_prediction_range)) %>% 
  mutate(.draw = row_number())

age_prior_predictions %>% 
  filter(.draw < 51) %>% 
  ggplot() +
  geom_segment(aes(x = -2, xend = 2, y = `-2`, yend = `2`, group = .draw),
               color = clr2, alpha = .2) +
  labs(x = "median age of marriage (std)", 
       y = "divorce rate (std)")
```

```{r}
age_seq <- seq(min(data_waffle$median_age_std),
               max(data_waffle$median_age_std),
               length.out = 101)
model_age_posterior_prediction_samples <- link(model_age, data = data.frame(median_age_std = age_seq)) %>% 
  as_tibble() %>% 
  set_names(nm = age_seq) %>% 
  pivot_longer(cols = everything(), names_to = "median_age_std", values_to = "divorce_std") %>% 
  mutate(median_age_std = as.numeric(median_age_std),
         MedianAgeMarriage = median_age_std * sd(data_waffle$MedianAgeMarriage) +
           mean(data_waffle$MedianAgeMarriage),
         Divorce = divorce_std * sd(data_waffle$Divorce) +
           mean(data_waffle$Divorce)) 

model_age_posterior_prediction_pi <- model_age_posterior_prediction_samples %>% 
  group_by(median_age_std, MedianAgeMarriage) %>% 
  summarise(mean = mean(Divorce),
            PI_lower = PI(Divorce)[1],
            PI_upper = PI(Divorce)[2]) %>% 
  ungroup()

model_age_posterior_prediction_simulation <- sim(model_age,
                                                 data = data.frame(median_age_std = age_seq), 
                                                 n = 5e3) %>% 
  as_tibble() %>% 
  set_names(nm = age_seq) %>% 
  pivot_longer(cols = everything(), names_to = "median_age_std", values_to = "divorce_std") %>% 
  mutate(median_age_std = as.numeric(median_age_std),
         MedianAgeMarriage = median_age_std * sd(data_waffle$MedianAgeMarriage) +
           mean(data_waffle$MedianAgeMarriage),
         Divorce = divorce_std * sd(data_waffle$Divorce) +
           mean(data_waffle$Divorce))
  
model_age_posterior_prediction_simulation_pi <- model_age_posterior_prediction_simulation %>% 
  group_by(median_age_std, MedianAgeMarriage) %>% 
  summarise(mean = mean(Divorce),
            PI_lower = PI(Divorce)[1],
            PI_upper = PI(Divorce)[2]) %>% 
  ungroup()

p_age <- ggplot(mapping = aes(x = MedianAgeMarriage)) +
  geom_ribbon(data = model_age_posterior_prediction_simulation_pi,
              aes(ymin = PI_lower, ymax = PI_upper), fill = clr0d, alpha = .35)  +
  geom_smooth(data = model_age_posterior_prediction_pi, stat = "identity",
              aes(y = mean, ymin = PI_lower, ymax = PI_upper),
              color = clr3, fill = fll3, size = .4) +
  geom_point(data = data_waffle, aes(y = Divorce), color = rgb(0,0,0,.5), size = .6) +
  labs(y = " divorce")
```

Marriage Model: alternative model (divorce rate depends on marriage rate)

$$
\begin{array}{cccr} 
D_i & {\sim} & Normal(\mu, \sigma) & \textrm{[likelihood]}\\
\mu_i & = & \alpha + \beta_{M} M_{i} & \textrm{[linear model]}\\
\alpha & \sim & Normal(0, 0.2) & \textrm{[$\alpha$ prior]}\\
\beta_{M} & \sim & Normal(0, 0.5) & \textrm{[$\beta$ prior]}\\
\sigma & \sim & Exponential(1) & \textrm{[$\sigma$ prior]}
\end{array}
$$

```{r}
model_marriage <- quap(
  flist = alist(
    divorce_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_M * marriage_std ,
    alpha ~ dnorm( 0, 0.2 ),
    beta_M ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_waffle
)
```

```{r}
marriage_seq <- seq(min(data_waffle$marriage_std),
               max(data_waffle$marriage_std),
               length.out = 101)
model_marriage_posterior_prediction_samples <- link(model_marriage,
                                               data = data.frame(marriage_std = marriage_seq)) %>% 
  as_tibble() %>% 
  set_names(nm = marriage_seq) %>% 
  pivot_longer(cols = everything(), names_to = "marriage_std", values_to = "divorce_std") %>% 
  mutate(marriage_std = as.numeric(marriage_std),
         Marriage = marriage_std * sd(data_waffle$Marriage) +
           mean(data_waffle$Marriage),
         Divorce = divorce_std * sd(data_waffle$Divorce) +
           mean(data_waffle$Divorce)) 

model_marriage_posterior_prediction_pi <- model_marriage_posterior_prediction_samples %>% 
  group_by(marriage_std, Marriage) %>% 
  summarise(mean = mean(Divorce),
            PI_lower = PI(Divorce)[1],
            PI_upper = PI(Divorce)[2]) %>% 
  ungroup()

model_marriage_posterior_prediction_simulation <- sim(model_marriage,
                                                 data = data.frame(marriage_std = marriage_seq), 
                                                 n = 5e3) %>% 
  as_tibble() %>% 
  set_names(nm = marriage_seq) %>% 
  pivot_longer(cols = everything(), names_to = "marriage_std", values_to = "divorce_std") %>% 
  mutate(marriage_std = as.numeric(marriage_std),
         Marriage = marriage_std * sd(data_waffle$Marriage) +
           mean(data_waffle$Marriage),
         Divorce = divorce_std * sd(data_waffle$Divorce) +
           mean(data_waffle$Divorce))
  
model_marriage_posterior_prediction_simulation_pi <- model_marriage_posterior_prediction_simulation %>% 
  group_by(marriage_std, Marriage) %>% 
  summarise(mean = mean(Divorce),
            PI_lower = PI(Divorce)[1],
            PI_upper = PI(Divorce)[2]) %>% 
  ungroup()

p_marriage <- ggplot(mapping = aes(x = Marriage)) +
  geom_ribbon(data = model_marriage_posterior_prediction_simulation_pi,
              aes(ymin = PI_lower, ymax = PI_upper), fill = clr0d, alpha = .35)  +
  geom_smooth(data = model_marriage_posterior_prediction_pi, stat = "identity",
              aes(y = mean, ymin = PI_lower, ymax = PI_upper),
              color = clr1, fill = fll1, size = .2) +
  geom_point(data = data_waffle, aes(y = Divorce), color = rgb(0,0,0,.5), size = .6) +
  labs(y = " divorce")
```

Waffle Model:

```{r}
model_waffle <- quap(
  flist = alist(
    divorce_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_W * waffle_pop ,
    alpha ~ dnorm( 0, 0.2 ),
    beta_W ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_waffle
)

waffle_seq <- seq(min(data_waffle$waffle_pop),
               max(data_waffle$waffle_pop),
               length.out = 101)
model_waffle_posterior_prediction_samples <- link(model_waffle,
                                               data = data.frame(waffle_pop = waffle_seq)) %>% 
  as_tibble() %>% 
  set_names(nm = waffle_seq) %>% 
  pivot_longer(cols = everything(), names_to = "waffle_pop", values_to = "divorce_std") %>% 
  mutate(waffle_pop = as.numeric(waffle_pop),
         Divorce = divorce_std * sd(data_waffle$Divorce) +
           mean(data_waffle$Divorce)) 

model_waffle_posterior_prediction_pi <- model_waffle_posterior_prediction_samples %>% 
  group_by(waffle_pop) %>% 
  summarise(mean = mean(Divorce),
            PI_lower = PI(Divorce)[1],
            PI_upper = PI(Divorce)[2]) %>% 
  ungroup()

p_waffle <- ggplot(mapping = aes(x = waffle_pop)) +
  geom_smooth(data = model_waffle_posterior_prediction_pi, stat = "identity",
              aes(y = mean, ymin = PI_lower, ymax = PI_upper),
              color = clr2, fill = fll2, size = .2) +
  geom_point(data = data_waffle, aes(y = Divorce), color = rgb(0,0,0,.5), size = .6) +
  labs(y = " divorce")
```

```{r, fig.asp = .4}
p_waffle + 
p_marriage + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
  p_age + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) &
  lims(y = c(4, 15))
```

## Directed Acyclic Graphs

```{r, fig.asp = .4}
dag1 <- dagify(
  D ~ A + M,
  M ~ A,
  exposure = "A",
  outcome = "M") %>% 
  tidy_dagitty(.dagitty = .,layout = tibble(x = c(0,1,.5), y = c(1,1, .4))) %>%
  mutate(stage = if_else(name == "D", "response",
                         if_else(name %in% c("A", "M"),
                                 "predictor", "confounds")))

dag2 <- dagify(
  D ~ A,
  M ~ A,
  exposure = "A",
  outcome = "M") %>% 
  tidy_dagitty(.dagitty = .,layout = tibble(x = c(0,.5,1), y = c(1, .4, 1))) %>%
  mutate(stage = if_else(name == "D", "response",
                         if_else(name %in% c("A", "M"),
                                 "predictor", "confounds")))

plot_dag(dag1, clr_in = clr3) + 
  plot_dag(dag2, clr_in = clr3) &
  scale_y_continuous(limits = c(.35, 1.05)) &
  coord_equal()
```

DAG notation:

- $Y \perp \!\!\! \perp X | Z$: *"$Y$ is independent of $X$ conditional on $Z$"*
- $D \not\!\perp\!\!\!\perp A$: "*$D$ is associated with $A$"*

Check pair wise correlations with `cor()`:

```{r}
data_waffle %>% 
  dplyr::select(divorce_std,marriage_std, median_age_std) %>% 
  cor() %>% 
  as.data.frame(row.names = row.names(.)) %>% 
  round(digits = 2) %>% 
  knitr::kable()
```

```{r}
library(dagitty)

dagitty('dag{ D <- A -> M -> D}') %>% 
  impliedConditionalIndependencies()

dagitty('dag{ D <- A -> M }') %>% 
  impliedConditionalIndependencies()
```

## Multiple Regression notion

$$
\begin{array}{cccr} 
D_i & {\sim} & Normal(\mu, \sigma) & \textrm{[likelihood]}\\
\mu_i & = & \alpha + \beta_{M} M_{i} + \beta_{A} A_{i} & \textrm{[linear model]}\\
\alpha & \sim & Normal(0, 0.2) & \textrm{[$\alpha$ prior]}\\
\beta_{M} & \sim & Normal(0, 0.5) & \textrm{[$\beta_M$ prior]}\\
\beta_{A} & \sim & Normal(0, 0.5) & \textrm{[$\beta_A$ prior]}\\
\sigma & \sim & Exponential(1) & \textrm{[$\sigma$ prior]}
\end{array}
$$

or compact notion 

$$
\mu_i = \alpha + \sum_{j = 1}^{n} \beta_jx_{ji}
$$

or even *matrix* notion

$$
m = Xb
$$

```{r}
model_multiple <- quap(
  flist = alist(
      divorce_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_M * marriage_std + beta_A * median_age_std,
    alpha ~ dnorm( 0, 0.2 ),
    beta_A ~ dnorm( 0, 0.5 ),
    beta_M ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_waffle
)

precis(model_multiple) %>% 
  round(digits = 2) %>% 
  as.matrix() %>% 
  knitr::kable()
```

```{r, fig.asp = .3}
ct <- coeftab(model_age, model_marriage, model_multiple,se = TRUE)
plot_coeftab(ct)
```

`beta_A` doesn't really change, it only grows more uncertain, yet `beta_M` is only associated with divorce, when marriage rate is missing from the model.

> "Once we know the median age at marriage for a State, there is little to no additional predictive power in also knowing the rate of marriage at that State."

$\rightarrow$ $D \perp \!\!\! \perp M | A$

simulating the divorcee example

```{r}
n <- 50
data_divorce_sim <- tibble(median_age_std = rnorm(n),
                           marriage_std = rnorm(n, mean = -median_age_std),
                           divorce_std = rnorm(n, mean = median_age_std),
                           divorce_codep = rnorm(n, mean = median_age_std + marriage_std))

p1 <- ggpairs(data_divorce_sim %>% dplyr::select(-divorce_codep),
        lower = list(continuous = wrap(ggally_points, colour = clr1, size = .5, alpha = .7)),
        diag = list(continuous = wrap("densityDiag", fill = fll1, color = clr1, adjust = 1)),
        upper = list(continuous = wrap(ggally_cor, size = 5, color = "black", family = "Josefin sans")))
p2 <- ggpairs(data_divorce_sim %>% dplyr::select(-divorce_std),
        lower = list(continuous = wrap(ggally_points, colour = clr2, size = .5, alpha = .7)),
        diag = list(continuous = wrap("densityDiag", fill = fll2, color = clr2, adjust = 1)),
        upper = list(continuous = wrap(ggally_cor, size = 5, color = "black", family = "Josefin sans")))

cowplot::plot_grid(ggmatrix_gtable(p1), ggmatrix_gtable(p2))
```

simulating the right DAG ($D \perp \!\!\! \perp M | A$)

```{r}
model_multiple_sim <- quap(
  flist = alist(
      divorce_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_M * marriage_std + beta_A * median_age_std,
    alpha ~ dnorm( 0, 0.2 ),
    beta_A ~ dnorm( 0, 0.5 ),
    beta_M ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_divorce_sim
)

model_age_sim <- quap(
  flist = alist(
      divorce_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_A * median_age_std,
    alpha ~ dnorm( 0, 0.2 ),
    beta_A ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_divorce_sim
)

model_marriage_sim <- quap(
  flist = alist(
      divorce_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_M * marriage_std,
    alpha ~ dnorm( 0, 0.2 ),
    beta_M ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_divorce_sim
)
```

```{r, fig.asp = .3}
ct_sim <- coeftab(model_age_sim, model_marriage_sim, model_multiple_sim, se = TRUE)
plot_coeftab(ct_sim)
```

simulating the left DAG ($D \not\!\perp\!\!\!\perp M | A$)

```{r}
model_multiple_sim_codep <- quap(
  flist = alist(
      divorce_codep ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_M * marriage_std + beta_A * median_age_std,
    alpha ~ dnorm( 0, 0.2 ),
    beta_A ~ dnorm( 0, 0.5 ),
    beta_M ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_divorce_sim
)

model_age_sim_codep <- quap(
  flist = alist(
      divorce_codep ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_A * median_age_std,
    alpha ~ dnorm( 0, 0.2 ),
    beta_A ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_divorce_sim
)

model_marriage_sim_codep <- quap(
  flist = alist(
      divorce_codep ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_M * marriage_std,
    alpha ~ dnorm( 0, 0.2 ),
    beta_M ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_divorce_sim
)
```


```{r, fig.asp = .3}
ct_sim_codep <- coeftab(model_age_sim_codep, model_marriage_sim_codep, model_multiple_sim_codep,
                        se = TRUE)
plot_coeftab(ct_sim_codep)
```


### Visualizations for multivariate regressions

- *Predictor residual plots.* useful for understanding the model, but not much else
- *Posterior prediction plots.* checking fit and assessing predictions
- *Counterfactual plots.* implied predictions for imaginary experiments

#### Predictor residual plots

predictor residual plot for *marriage rate*

```{r}
pred_res_marriage <- quap(
  flist = alist(
    marriage_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_AM * median_age_std,
    alpha ~ dnorm( 0, 0.2 ),
    beta_AM ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_waffle
)

residuals_marriage <- link(pred_res_marriage) %>% 
  as_tibble() %>% 
  set_names(nm = seq_along(data_waffle$median_age_std)) %>% 
  pivot_longer(cols = everything(),
               names_to = "row_idx", 
               values_to = "fit_marriage") %>%
  group_by(row_idx) %>% 
  summarise(mean_marriage = mean(fit_marriage),
            lower_pi = PI(fit_marriage)[1],
            upper_pi = PI(fit_marriage)[2]) %>% 
  ungroup() %>% 
  mutate(row_idx = as.numeric(row_idx)) %>% 
  left_join(data_waffle %>%  mutate(row_idx = row_number()),. ) %>% 
  mutate(residual_marriage  = marriage_std - mean_marriage)

p_11 <- residuals_marriage %>% 
  ggplot(aes(x = median_age_std)) +
  geom_segment(aes(xend = median_age_std, y = mean_marriage, yend = marriage_std),
               color = rgb(0,0,0,.6), linetype = 3) +
  geom_line(aes(y = mean_marriage), color = clr1) +
  geom_point(aes(y = marriage_std),
             color = clr1, fill = clr_lighten(clr1, .35), shape = 21) +
  geom_text(data = residuals_marriage %>% filter(Loc %in% c("DC", "HI", "ND", "ME", "WY")),
            aes(x = median_age_std - .1, y = marriage_std, label = Loc), hjust = 1) 

pred_res_marriage_mu <- quap(
  flist = alist(
    divorce_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta * residual_marriage,
    alpha ~ dnorm( 0, 0.2 ),
    beta ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = residuals_marriage
)

seq_res <- seq(min(residuals_marriage$residual_marriage), max(residuals_marriage$residual_marriage), length.out = 101)

residual_lm_posterior <- link(pred_res_marriage_mu, data = data.frame(residual_marriage = seq_res)) %>% 
  as_tibble() %>% 
  set_names(nm = seq_res) %>% 
  pivot_longer(cols = everything(), names_to = "residual_marriage", values_to = "divorce_std") %>% 
  mutate(residual_marriage = as.numeric(residual_marriage)) %>% 
  group_by(residual_marriage) %>% 
  summarise(mean = mean(divorce_std),
            PI_lower = PI(divorce_std)[1],
            PI_upper = PI(divorce_std)[2]) %>% 
  ungroup()

p_12 <- ggplot(mapping = aes(x = residual_marriage)) +
  geom_vline(xintercept = 0, lty = 3, color = rgb(0,0,0,.6)) +
  geom_smooth(data = residual_lm_posterior, aes(y = mean, ymin = PI_lower, ymax = PI_upper),
              stat = "identity", color = clr1, fill = fll1, size = .4) +
  geom_point(data = residuals_marriage, aes(y = divorce_std),
             color = clr1, fill = clr_lighten(clr1,.35), shape = 21) +
  geom_text(data = residuals_marriage %>% filter(Loc %in% c("DC", "HI", "ND", "ME", "WY")),
            aes(y = divorce_std - .2, label = Loc)) +
  labs(y = "divorce_rate (std)")
```

predictor residual plot for *age at marriage*

```{r}
pred_res_age <- quap(
  flist = alist(
    median_age_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta_MA * marriage_std,
    alpha ~ dnorm( 0, 0.2 ),
    beta_MA ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = data_waffle
)

residuals_age <- link(pred_res_age) %>% 
  as_tibble() %>% 
  set_names(nm = seq_along(data_waffle$marriage_std)) %>% 
  pivot_longer(cols = everything(),
               names_to = "row_idx", 
               values_to = "fit_age") %>%
  group_by(row_idx) %>% 
  summarise(mean_age = mean(fit_age),
            lower_pi = PI(fit_age)[1],
            upper_pi = PI(fit_age)[2]) %>% 
  ungroup() %>% 
  mutate(row_idx = as.numeric(row_idx)) %>% 
  left_join(data_waffle %>%  mutate(row_idx = row_number()),. ) %>% 
  mutate(residual_age  = median_age_std - mean_age)

p_21 <- residuals_age %>% 
  ggplot(aes(x = marriage_std)) +
  geom_segment(aes(xend = marriage_std, y = mean_age, yend = median_age_std),
               color = rgb(0,0,0,.6), linetype = 3) +
  geom_line(aes(y = mean_age), color = clr2) +
  geom_point(aes(y = median_age_std),
             color = clr2, fill = clr_lighten(clr2, .35), shape = 21) +
  geom_text(data = residuals_marriage %>% filter(Loc %in% c("DC", "HI", "ID")),
            aes(x = marriage_std - .1, y = median_age_std, label = Loc), hjust = 1) 

pred_res_age_mu <- quap(
  flist = alist(
    divorce_std ~ dnorm( mu, sigma ) ,
    mu <- alpha + beta * residual_age,
    alpha ~ dnorm( 0, 0.2 ),
    beta ~ dnorm( 0, 0.5 ),
    sigma ~ dexp( 1 )
  ),
  data = residuals_age
)

seq_res_age <- seq(min(residuals_age$residual_age), max(residuals_age$residual_age), length.out = 101)

residual_lm_posterior_age <- link(pred_res_age_mu, data = data.frame(residual_age = seq_res_age)) %>% 
  as_tibble() %>% 
  set_names(nm = seq_res_age) %>% 
  pivot_longer(cols = everything(), names_to = "residual_age", values_to = "divorce_std") %>% 
  mutate(residual_age = as.numeric(residual_age)) %>% 
  group_by(residual_age) %>% 
  summarise(mean = mean(divorce_std),
            PI_lower = PI(divorce_std)[1],
            PI_upper = PI(divorce_std)[2]) %>% 
  ungroup()

p_22 <- ggplot(mapping = aes(x = residual_age)) +
  geom_vline(xintercept = 0, lty = 3, color = rgb(0,0,0,.6)) +
  geom_smooth(data = residual_lm_posterior_age, aes(y = mean, ymin = PI_lower, ymax = PI_upper),
              stat = "identity", color = clr2, fill = fll2, size = .4) +
  geom_point(data = residuals_age, aes(y = divorce_std),
             color = clr2, fill = clr_lighten(clr2,.35), shape = 21) +
  geom_text(data = residuals_age %>% filter(Loc %in% c("DC", "HI", "ID")),
            aes(y = divorce_std - .2, label = Loc)) +
  labs(y = "divorce_rate (std)")
```

```{r, fig.asp = .7}
p_11 + p_21 +
  p_12 + p_22
```

#### Posterior Preediction Plots

```{r}
posterior_prediction <- link(model_multiple) %>% 
  as_tibble() %>% 
    set_names(nm = seq_along(data_waffle$divorce_std)) %>% 
  pivot_longer(cols = everything(),
               names_to = "row_idx", 
               values_to = "divorce_predicted") %>%
  group_by(row_idx) %>% 
  summarise(divorce_predicted_mean = mean(divorce_predicted),
            lower_pi = PI(divorce_predicted)[1],
            upper_pi = PI(divorce_predicted)[2]) %>% 
  ungroup() %>% 
  mutate(row_idx = as.numeric(row_idx)) %>% 
  left_join(data_waffle %>%  mutate(row_idx = row_number()), . ) 

posterior_simmulation <- sim(model_multiple) %>% 
  as_tibble() %>% 
    set_names(nm = seq_along(data_waffle$divorce_std)) %>% 
  pivot_longer(cols = everything(),
               names_to = "row_idx", 
               values_to = "divorce_predicted") %>%
  group_by(row_idx) %>% 
  summarise(lower_pi = PI(divorce_predicted)[1],
            upper_pi = PI(divorce_predicted)[2]) %>% 
  ungroup() %>% 
  mutate(row_idx = as.numeric(row_idx)) %>% 
  left_join(data_waffle %>%  mutate(row_idx = row_number()), . ) 

ggplot(mapping = aes(x = divorce_std)) +
  geom_abline(slope = 1, size = .7, lty = 3, color = rgb(0,0,0,.6)) +
  geom_linerange(data = posterior_prediction,
                aes(ymin = lower_pi, ymax =  upper_pi), color = clr2)+
  geom_point(data = posterior_prediction, aes(y = divorce_predicted_mean), 
             color = clr2, fill = clr_lighten(clr2,.6), shape = 21, size = 1.5)+
  geom_text(data = posterior_prediction %>% filter(Loc %in% c("ID", "ME", "RI", "UT")),
            aes(x = divorce_std - .15, y = divorce_predicted_mean, label = Loc)) 
  
```


## Homework

**E1**

**E2**

**E3**

**E4**

**E5**

**E6**

**E7**

**M1**

**M2**

**M3**

**M4**

**M5**

**M6**

**H1**

**H2**

**H3**

**H4**

**H5**


## {brms} section

---

<div id="myModal" class="modal">
  <span class="close">&times;</span>
  <img class="modal-content" id="img01">
  <div id="caption"></div>
</div>

<script src="./js/zoom.js"></script>